
\chapter{Actions}
\label{ch:actions}

Historically, groups have appeared because they can ``act'' on a set
(or more general objects), that is to say, they collect some of the
symmetries of the set. This is a point of view that we will return to
many times and we give the basic theory in \cref{sec:gsets}.
This section should remind the reader of the material in \cref{cha:circle},
where we dealt with the special case of the group of integers.
More generally, connected \coverings now reappear in the guise of
``transitive $G$-sets'', and these are intimately related to
the set of subgroups of a group.
These also generalize the bicycles of~\cref{sec:bicycles},
from which we lift the notion of ``normality''.

A central notion in this chapter is that of ``$G$-torsor'', introduced in
\cref{sec:gsets}. The main result about $G$-torsors is that their type can be
identified with the classifying type of the group $G$. In other words, the
group $G$ can be completely recovered from its torsors. This idea will be
essential in~\cref{ch:absgroup} where we show that our definition of concrete
groups is equivalent to the abstract version taught in most algebra classes.


\section{Brief overview of the chapter}

After setting things up in~\cref{sec:gsets}, and
studying subgroups in~\cref{sec:subgroups},
we introduce the important
operations of taking \emph{invariant maps} and \emph{orbits} of an action
in~\cref{sec:fixpts-orbits}.
The fundamental equivalence between the classifying type $\BG$ of a group $G$
and the type of $G$-torsors is constructed in~\cref{sec:torsors}. 
In \cref{sec:groupssubperm} we apply $G$-torsors to prove Cayley's Theorem
for our groups, and in \cref{sec:burnsides-lemma} we
begin the study of the combinatorics of group actions.
This allows us to count, for instance, how many ways there are of 
``coloring'' objects acted on by groups,
and it lays the groundwork for the combinatorics of finite groups
we'll be looking at in~\cref{ch:fingp}.

\section{Group actions ($G$-sets)}
\label{sec:gsets}

One of the most important features of groups: they can \emph{act} on things
(giving rise to manifestations of symmetries)! For example, in the circle $\Sc$,
the loop $\Sloop$ is a non-trivial transformation of the base point
$\base$. This can \emph{act} on other mathematical entities: Take for example
the $2$-fold \covering of $\Sc$ whose fiber at $\base$ is $\bn 2$. The symmetry
$\Sloop$ acts on its two elements by swapping them. The following definition
decribes that phenomenon for arbitrary groups.

%\MB{Before we handle the more complex case of abstract groups,
%let us see what this looks like for groups. Leave out?}

\begin{definition}\label{def:Gset}
  For $G$ a group, a \emph{$G$-set} is a function\footnote{Recall that, to make
    things more readable, we write $\BG\to\Set$ instead of the more correct
    $\BG_\div\to\Set$.}  \index{group!acting on a set}\index{group action!of
    $G$-set}
  \[
    X : \BG\to\Set,
  \]
  and $X(\sh_G)$ is referred to as the \emph{underlying set}.  If $p:x\eqto y$
  in $\BG$, then the transport function $\trp{X}(p) : X(x)\equivto X(y)$ is also
  denoted by $X(p)$.  We denote $X(p)(a)$ by $p\cdot_X a$.  The operation
  $\cdot_X$ is called the \emph{group action} of $X$.  When $X$ is clear from
  the context we may leave out the subscript $X$.\footnote{%
    The reader might worry that this notation conflicts with that of path
    composition. The type of the operation
    ${}\cdot{}: (x\eqto y) \to X(x) \to X(y)$ helps identify which is which.
    However, see \cref{def:principaltorsor} for a special case where both
    conflates (i.e., where $\cdot_X$ is indeed path composition).}  In
  particular, if $g:\USymG$, then $X(g)$ is a permutation of the underlying set
  $X(\sh_G)$ of $X$.

  The type of $G$-sets is\index{GSet@$G$-set (of group)}
  \glossary(GSet){\protect{$\GSet$}}{type of $G$-sets}
  \[
    \GSet\defequi(\BG\to\Set).\qedhere
  \]
\end{definition}
\marginnote{
  Much of what follows will work equally well for $\infty$-groups;
  if $G$ is (a group or) an infinity group,
  a \emph{$G$-type} is a function $X : \BG\to\UU$,
  with \emph{underlying type} $X(\sh_G)$.
  This is an \emph{action in $\UU$}, and
  more generally, an action of $G$ on an element of type $A$
  is a function $X : \BG\to A$, see~\cref{sec:actions} below.}

\begin{example}\label{def:trivGset}
  If $G$ is a group and $X$ is a set, then $\triv_G X$ defined by
  \[\triv_G X(z)\defequi X, \quad\text{for all $z:\BG$,}\]
  is a $G$-set.
  Examples of this sort (regardless of $X$) are called \emph{trivial $G$-sets}.
\end{example}
\begin{remark}
  \label{rem:G-set-vs-set-bundle}
The reader may have noticed that the type of $G$-sets is equivalent to the
type of \coverings over $\BG$.
The reason we have allowed ourselves two names is that our focus is different: for a $G$-set $X:\BG\to\Set$ we focus on the sets $X(z)$, whereas when talking about \coverings the first projection $\sum_{z:\BG}X(z)\to \BG$ takes center stage.  Each focus has its advantages.
\end{remark}

\begin{example}\label{def:principaltorsor}
  If $G$ is a group, then
  \[
    \princ G:\BG\to\Set,
    \qquad\princ G(z)\defequi(\sh_G \eqto z)
  \]
  is a $G$-set called the \emph{principal $G$-torsor}.\footnote{%
    The term ``$G$-torsor'' will reappear several times and will mean nothing but a $G$-set in the component of $\princ G$ -- a ``twisted'' version of $\princ G$.}
  We've seen this family before in the guise of (preimages of) the
  ``universal \covering'' of \cref{def:universalcover}.

  There is nothing sacred about starting the identification
  $\sh_G \eqto z$ at $\sh_G$.
  Define more generally
  \begin{equation}\label{eq:pathsp}
    \pathsp{\blank}:\BG\to\GSet,
    \qquad
    \pathsp{y} \defeq (z \mapsto (y\eqto z)),
  \end{equation}
  Applying $\pathsp{\blank}$ to a path $q:y\eqto y'$
  induces an equivalence from $\pathsp y$ to $\pathsp {y'}$ that sends $p:y \eqto z$
  to $pq^{-1}:y'\eqto z$.
  As a matter of fact, \cref{lem:BGbytorsor} will identify $\BG$ with the type of
  $G$-torsors via the map $\pathsp{\blank}$,
  using the full transport structure of the identity 
  type $\pathsp y(z)\jdeq(y \eqto z)$.

  Note that the underlying set of $\princ G$ is
  \[
    \princ G(\sh_G) \jdeq
    (\sh_G \eqto \sh_G) \jdeq \USymG,
  \]
  the underlying symmetries of $G$. %
\end{example}

\begin{example}\label{def:adjointrep}
  If we vary both ends of the identifications simultaneously, we get another
  $G$-set: If $G$ is a group, then
  \[
    \Ad_G:\BG\to\Set,\qquad\Ad_G(z)\defequi(z\eqto z)
  \]
  is a $G$-set called
  the \emph{adjoint $G$-set}.\footnote{%
    Note that $\Ad_G$ also makes sense for $\infty$-groups.
    With the name ``adjoint'' we conform to usual terminology.
    The action of $\Ad_G$ works as conjugation: if $p:y \eqto z$,
    then $\Ad_G(p):(y\eqto y) \equivto (z\eqto z)$ is given by:
    \[
      \Ad_G(p)(q)\eqto pqp^{-1} \text{ in } z \eqto z.
    \]
    The picture
    \[
      \begin{tikzcd}[ampersand replacement=\&]
        y \ar[r,eqr,"p"]\ar[d,eql,"q"'] \& z \ar[d,eqr,"\Ad_G(p)(q)"] \\
        y \ar[r,eql,"p"'] \& z
      \end{tikzcd}
    \]
    is a mnemonic device illustrating that it couldn't have been different,
    and should be contrasted with the picture for
    $\princ G (p):(\sh_G\eqto y)\equivto (\sh_G\eqto z)$:
    \[
      \begin{tikzcd}[ampersand replacement=\&]
        \sh_G \ar[r,eqr,"{\refl{\sh_G}}"]\ar[d,eql,"q"']
          \& \sh_G \ar[d,eqr,"\princ G(p)(q)"] \\
        y \ar[r,eql,"p"'] \& z.
      \end{tikzcd}
    \]
  }\label{ft:adjoint-transport}
By the induction principle for the circle, the type
\[
  \sum_{z:\BG}\Ad_G(z) \jdeq \sum_{z:\BG}(z \eqto z)
\]
is equivalent to the type of (unpointed!) maps $\Sc\to\BG$, known in other
contexts as the \emph{free loop space} of $\BG$, an apt name given that it is
the type of ``all symmetries in $\BG$.''  Under that equivalence, the first
projection $\sum_{z:\BG}\Ad_G(z)\to \BG$ correspond to the evaluation function
$(\Sc\to\BG)\to\BG$ at the base point $\base$.
\end{example}
\begin{example}
  \label{ex:HomHGasGset}
  Let $G$ and $H$ be groups. Recall that the type $\Hom(H,G)$ of homomorphisms from
  $H$ to $G$ is a set (\cref{lem:hom-is-set}). We will define group actions on 
  $\Hom(H,G)$ by moving the shapes of $G$ and $H$ as
  in \cref{exa:conj-concrete}: Reusing the notation
  $\Hom(H,G)$, define for any $x:\BH$ and $y:\BG$
  \[
    \Hom(H,G)(x,y)\defequi \Hom(\mkgroup(\BH_\div,x),\mkgroup(\BG_\div,y)).
  \]
  Alternatively, by \cref{def:pointedtypes} and
  \cref{def:grouphomomorphism}, we have
  \[
    \Hom(H,G)(x,y)\jdeq 
    \Copy_{\mkgroup}\biggl(\sum_{f:\BH_\div\to \BG_\div}(y\eqto f(x))\biggr).
  \]
  The type $\Hom(H,G)$ may be considered to be a $(H\times G)$-set:
  \[
    \Hom(H,G) : (\BH\times \BG)\to\Set,
  \]
  and we shall be particularly interested in the restriction to $G$,
  giving a $G$-set for which we again reuse the notation:
  \[
    \Hom(H,G)(y)\defequi\Hom(H,G)(\sh_H,y).\qedhere
  \]
\end{example}
\begin{xca}
  \label{xca:HomZGvsAdG}
  Provide an identification between the $G$-sets
  $\Ad_G$  and $\Hom(\ZZ,G)$
  of \cref{def:adjointrep,ex:HomHGasGset}.\footnote{%
    Hint: This is similar to \cref{ex:Zinitial}:
    identify $\Hom(\ZZ,G)(y)$ with $\sum_{z:\BG}\sum_{p:z\eqto z}(y \eqto z)$
    and use~\cref{lem:contract-away}.}
\end{xca}

\begin{definition}\label{def:map-of-Gsets}
  If $G$ is a group and $X,Y$ are $G$-sets,\footnote{%
  This definition generalizes to \inftygps and $G$-types.}
  then a
  \emph{map from $X$ to $Y$} is an element of the set
  \[
    \Hom_G(X,Y) \defeq \prod_{z:\BG}(X(z)\to Y(z)).
  \]
  When $f$ is such a map, we may write $f_z$ for $f(z)$.\footnote{Such a map $f$
    is called a map of famnilies in \cref{def:fiberwise}.}  \index{map!of
    $G$-sets}\index{Gsubset@$G$-subset}
  \glossary(HomG){$\protect\Hom_G(X,Y)$}{type of maps of $G$-sets}
\end{definition}

\begin{xca}\label{xca:equivariant-map-totalization}
  Show that $\Hom_G(X,Y)$ is equivalent to the type of functions from the
  totalization of $X$ to that of $Y$ that commute to the projections, i.e.,
  $\sum_{f : \Tot(X) \to \Tot(Y)}\fst \eqto \fst \circ f$.
\end{xca}


\begin{remark}\label{rem:map-of-Gsets}
  Given $G$-sets $X,Y$ and a map $f$ from $X$ to $Y$, we have
  $f_w(g\cdot_X x) = g\cdot_Y f_z(x)$ for all $z,w:\BG$, $x:X(z)$, $g:z\eqto
  w$. In other words, the diagram on the right commutes:
\[
\begin{tikzcd}
  z\ar[d,eql,"g"'] &X(z) \ar[r,"f_z"] \ar[d,"g\cdot_X\,\blank"']
                  &Y(z) \ar[d,"g\cdot_Y\,\blank"] \\
  w               &X(w) \ar[r,"f_w"']                & Y(w)
\end{tikzcd}
\]
An important special case is when $Y$ is the trivial $G$-set $\triv_G\Prop$ that
is constantly the set $\Prop$: Given a map $P$ from $X$ to $\triv_G\Prop$, we
have $P_w(g\cdot x)$ if and only if $P_z(x)$ for all $z,w:\BG$, $x:X(z)$,
$g:z\eqto w$.  This applies to the following definition.
\end{remark}

\begin{definition}\label{def:Gsubset}
  A \emph{$G$-subset} of a $G$-set $X$ is a map from $X$ to the $G$-set 
  $\triv_G\Prop$. 
  The type of all such maps is denoted\footnote{%
  Recall \cref{def:subtype}: $\Sub(T)\jdeq(T\to\Prop)$.}
  \glossary(SubGX){$\protect\Sub_G(X)$}{set of $G$-subsets of $X$}
  \[
   \Sub_G(X)\defeq \Hom_G(X,\triv_G\Prop) 
%   \jdeq \bigl(\prod_{z:\BG}(X(z)\to\Prop)\bigr)
   \jdeq \prod_{z:\BG}\Sub(X(z)).
  \]
  Similarly to \cref{cor:subtype-same-level}, $\Sub_{G}(X)$ is a set.\footnote{%
  \label{ft:SubTotX} The type $\Sub_{G}(X)$ can be uncurried 
  (\cref{xca:Sigma-curry}) as $\Tot(X)\to\Prop$, 
  the type of subtypes of $\Tot(X)\jdeq\sum_{z:\BG}X(x)$ (\cref{def:subtype}).}
  If $P$ is a $G$-subset of $X$, then the \emph{underlying $G$-set of $P$},
  denoted by $X_P$, is defined by
  \[
  X_P(z) \defeq \sum_{x:X(z)}P(z)(x),\quad\text{for all $z:\BG$}.\qedhere
  \]
  \glossary(Gset){$X_P$}{underlying $G$-set of $P$}
\end{definition}

\begin{xca}\label{xca:SubGX-closedSubXshG}
Show that evaluation at $\sh_G$ is an equivalence from $\Sub_G(X)$ to
\[
  \sum_{Q:\Sub(X(\sh_G))}
  \prod_{x:X(\sh_G)}
  \Bigl(Q(x)\to\prod_{g:\USymG} Q(g\cdot x)\Bigr).
\]
The latter type is the type of all subsets of $X(\sh_G)$ that are closed under
the group action.
\end{xca}

The following exercise wil be used in the subsequent remark.
\begin{xca}\label{xca:ptd-conn-to-comp}
Let $(A,a)$ and $(B,b)$ be pointed types and let $A$ be connected.
Give an equivalence from $(A,a)\ptdto(B,b)$ to
$(A,a)\ptdto(B_{(b)},(b,!))$.
% $(f,p)\mapsto((x\mapsto (f(x),!_x)),(p,!)$, with $!_x$ by
% truncation induction from $\ap{f}(r_x) p$, $r_x: a\eqto x$.
\end{xca}

\begin{remark}
  \label{remark:GsetsareGsets}
  A $G$-set $X$ is often presented by focusing on the underlying set $X(\sh_G)$
  and providing it with a structure relating it to $G$ determining
  the entire function $X : \BG\to\Set$.
  More precisely, since $\BG$ is connected, 
  using \cref{xca:ptd-conn-to-comp},
  we have the following chain of easy equivalences:
  \begin{align*}
\GSet &\jdeq (\BG_\div\to\Set) \\
&\we \sum_{S:\Set} \sum_{X:(\BG_\div\to\Set)}(S\eqto X(\sh_G)) \\
&\jdeq \sum_{S:\Set} (\BG\ptdto(\Set,S)) \\
&\we \sum_{S:\Set} (\BG\ptdto(\Set_{(S)},(S,!))) \\
&\we \sum_{S:\Set} \Hom(G,\SG_S)
  \end{align*}
  Hence a $G$-set $X$ can, without loss of information, be considered as a set
  $X(\sh_G)$ and a homomorphism from $G$ to the permutation group of
  $X(\sh_G)$. Conversely, a set $S$ together with a homomorphism $f$ from $G$ to
  $\SG_S$ can be described as a $G$-set with underlying set $S$ and whose group
  action $(g \cdot \blank)$ for $g : \USymG$ is $f(g)$.\footnote{The $G$-set
    associated to the homomorphism $f$ derives from the ``delooping'' of $f$,
    and it will be formally defined in \ref{sec:delooping}.}
\end{remark}

\begin{definition}\label{def:Gaction}
  If $G$ is a group and $S$ is a set, then an \emph{action}
  of $G$ on $S$
  is a homomorphism from $G$ to the permutation group of $\SG_S$ of $S$.%
  \index{action!of a group on a set}
\end{definition}
By the construction in~\cref{remark:GsetsareGsets} we identify $G$-sets
and sets with an action of $G$ on a set.

\begin{xca}\label{xca:Ad-triv-abelian}
  Prove that a group $G$ is abelian if and only if the $G$-sets $\Ad_G$ and
  $\triv_G(\USymG)$ are identical.
\end{xca}

\begin{xca}\label{xca:Ad-princ-trivial}
  Prove that a group $G$ is the trivial group if and only if 
  the $G$-sets $\Ad_G$ and $\princ G$ are identical.
\end{xca}

\begin{definition}\label{def:finite-G-set}
Let $G$ be a group and $X:\BG\to\Set$ a $G$-set.
We say $X$ is \emph{finite} if the underlying set $X(\sh_G)$ is finite.\footnote{Since $\BG$ is connected, $X(\sh_G)$ is finite if and only if $X(z)$ is finite for any $z:\BG$. Similarly, for any $n:\mathbb N$, $X(\sh_G)$ is an $n$-element set if and only if $X(z)$ is such for any $z:\BG$.}
\index{finite $G$-set}
For any finite $G$-set $X$ we denote the number of elements
in $X(\sh_G)$ by $\Card(X)$, also called the \emph{cardinality} of $X$.
\index{cardinality!of finite $G$-set}
\end{definition}

\subsection{Transitive $G$-sets}
\label{sec:transitiveGsets}
We saw in~\cref{cha:circle} that \emph{connected \coverings}
play a special role:
In the case of the circle, classifying the group of integers $\ZZ$,
they correspond to cycles (\cref{thm:cycset-connS1cover}).

We hinted there that they are connected to subgroups, so
we now study them over a general group $G$.
When seen as $G$-sets, connected \coverings over $\BG$ are called transitive $G$-sets:
% Classically, an $\abstr(G)$-set (a notion \emph{we} have yet not defined) $\mathcal X$ is said to be \emph{transitive} if there exists some $x:\mathcal X$ such that for all $y:\mathcal X$ there exists a $g:\mathcal X$ with $x=g\cdot y$.
% In our world this translates to:
\begin{definition}\label{def:transitiveGset}
  A $G$-set $X:\BG\to\Set$ is \emph{transitive}\index{transitive $G$-set} if the proposition\footnote{This is a natural translation in the concrete world of the usual definition of transitivity in ordinary group theory: An action of an abstract group of a set $\mathcal X$ is transitive when there exists $x$ in $\mathcal X$ such that for all $y$ in $\mathcal X$ there exists an element $g$ of the group such that $x=g\cdot y$.}
  \[
    \istrans(X) \defequi
    \exists_{x:X(\sh_G)} \prod_{y:X(\sh_G)} \exists_{g:\USymG} (x=g\cdot y)
  \]
  holds.
\end{definition}
\begin{remark}
  In other words, $X$ is transitive if and only if there exists
  some $x:X(\sh_G)$ such that the map $\blank\cdot x:\USymG\to X(\sh_G)$ is
  surjective.

  Note also that by connectedness (cf.~\cref{xca:component-connected})
  it is equivalent to demand this over all $z:\BG$:
  \begin{equation}\label{eq:Gset-trans-gen}
    \prod_{z:\BG}\exists_{x:X(z)}
    \prod_{y:X(z)}\exists_{g:z\eqto z}(x=g\cdot y).
  \end{equation}

  Yet another equivalent way of expressing that $X$ is transitive is to say
  that $X(\sh_G)$ is nonempty and for any $x,y:X(\sh_G)$ there
  exists some $g:\USymG$ with $x = g\cdot y$.
  Note that the empty $G$-set is not transitive.
\end{remark}

\begin{lemma}
  \label{lem:conistrans}
  A $G$-set is transitive if and only if 
  the associated \covering is connected (see \cref{def:covering}).
\end{lemma}
\begin{proof}
  Consider a $G$-set $X:\BG\to\Set$ and its associated \covering
  $\fst:\Tot(X)\to\BG$. Recall that $\Tot(X)\defequi\sum_{w:\BG}X(w)$.  Now,
  $\Tot(X)$ is connected if and only if it is nonempty and for all $z:\BG$ and
  $x:X(z)$ and all $w:\BG$ and $y:X(w)$ there exists $g:z\eqto w$ such that
  $y=g\cdot x$. Since $\BG$ is connected, this is equivalent to asserting that
  $\Tot(X)$ is nonempty and for all $x,y:X(\sh_G)$ there exists $g:\USymG$ such
  that $y=g\cdot x$.
\end{proof}

The next lemma is an analog of~\cref{cor:ConnCycles} (for cycles),
and a generalization of~\cref{lem:evisinj-bicycle} (for bicycles).
The action in \cref{fig:not-normal}
corresponds to the bicycle back in~\cref{fig:abnormal-bicycle}
(and reproduced in~\cref{fig:not-normal-graph})
illustrates what can go wrong.
\begin{marginfigure}
  \noindent\begin{tikzpicture}[scale=.1]
    \coordinate (two)   at (0, 10);
    \coordinate (one)   at (0, 6);
    \coordinate (zero)  at (0, 2);
    \coordinate (base)  at (0,-5);

    \pgfmathsetmacro\cc{.55228475}% = 4/3*tan(pi/8)
    \pgfmathsetmacro\cy{2*\cc}%
    \pgfmathsetmacro\cx{10*\cc}%
    \pgfmathsetmacro\intx{3.5}%
    \pgfmathsetmacro\inty{1.5}%
    \pgfmathsetmacro\ay{.35165954}%

    % right 3-cycle
    \draw[casblue] (zero) .. controls ++(0,-\cy+\ay) and ++(-\cx,-\ay)
    .. (10,1) .. controls ++(\cx,+\ay) and ++(0,-\cy-\ay)
    .. (20,4)
    \foreach \y in {4,8} {
      .. controls ++(0,\cy + \ay) and ++(\cx,-\ay)
      .. (10,3 + \y) .. controls ++(-\cx,\ay) and ++(0,\cy-\ay)
      .. (0,2 + \y) .. controls ++(0,-\cy+\ay) and ++(-\cx,-\ay)
      .. (10,1 + \y) .. controls ++(\cx,\ay) and ++(0,-\cy-\ay)
      .. (20,4 + \y) }
    .. controls ++(0,+\cc) and ++(\cx,\ay)
    .. (10+\intx,12 + \inty) .. controls ++(-\cx,-\ay) and ++(\cx,\ay)
    .. (10-\intx,2 + \inty) .. controls ++(-\cx,-\ay) and ++(0,\cc)
    .. (zero);

    % left 2-cycle
    \draw[casred] (one) .. controls ++(0,-\cy+\ay) and ++(\cx,-\ay)
    .. (-10,5) .. controls ++(-\cx,+\ay) and ++(0,-\cy-\ay)
    .. (-20,8) .. controls ++(0,\cy + \ay) and ++(-\cx,-\ay)
    .. (-10,11) .. controls ++(+\cx,\ay) and ++(0,\cy-\ay)
    .. (two) .. controls ++(0,-\cy+\ay) and ++(\cx,-\ay)
    .. (-10,9) .. controls ++(-\cx,\ay) and ++(0,-\cy-\ay)
    .. (-20,12) .. controls ++(0,+\cc) and ++(-\cx,\ay)
    .. (-10-\intx,12 + \inty) .. controls ++(\cx,-\ay) and ++(-\cx,\ay)
    .. (-10+\intx,6 + \inty) .. controls ++(\cx,-\ay) and ++(0,\cc)
    .. (one);

    % left 1-cycle
    \draw[casred] (zero) .. controls ++(0,\cy) and ++(\cx,0)
    .. (-10,4) .. controls ++(-\cx,0) and ++(0,\cy)
    .. (-20,2) .. controls ++(0,-\cy) and ++(-\cx,0)
    .. (-10,0) .. controls ++(\cx,0) and ++(0,-\cy)
    .. (zero);

    % base right
    \draw (base) .. controls (0,-5+\cy) and ++(-\cx,0)
    .. (10,-3) .. controls ++(\cx,0) and ++(0,\cy)
    .. (20,-5) .. controls ++(0,-\cy) and ++(\cx,0)
    .. (10,-7) .. controls ++(-\cx,0) and ++(0,-\cy) .. (base);
    % base left
    \draw (base) .. controls (0,-5 + \cy) and (-10+\cx,-3)
    .. (-10,-3) .. controls (-10-\cx,-3) and (-20,-5 + \cy)
    .. (-20,-5) .. controls (-20,-5 - \cy) and (-10-\cx,-7)
    .. (-10,-7) .. controls (-10+\cx,-7) and (0,-5 - \cy)
    .. (base);

    % draw dots last
    \node[dot,label=above:$x$] (ntwo) at (two) {};
    \node[dot] (none)   at (one) {};
    \node[dot] (nzero)  at (zero) {};
    \node[dot] (nbase)  at (base) {};
  \end{tikzpicture}
  \caption{An $\mkgroup(\Sc\vee\Sc)$-set $X$ for which $\protect\ev_x$ is not
   surjective. At the bottom the type $\Sc\vee\Sc$ is visualized as
   two circles with a common base point. Note that the underlying
   set of $X$ with the red and the blue permutation is a bicycle
   in the sense of \cref{def:bicycle}.}
  \label{fig:not-normal}
\end{marginfigure}

\begin{marginfigure}
  \noindent\begin{tikzpicture}
    \pgfmathsetmacro{\len}{1}
    \node[vertex] (n1) at (0:\len) {};
    \node[vertex,label=above:$x$] (n2) at (120:\len) {};
    \node[vertex] (n3) at (240:\len) {};
    \begin{scope}[every to/.style={bend right=22}]
      % generator a
      \draw[genb] (n1) to (n2);
      \draw[genb] (n2) to (n3);
      \draw[genb] (n3) to (n1);
    \end{scope}
    % generator b
    \draw[gena] (n1) to[out=-30,in=30,looseness=25] (n1);
    \draw[gena,out=205,in=155] (n2) to (n3);
    \draw[gena,out=45,in=-45] (n3) to (n2);
  \end{tikzpicture}
  \caption{Alternative representation of the $\mkgroup(\Sc\vee\Sc)$-set
    $X$ from~\cref{fig:not-normal},
    using colors and arrows to represent which
    parts lies over which circle in which orientation.}
  \label{fig:not-normal-graph}
\end{marginfigure}

\begin{lemma}
  \label{lem:evisinjwhentransitive}
  Let $X,Y:\BG\to\Set$ be $G$-sets. Let $z:\BG$ and $x:X(z)$.
  If $X$ is transitive, then the evaluation map
  \[
    \ev_x:\Hom_G(X, Y)\to Y(z),\qquad \ev_x(f)\defequi f_z(x)
  \]
  is injective.\footnote{%
    Recall that for type families $X,Y:T\to\UU$, and
    $f:\prod_{z:T}(X(z)\to Y(z))$, we may write $f_z:(X(z)\to Y(z))$
    (instead of the more correct $f(z)$) for its evaluation at $z:T$.}
\end{lemma}
\begin{proof}
  We show that for any $y:Y(z)$, there is at most one
  $f:\Hom_G(X,Y)$ such that $f_z(x)=y$.
  Let $f,f':\Hom_G(X,Y)$ such that $f_z(x)=y=f'_z(x)$.
  Let $w:\BG$ and  $x':X(w)$. It suffices to show that $f_w(x')=f'_w(x')$.
  Since the latter is a proposition, we may assume (by the transitivity 
  of $X$, using \cref{lem:conistrans}) that we have a $g:z\eqto w$ such 
  that $g\cdot_X x=x'$. Using~\cref{rem:map-of-Gsets}, we have
  \[
    f_w(x')=f_w(g\cdot_X x)=g \cdot_Y f_z(x)=
    g \cdot_Y f'_z(x) = f'_w(g\cdot_X x) = f'_w(x').\qedhere
  \]
\end{proof}

In~\cref{sec:normal}, in our study of {\em normal} subgroups, we will see exactly when
we also get surjectivity of the evaluation map.


Via function extensionality,
the identity type $X \eqto Y$, for $G$-sets $X,Y$
is a subtype of the type $\Hom_G(X,Y)$.
Hence we also have that evaluation at some $x:X(z)$,
for any given $z:\BG$, is an injection
\[
  \ev_x:(X \eqto_{\GSet} Y)\to Y(z).
\]
\begin{xca}\label{xca:not-normal}
Reverse engineer the $\mkgroup(\Sc\vee\Sc)$-set $X$ in 
\cref{fig:not-normal,fig:not-normal-graph}.
Show that $X\eqto X$ is contractible.
Conclude that $\ev_x$, while injective, is not surjective.
(Hint: the induction principle for $\Sc\vee\Sc$ is a generalization
of the induction principle for the circle to two loops.)
\end{xca}
We can now generalize the definition of normal bicycle
from~\cref{def:normal-bicycle} to
transitive $G$-sets:
\begin{definition}\label{def:normal-action}
  A transitive $G$-set $X : \BG \to \Set$
  is \emph{normal} if the evaluation map
  \[
    \ev_x: \bigl(X \eqto_{\GSet} X) \to X(z),
    \qquad \ev_{x}(e)\defequi e_z(x)
  \]
  is an equivalence for all $z:\BG$ and $x:X(z)$.
\end{definition}
\begin{xca}\label{xca:normal-action-equiv}
  Show that if the evaluation map is an equivalence for some pair of a $z:\BG$ and a $x:X(z)$,
  then it is an equivalence for all such pairs.
  (This generalizes~\cref{xca:normal-bicycle-equiv}.)
\end{xca}

\subsection{Actions in a type}
\label{sec:actions}
Oftentimes it is interesting not to have an action on a set, 
but on an element in any given type (not necessarily the type of sets).  
For instance, a group can act on another, giving rise to the notion of 
the semidirect product in~\cref{sec:Semidirect-products}.  
We will return to these more general types of actions many times.

\begin{definition}\label{def:action}
  If $G$ is any group\footnote{%
  Even an $\infty$-group in the sense of \cref{sec:inftygps}.}
  and $A$ is any type,
  then we define an \emph{action of $G$ in $A$} as a function
  \[
    X : \BG \to A.
  \]
  The particular ``object of type $A$ being acted on'' is $X(\sh_G):A$,
  \index{action!of a group in a type}

  Fixing $a:A$ as the underlying object, 
  we define an \emph{action of $G$ on $a$}
  to be a homomorphism from $G$ to $\Aut_A(a)$.%
  \index{action!of a group on an element}
\end{definition}
This generalizes our earlier definition of $G$-sets $X : \BG \to \Set$
from~\cref{def:Gset},
and harmonizes with~\cref{remark:GsetsareGsets}, relating $G$-sets and
actions of $G$ on a set.
Indeed, we identify
an action of $G$ in $A$ with a pair of an underlying object
$a:A$ and an action of $G$ on $a$:
\[
  (\BG \to A) \equivto \sum_{a:A}\Hom(G,\Aut_A(a))
\]
This equivalence,
hinted at in~\cref{xca:BGtotype},
maps an action $X:\BG\to A$
to the pair consisting of $a \defeq X(\sh_G)$
and the homomorphism classified by the pointed map
from $\BG$ to the pointed component $\conncomp A a$ given by $X$.

\begin{definition}\label{std-action}
  The \emph{standard action} of $G$ on its designated shape $\sh_G$ is obtained by
  taking $A \defeq \BG$ and $X \defeq \id_{\BG}$.
\end{definition}

\begin{example}\label{ex:S2-acts-on-C3}
  The symmetric group $\SG_2$ acts on the cyclic group $\CG_3$ as follows.
  Given a $2$-element set $S$ consider the
  type $\sum_{X:\Set}(S \to (X\to X))$ of pairs $(X,f)$ of a set $X$
  and a ``pair'' of functions $f_s:X\to X$ (one for each $s:S$).
  In this type we have the element $(\bn 1 \amalg S,f)$,
  consisting of the $3$-element set $\bn 1 \amalg S$ and the function
  $f:S \to ((\bn 1 \amalg S)\to (\bn 1 \amalg S))$ defined by
  \begin{align*}
    f_s(\inl 0)     &\defeq \inr s,\\
    f_s(\inr s)        &\defeq \inr{\swap(s)},\\
    f_s(\inr{\swap(s)}) &\defeq \inl 0.
  \end{align*}
  Then $G(S) \defeq \Aut_{\sum_{X:\Set}S\to X\to X}(\bn1\amalg S,f)$ defines an action
  $\BSG_2 \to \Group$.\footnote{%
    If $S$ is $\set{s,s'}$, then we can picture the
    designated shape as follows,
    where the blue and red arrows denote $f_s$ and $f_{s'}$,
    respectively:\par
    \begin{tikzpicture}
    \draw (-.1,1) ellipse (.35 and .35);
    \node (X) at (0,1.5) {$\bn 1$};
    \draw (1,1) ellipse (.4 and 1);
    \node (Y) at (.9,2.2) {$S$};
    \node[dot,label=left:$0$] (x) at (0,1) {};
    \node[dot,label=above:$s$]  (s1) at (1,1.5) {};
    \node[dot,label=below:$s'$] (s2) at (1,.5) {};
    \draw[dashed] (0.6,1.1) ellipse (1.2 and 1.6);
    \begin{scope}[every to/.style={bend left=30}]
      % generator a
      \draw[gena] (x) to (s1);
      \draw[gena] (s1) to (s2);
      \draw[gena] (s2) to (x);
    \end{scope}
    % generator b
    \draw[genb] (x) to (s2);
    \draw[genb] (s2) to (s1);
    \draw[genb] (s1) to (x);
    \node (XY) at (-0.75,2.35) {$\bn1\amalg S$};
  \end{tikzpicture}}
  Furthermore, we identify $G(\bool)$ with $\BCG_3$ by mapping
  a shape $(X,f)$ in $\BG(\bool)$ to the $3$-cycle $(X,f_\yes)$
  and identifying the $3$-cycle $(\bn1\amalg\bool,f_\yes)$, for the $f$ defined above,
  with the standard $3$-cycle $(\bn3,\zs)$, correlating $\inl 0$ with $0:\bn 3$.
\end{example}
\begin{xca}\label{xca:AutC3}
  Show that action of $\SG_2$ on $\CG_3$ from~\cref{ex:S2-acts-on-C3}
  gives an identification $\SG_2 \eqto \Aut(\CG_3)$.
\end{xca}

\begin{example}
  By composing constructions we can build new actions
  starting from simple building blocks.
  For example, the standard action of symmetric group $\SG_n$
  is to permute the elements of the standard $n$-element set $\bn n$.
  Composing with the projection $\BSG_n \to \Set$,
  we get the corresponding standard $\SG_n$-set.\footnote{%
    Check that this action is transitive for $n>0$.}
  Composing further with the operation $\blank \to \bool : \Set \to \Set$,
  mapping any set $S$ to the set $S\to\bool$,
  we get the action of $\SG_n$ on the set of decidable subsets of $\bn n$.
\end{example}

\section{Subgroups}
\label{sec:subgroups}
In our discussion of the group $\ZZ\defequi\Aut_{\Sc}(\base)$ of integers in
\cref{cha:circle} we discovered that some of the symmetries of $\base$ were
picked out by the degree $m$ function $\dg{m}: \Sc\to\Sc$ (for some particular
natural number $m>0$, see \cref{def:dgm-map}).  At the level of the set
$\USym \ZZ \defeq (\base\eqto{}\base)$, the symmetries picked out are all the
iterates (positive or negative or even zero-fold) of $\Sloop^m$.  The important
thing is that we can compose or invert any of the iterates of $\Sloop^m$ and get
new symmetries of the same sort (because of distributivity
$mn_1+mn_2=m(n_1+n_2)$).%  So, while we do not get all symmetries of $\base$
% (unless $m=1$), we get what we'd like to call a subgroup of the group of
% integers.

The case of $m=0$ is special. The iterates of $\Sloop^0$, \ie of
$\refl{\base}$, can also be composed and inverted, never to give
something else than $\Sloop^0$ itself. % This is what we'd like to call
% the trivial subgroup of the group of integers. 
We can pick out the single symmetry $\Sloop^0$
by the constant map $\cst{\base} : \bn 1 \to \Sc$.

In every case, while we do not get all symmetries of $\base$ (unless $m=1$), we
get a set of these that behave, under the group laws of $\ZZ$, as the set of
symmetries of some shape. This phenomenon is captured by the notion of {\em
  subgroup}.

Formally speaking, we can sum up the situation for $\ZZ$ as follows. Both
$\dg{m}$ and $\cst{\base}$ can trivially be pointed to make them into
classifying maps of homomorphisms that are injections on the respective sets of
symmetries. Using \cref{cor:dgm-conncov}, each $\dg{m}$ is a pointed connected
\covering over the circle, and $\cst{\base}$ is even the universal \covering by
\cref{lem:univ-cover-of-groupoid}. Finally, \cref{lem:conistrans} lets us view
these as transitive $\ZZ$-sets with a chosen element in their underlying
sets. We adopt this last point of view to genereralize the notion to any group
$G$.

\subsection{Subgroups through $G$-sets}

% The idea of this approach is to take the total type of a transitive
% $G$-set $X$ and to choose a point $x:X(\sh_G)$ in the underlying set of $X$.
% Then the symmetries of $(\sh_G,x)$ are precisely the symmetries of $\sh_G$
% that keep the chosen point $x$ in place.

\begin{definition}\label{def:set-of-subgroups}
  For any group $G$, define the type of \emph{subgroups of $G$} as%
  \index{type!of subgroups of a group}%
  \glossary(SubG){$\protect\Sub(G)$}{type of subgroups of $G$}
  \[
    \Sub(G)\defequi\sum_{X:\BG\to\Set}{\,}X(\sh_G)
    \times\istrans(X). \qedhere
  \]
\end{definition}

\begin{xca}\label{xca:group-Xx!}
Show that, for any transitive $G$-set $X$, the type $\sum_{z:BG}X(z)$ above is a connected groupoid.
Hint: use \cref{lem:conistrans}.
\end{xca}

The reader could be surprised that the definition of a subgroup of a group $G$
is \textbf{not} that of a group with some relationship to $G$. Let us thus unfold a
bit the definition above. Given a subgroup\footnote{%
    To lighten the notation, we leave out the proof that $X$ is transitive.
    (Otherwise, we would write $(X,\pt,!):\Sub(G)$.)
    In~\cref{rem:notationsubgroup} below we'll set out further notational
    conveniences regarding subgroups.} $(X,\pt):\Sub(G)$ in the sense above,
we can consider the total type $\Tot(X) \defeq \sum_{z:BG}X(z)$ of the
transitive $G$-set $X$. Together with the first projection, it makes a connected
\covering over $\BG$ and its fiber (at $\sh_G$) contains a particular point,
namely $(\sh_G,\pt)$. \cref{xca:group-Xx!} above tells us that this choice of
point makes a group. The first projection then becomes (the classifying map of)
a homomorphism from that group to $G$ that embeds\footnote{More will be said on
  that ``embedding'' in~\cref{sec:subgroups-mono}.} the symmetries of
$(\sh_G,\pt)$ in $\Tot(X)$ into these of $\sh_G$ in $\BG$. This explain the
following definition:

\begin{definition}\label{def:underlying-group-of-subgroup}
  Let $G$ be a group. The \emph{underlying group} %
  \index{type!of subgroups of a group}%
  of a subgroup
  $(X,\pt) : \Sub(G)$ is
  \[
    \mkgroup \left(\Tot(X),(\sh_G,\pt)\right).\qedhere
  \]
\end{definition}

\begin{example}\label{ex:Rm0-subgroup} 
Recall from \cref{def:RmtoS1} the $\ZZ$-set
$R_m : \Sc\to\Set$ defined by $R_m(\base) \defeq \bn m$ and
$R_m(\Sloop) \defis \etop\zs$. Here $m>0$ so that we can point
$R_m$ by $0: R_m(\base)$.\footnote{Any element of $\bn m$ would do.}
Transitivity of $R_m$ is obvious.
Which symmetries $p: \base\eqto\base$ are picked out by $R_m$,
that keep the point $0: R_m(\base)$ in place? Those that 
satisfy $R_m(p)(0)=0$, \ie $p=\Sloop^{mk}$ for some integer $k$.
Given $\alpha_m$ in \cref{con:psi-alpha-m}, it should not come as a
surprise that these are precisely the symmetries picked out by $\dg{m}$.

The case of $m=0$ connects to another old friend: the $\ZZ$-set
$R : \Sc\to\Set$ defined by $R(\base) \defeq \zet$ and
$R(\Sloop) \defis \etop\zs$, see \cref{def:RtoS1}.
Again we point by $0: R(\base)$ and transitivity of $R$ is obvious.
The only symmetry that keeps $0$ in place is $\refl{\base}$,
since $R(\Sloop^k)(0) = \zs^k(0) = k = 0$ if and only if $k=0$.
Again, no surprise in view of the results in \cref{sec:symcirc}
identifying $R$ as the universal \covering over $\Sc$.
\end{example}

The following result is analogous to the fact that $\Sub(T)$ is
a set for any type $T$, see \cref{def:subtype}. It captures
that the essence of picking out symmetries (or picking out elements
of a type), is a predicate, like $R_m(p)(0)=0$ in \cref{ex:Rm0-subgroup}.

\begin{lemma}
  \label{lem:SubGisset}%
  For any group $G$, the type $\Sub(G)$ of subgroups of $G$ is a set.
\end{lemma}
\begin{proof}
  Let $G$ be a group, and let $(X,\pt)$ and $(X',\pt')$ be subgroups of $G$. Since
  $X$ is transitive we can use the remark following
  \cref{lem:evisinjwhentransitive} that $\ev_{\pt} : (X \eqto X') \to X'(\sh_G)$ is
  injective. In particular, the preimage of $\pt'$ is thus a proposition. Now, we
  have a chain of equivalences:
  \begin{eqnarray*}  
    \left((X,\pt)\eqto(X',\pt')\right)
    & \equivto & \displaystyle\sum_{p:X \eqto X'} \pt' = \trp p {\pt} \\
    & \equivto & \displaystyle\sum_{p:X \eqto X'} \pt' = \ev_{\pt}(p) \\
    & \jdeq & {\ev_{\pt}}^{-1} (\pt'). \qedhere
  \end{eqnarray*}
\end{proof}

\begin{example}
  \label{exa:fix1subSGn}%
  Consider the symmetric group $\SG_n$ from~\cref{ex:groups}\ref{ex:permgroup},
  for some $n>0$. The $\SG_n$-set $X : \BSG_n \to\Set$ given by $X(A,!)\defeq A$
  for any $A:\Set$ together with $! : \Trunc{\bn n  \eqto A}$ is obviously transitive
  (through univalence).  For any $k:\bn n$, we can point $X$ by
  $k: X(\sh_{\SG_n}) \jdeq \bn n$.\footnote{The choice of the point does matter
    for the symmetries that are picked out.}  Thus we have $(X,k):\Sub(\SG_n)$.
  The symmetries that are picked out are those $\pi : \bn n \eqto \bn n$ that
  satisfy $(\pi \cdot_X k) = k$. % \footnote{%
    % This uses the alternative notation for the group action of $X$ introduced in
    % \cref{def:Gset}.}
  In other words, $\pi$ keeps $k$ in place and can be any
  permutation of the other elements of $\bn n$.  From the
  next~\cref{xca:n-is-ptd-n+1} we get that the underlying group of each $(X,k)$
  is isomorphic to $\SG_{n-1}$.
\end{example}

\begin{xca}\label{xca:n-is-ptd-n+1}
  Give an equivalence from the type of
  $n$-element sets to the type of pointed $(n{+}1)$-element sets.
  Hint: use~\cref{xca:finsets-decidable}.
\end{xca}

\begin{xca} \label{xca:A-is-A-1+1}
  For any set $A$ with decidable equality, give
  an equivalence from $A$ to $\sum_{B:\UU}(A\eqto(B+\bn1))$.
\end{xca}

\begin{example}
  \label{exa:C3subC6}%
Recall from \cref{ex:cyclicgroups} the definition 
$\CG_6\defeq\Aut_\Cyc(\bn6,\zs)$ of the cyclic group of order $6$.
This group can be visualized as the rotational symmetries of a regular
hexagon,  \ie the rotations by $2\pi\cdot m /6$, where $m=0,1,2,3,4,5$.
The symmetries of the regular triangle (rotations by $2\pi\cdot m/3$, 
where $m=0,1,2$) can also be viewed as symmetries of the hexagon,
see \cref{fig:C3inC6}.
Thus there is a subgroup of $\CG_6$ which, as a group, is isomorphic to 
$\CG_3$, and which we now construct.

\begin{marginfigure}
\begin{tikzpicture}[scale=.9]
  \newdimen\R
\R=2cm
  \draw (0:\R) \foreach \x in {60,120,...,360} { -- (\x:\R) } ;
  \draw (0:\R) \foreach \x in {120,240,360} { -- (\x:\R) } ;
\end{tikzpicture}
\caption{\label{fig:C3inC6}Geometrical shapes illustrating 
$\protect\CG_3$ as subgroup of $\protect\CG_6$.}
\end{marginfigure}

In order to obtain $\CG_3$ as (the underlying group of) a subgroup of $\CG_6$ we
can define $F : \CG_6 \to \Set$ by $F(X,t) \defeq X/2$ for all $(X,t):\BCG_6$,
where $X/2$ (defined in \cref{sec:mthroot}) is the quotient of $X$ modulo
identifying elements that are an even power of $t$ away from each other.
Clearly, $F$ is a transitive $G$-set.  On symmetries, $F$ maps
$\pi : (\bn6,\zs) \eqto (\bn6,\zs)$ to
$([k] \mapsto [\pi(k)]) : \bn6 /2 \eqto \bn6 /2$.\footnote{%
  The function $[k] \mapsto [\pi(k)]$ is well-defined since permutations that
  commute with $\zs$ preserve distances.}  Thus defines a subgroup
$(F,[0]) : \Sub(\CG_6)$. The symmetries it picks out are those $\pi$ satisfying
$F(\pi)([0])=[0]$ and are the even powers of $\zs$.\footnote{%
  In view of \cref{cor:id-m-cycle}, these symmetries can be visualized by the
  vertices of the regular triangle, see \cref{fig:C3inC6}.  The same is true for
  the symmetries picked out by $F(\pi)([1])=[1]$.  Both $F(\pi)([0])=[1]$ and
  $F(\pi)([1])=[0]$ give the other inscribed regular triangle.} The underlying
group of $(F,[0])$ is
\begin{displaymath}
  \mkgroup\biggl(\sum_{(X,t):\Cyc_6} X/2, ((\bn6,\zs),[0])\biggr).
\end{displaymath}

Using $\rho_2: \BCG_3 \ptdto \BCG_6$ from \cref{lem:deg-m-on-Cyc}, the
equivalence between $X/2$ and $\inv\rho_2 (X,t)$ from \cref{thm:fiber-cdg}, and
the equivalence from \cref{lem:sum-of-fibers}, we get an equivalence between the
underlying group of $(F,[0])$ and $\CG_3$.
\end{example}

\begin{definition}\label{def:decidable-subgroup}
  A subgroup $(X,\pt)$ of a group $G$ is called
  \emph{decidable}\index{decidable!subgroup} when $X(\sh_G)$ has decidable
  equality.
\end{definition}

The subgroup of $\CG_6$ explained above is a decidable subgroup. There are other
decidable subgroups of $\CG_6$, and they are accounted for simply by the various
factorizations of the number $6$.

\subsection{Subgroups as monomorphisms}\label{sec:subgroups-mono}

%Three old para's that could still be useful as intro to this 
%A monomorphism into $G$ is given by a pointed connected groupoid  $\BH=(\BH_\div,\pt_H)$, a function $F:\BH_\div\to\BG_\div$ whose fibers are sets (a \covering) and an identification $p_f:\sh_G\eqto{}F(\sh_H)$.  There is really no need to specify that $\BH_\div$ is a groupoid: if $F:T\to \BG$ is a \covering, then $T$ is automatically a groupoid.

%On the other hand,  the type of \coverings over $\BG$ is equivalent to the type of $G$-sets: if $X:\BG\to\Set$ is a $G$-set, then the \covering is given by the first projection $\tilde X\to \BG$ where $\tilde X\defequi\sum_{y:\BG}X(y)$ and the inverse is obtained by considering the fibers of a \covering.  Furthermore, we saw in \cref{lem:conistrans} that $\tilde X$ being connected is equivalent to the condition $\istrans(X)$ of \cref{def:transitiveGset} claiming that the $G$-set $X$ is transitive.

%Hence, the type (set, really) $\typemono(G)$ of monomorphisms into $G$ is equivalent to the type of pointed connected \coverings over $\BG$, which again is equivalent to the type $\Sub(G)$ of transitive $G$-sets $X:\BG\to\Set$ together with a point in $X(\sh_G)$.

For many purposes it is useful to have an alternative point-of-view on
subgroups.  We thus give here a second, equivalent definition of a
subgroup. Going back to the motivating examples $\dg{m}$ and $\cst{\base}$ from
the introduction of this chapter, recall that both $\USym\dg{m}$ and
$\USym\cst{\base}$ are injective: they ``pick out'' (in a nice controlled way)
some symmetries of $\base$ in $\Sc$.  Also recall
from~\cref{cor:fib-vs-path}\ref{set-fib-vs-path-point} that $\USymf$ is
injective iff $\Bf$ is a \covering, for any homomorphism $f$.

\begin{definition}
  \label{def:typeofmono}
  Let $G$ and $H$ be groups. We say that a homomorphism $i: \Hom(H,G)$ is
  a \emph{monomorphism},\index{monomorphism} denoted $\ismono(i)$,%
  \glossary(isMono){$\protect\ismono(i)$}{proposition stating that
    $i$ is a monomorphism of groups}
  if $\USymi:\USymH\to \USymG$ is injective  
  (\ie all preimages of $\USymi$ are propositions).
  
  The \emph{type of monomorphisms into $G$}\footnote{%
    The similarity of this type with the type of subtypes
    $\Inj(T) \defeq \sum_{S:\UU}\sum_{f:S\to T}\isinj(f)$ in
    \cref{def:injtype} is not coincidental, and the remarks made 
    there in \cref{ft:incl-vs-inj} apply here as well.
    
    In particular, the identity type of $\typemono(G)$ identifies precisely the
    triples that define the same subset of $\USym G$, namely when their
    homomorphisms differ by precomposition by an identification of their
    underlying groups.
  }
  \index{type! of monomorphisms into a group}
  \glossary(MonoG){$\protect{\typemono(G)}$}{type of monomorphisms%
  into the group $G$} 
  is
  \[
  \typemono(G)\defequi\sum_{H:\typegroup}\sum_{i:\Hom(H,G)}\ismono(i).
  \]
  
  We call $H$ the \emph{underlying group} of any monomorphism\footnote{As for
    subgroups, we abuse notation and leave out the propositional component, that
    is the element $!:\ismono(i)$.}  $(H,i)$ into $G$.
  
  A monomorphism $(H,i)$ into $G$ is:
      \begin{enumerate}
      \item \emph{trivial}\index{trivial monomorphism} 
        if $H$ is the trivial group;\footnote{This amounts to $Bi$
      being the universal \covering over $BG$, see \cref{def:univ-cover}}
%.contractible (or, equivalently, if $\USymH$ is contractible),
      \item \emph{proper}\index{proper monomorphism} if $i$ is 
      not an isomorphism.\qedhere
      \end{enumerate}
\end{definition}
    
\begin{example}
  \label{ex:SGninSGn+1}
   \marginnote{
     That $i:\SG_2\to\SG_3$ is a monomorphism can visualized as follows:
     if $\SG_3$ represent all symmetries of an equilateral triangle in the
     plane (with vertices $1$, $2$, $3$), then $i$ is represented by the
     inclusion of the symmetries leaving $3$ fixed; \ie reflection through
     the line marked with dots in the picture.
\[ \begin{tikzpicture}[scale=.9]
  \newdimen\R
\R=2cm
  \newdimen\Rl
\Rl=2.3cm
  \draw (90:\R) \foreach \x in {210,330,450} { -- (\x:\R) } ;
  \node at (90:\Rl) {3};
  \node at (210:\Rl) {1};
  \node at (330:\Rl) {2};
  \draw[dotted] (90:\R) -- (270:\R * 0.5) ;
\end{tikzpicture}
\]}
We will present the subgroups from \cref{exa:fix1subSGn} with
monomorphisms. For each $n:\NN$, consider the  homomorphism 
$i_n:\Sigma_n\to\SG_{n+1}$ of permutation groups
with $\Bi_n$ sending $A:\BSG_n\defequi \FinSet_n$ to $A\coprod\true:\BSG_{n+1}$.
As pointing path we take the reflexivity path.
This is a monomorphism since $\USymi_n:\USym\SG_n\to\USym\SG_{n+1}$ 
is an injection, extending any permutation $\pi$ of $\bn{n}$ to 
a permutation of $\bn{n}+\bn{1}$ by adding the last element as a
fixed point.

In the picture in the margin we have taken $n=2$. The elements of $\bn 2$ are
named $1$ and $2$; those of $\bn 3$ are named $1$, $2$ and $3$. The symmetry
through the dotted line is the non-identity symmetry in $\USym\SG_3$ picked out
by $\USymi_2$.

Two other subgroups can be obtained by varying the pointing path of
$i_3$. These pointing paths are induced by the permutations of $\bn{3}$. In
\cref{xca:SG2subSG3} you are asked to elaborate each case.

There are other subgroups of $\SG_3$, see \cref{xca:C3subSG3} for example.
\end{example}

\begin{xca}\label{xca:SG2subSG3}
Calculate $\im(\USymi_3)$ for each pointing path $\pi:\bn{3}\eqto\bn{3}$.
\end{xca}

\begin{xca}\label{xca:C3subSG3}
Define monomorphisms $j,j':\CG_3\to\SG_3$ such that $\USymj\neq \USymj'$
while $(\CG_3,j)$ and $(\CG_3,j')$ can be identified in $\typemono(\SG_3)$.
\end{xca}

\begin{example}
  \label{ex:prodinclismono}
  If $G$ and $H$ are groups, then $i_G : G \to G\times H$, classified by
  $\Bi_G : \BG \ptdto \BG\times\BH$ with $\Bi_G(z) \defeq (z,\sh_H)$, pointed by
  reflexivity, is a monomorphism: $\USymi_G$ maps $g:\USymG$ to
  $(g,\refl{\sh_H})$ and is obviously injective.%
  \footnote{Alternatively, the preimage of $(z,t)$ by $\Bi_G$ is
    $\sum_{z':\BG} (z \eqto z') \times (t \eqto \sh_H)$ which contracts away to
    the set $t \eqto \sh_H$. Hence $\Bi_G$ is a set bundle.} %
  We call $i_G$ the \emph{first inclusion} and we have a similar \emph{second
    inclusion} $i_H : H \to G\times H$.
\end{example}

\begin{lemma}\label{lem:SubG=MonoG}
  Let $G$ be a group.  The map $F: \Sub(G)\to \typemono(G)$ sending the subgroup
  $(X,\pt) : \Sub(G)$ to the element
  \begin{displaymath}
    \bigl(\mkgroup\bigl(\Tot(X),(\sh_G,\pt)\bigr),\mkgroup(\fst)\bigr) : \typemono(G),
  \end{displaymath}
  where $\fst: \bigl(\Tot(X),(\sh_G,\pt)\bigr)\ptdto\BG$ is pointed by
  reflexivity, is an equivalence.% \footnote{%
    % Recall that we may omit ``!''s: propositional data never dies,
    % it just fades away!}
  % \[
  %   F: \Sub(G)\to \typemono(G) : (X,x) \mapsto
  %   \biggl(\mkgroup \Bigl(\sum_{z:\BG}X(z),(\sh_G,\pt)\Bigr),\mkgroup\fst\biggr).\qedhere
  % \]
\end{lemma}

\begin{proof}
  Given a monomorphism $i:\Hom(H,G)$, write $(\Bi_\div,\Bi_\pt)$ for the two
  components of $\Bi :\BH\ptdto\BG$. Then, since $i$ is a monomorphism,
  $\Bi_\div$ is a \covering whose domain $\BH_\div$ is connected, and thus the
  preimage function $\Bi_\div^{-1} : \BG \to \Set$ is a transitive
  $G$-set. Moreover, the pair $(\sh_H,\Bi_\pt)$ is an element of
  $\Bi_\div^{-1}(\sh_G) \defequi \sum_{x:\BH}(\sh_G\eqto{}\Bi_\div(x))$. Hence,
  we can define a map $E$ as follows:%
  \glossary(E){$E$}{equivalence from $\protect\typemono(G)$ to
    $\protect\Sub(G)$}
  \[
    E:\typemono(G)\to\Sub(G),\qquad
    (H,i)\mapsto E(H,i)\defequi (\Bi_\div^{-1},(\sh_H,\Bi_\pt)),
  \]
  We claim that $E$ is an inverse equivalence for $F$. The details are left to
  the reader: see \cref{xca:SubG=MonoG} below.
\end{proof}

 \begin{example}\label{exa:EforSG3}
  In this example we explain how the equivalence between
  $\typemono(G)$ and $\Sub(G)$ works in the special case
  $G\jdeq\SG_3$ and with two versions of the same subgroup.
  
  Recall $(\SG_2,i_2):\typemono(\SG_3)$ with
  $i_2: \SG_2\ptdto\SG_3 : B\mapsto (B\coprod\true)$ from \cref{ex:SGninSGn+1}.
  The preimage function $\inv\Bi_2$ maps any $A:\BSG_3$
  to $\sum_{B:\BSG_2}(A\eqto{}(B\coprod\true))$.
  In particular we have $(\bn2,\refl{\bn3}):\inv\Bi_2(\bn3)$
  (recall that $i_2$ is pointed by reflexivity).
   
  We have $E(\SG_2,i_2)\jdeq(\inv\Bi_2,(\bn2,\refl{\bn3}))$. Applying $F$ to
  the result, we get
  $\bigl(\mkgroup\bigl(\sum_{A:\BSG_3}\inv\Bi_2(A)\bigr),\mkgroup(\fst)\bigr)$. Using
  \cref{lem:sum-of-fibers} one sees that, indeed, the latter monomorphism can be
  identified with $(\SG_2,i_3)$.
  
  Why do we say that $(X,3):\Sub(\SG_3)$ from \cref{exa:fix1subSGn} defines the
  ``same subgroup'' as $(\SG_2,i_2):\typemono(\SG_3)$ from \cref{ex:SGninSGn+1}?
  The reason is that they pick out the same symmetries in $\SG_3$, as argued in
  these examples.  Moreover, $(X,3)$ and $E(\SG_2,i_2)$ can be identified in
  $\Sub(\SG_3)$.\footnote{Equivalently, $F(X,3)$ and $(\SG_2,i_2)$ can be
    identified in $\typemono(\SG_3)$.}  Indeed, note that $X(A,!) \jdeq A$ and
  $\inv\Bi_2(A,!) \jdeq \sum_{B:\BSG_2}(A\eqto{}(B\coprod\true))$: Apply
  \cref{xca:A-is-A-1+1} and verify that the points correspond.
  \cref{lem:E-preserves-symms} below offers a general result of this kind.
  \end{example}

\begin{xca}\label{xca:SubG=MonoG}
Complete the details of the proof of \cref{lem:SubG=MonoG} above using 
\cref{cor:fib-vs-path}\ref{set-fib-vs-path-point},
\cref{lem:sum-of-fibers},
\cref{lem:conistrans}.
\end{xca}

Since $\Sub(G)$ is a set by \cref{lem:SubGisset}, 
\cref{lem:SubG=MonoG} allows us to conclude:

\begin{corollary}\label{lem:setofsubgroups}
Let $G$ be a group. Then $\typemono(G)$ is a set.
\end{corollary}

The following lemma states that the equivalences in \cref{lem:SubG=MonoG} preserve the subsets of symmetries that are picked out.
\begin{lemma}\label{lem:E-preserves-symms}
  Let $G$ be a group and $g:\USymG$ a symmetry.  Recall the equivalence $F$ from
  \cref{lem:SubG=MonoG}.  For all $(X,\pt) : \Sub(G)$ and $(H,i) : \typemono(G)$
  such that $(H,i)=F(X,\pt)$, we have $g \cdot_X \pt = \pt$ in $X_{\sh_G}$ if and only
  if $g$ is in the image\footnote{I.e., there exists $h:\USymH$ such that
    $g = \USymi(h)$ in $\USymG$.} of $\USymi$.
\end{lemma}
\begin{proof}
Let $(X,\pt):\Sub(G)$. It suffices to prove the lemma for 
$(H,i)\jdeq F(X,\pt) : \typemono(G)$.
This means $\BH\jdeq(\Tot(X),(\sh_G,\pt))$ and $\Bi\jdeq\fst$.
We have to prove: $g \cdot_X \pt = \pt$ iff there exists an
$h:(\sh_G,\pt)\eqto(\sh_G,\pt)$ such that $g = \USymi(h)\jdeq\loops\fst(h)$.

If $g \cdot_X \pt = \pt$, then we can simply take $h$ to the symmetry defined by the
pair $(g,\refl \pt)$.

For the converse, assume there exists an $h:(\sh_G,\pt)\eqto(\sh_G,\pt)$ such that
$g = \USymi(h)$.  Then $h$ is defined by a pair of the form $(g,p)$ for some
$p: \pathover {\pt} {X} g {\pt}$.\footnote{This path $p$ is in fact equal to
  $\refl{\pt}$ since $X_{\sh_G}$ is a set.} But $\trp [X] p {\pt}$ is precisely
$X(g)(\pt)$ (\ie $g \cdot_X {\pt}$).
\end{proof}

\marginnote{%
  Which of the equivalent sets $\typemono(G)$ and $\Sub(G)$ is allowed to be called ``the set of subgroups of $G$'' is, of course, a choice.  It could easily have been the other way around and we informally refer to elements in either sets as ``subgroups'' and use the given equivalence $E$ as needed.
}%
\marginnote{%
  An argument for our choice can be
 as follows.  In set-based mathematics one has two options for defining "subgroup": either as a certain subset (uniquely given by its characteristic function to $\Prop$) or as an equivalence class of injections (taking care of size issues since the class of monomorphisms will not form a small set).  The former is the usual choice and is the one we model here with $\Sub(G)$, whereas the other corresponds to $\typemono(G)$. \wip{I strongly disagree with this note: $\typemono(G)$ is akin to the collection of injections, not the set of equivalence classes of those; and our $\Sub(G)$ is certainly not akin to the subset definition since it is equivalent to $\typemono(G)$ (and thus have the same ``size'').}
  % that the identity type in $\Sub(G)$ seems more transparent than the one in $\typemono(G)$  (``more things are equal'' in $\typemono(G)$?), just as  $A\to\Prop$ gives more the intuition of picking out a subset by means of a characteristic function than what you get when considering the equivalent type of injections into $A$.
}


Through the equivalence $E$ we can translate the concepts in
\cref{def:typeofmono} to subgroups in $\Sub(G)$.  First, observe that the
underlying groups of an element in $\typemono(G)$ and of its image under $E$ in
$\Sub(G)$ can be identified.

\begin{definition}\label{def:triv-proper-Mono}
We say that a subgroup $(X,\pt):\Sub(G)$ is:
\begin{enumerate}
\item \emph{trivial}\index{trivial subgroup} if the underlying group
  $\mkgroup(\Tot(X),(\sh_G,\pt))$ is trivial, \ie when $\Tot(X)$ is contractible;
\item \emph{proper}\index{proper subgroup} if $X(\sh_G)$ is not
  contractible.\qedhere
\end{enumerate}
\end{definition}
      
\begin{remark}\label{rem:notationsubgroup}
  A note on classical notation is in order.  If $(X,\pt)$ is a subgroup
  corresponding to a monomorphism $(H,i)$ into a group $G$, tradition would
  permit us to relax the burden of notation and we could write ``a subgroup
  $i:H\subseteq\wip{\leq} G$'', or, if we didn't need the name of $i:\Hom(H,G)$, simply
  ``a subgroup $H\subseteq\wip{\leq} G$'' or ``a subgroup $H$ of $G$''.
    \end{remark}
    
\begin{example}
  \label{ex:prodinclisGset}
  We saw in \cref{ex:prodinclismono} that the first 
  inclusion $i_1:G\to G\times G'$ is a monomorphism.
  The corresponding $G\times G'$-set is the composite of the second projection
  $\mathrm{proj}_2:\BG_\div\times\BG'_\div\to \BG'_\div$ followed by the
  principal $G'$-torsor $\princ {G'}:\BG'\to\Set: z'\mapsto (\sh_{G'}\eqto z')$ 
  of \cref{def:principaltorsor}.

  Indeed, let us compute the image through $F$ of the latter. First,
  \begin{eqnarray*}
    \Tot(\princ {G'}\circ \mathrm{proj}_2)
    & \jdeq & \sum_{(z,z') : \BG_\div\times\BG'_\div}(\sh_{G'}\eqto z') \\
    & \equivto & \BG_\div \times \sum_{z' : \BG'_\div}(\sh_{G'}\eqto z') \\
    & \equivto & \BG_\div
  \end{eqnarray*}
  Notice that this equivalence $\phi$ can easily be pointed by $\refl{\sh_G}$
  and that, as pointed maps, we have an identification of type
  $\fst \eqto i_1 \circ \phi$. In other words, $\phi$ gives rise to an equality
  of pointed set bundles over $\BG\times\BG'$. This, in turn, gives rise to the
  equality $(\BG, i_1) = F(\princ {G'}\circ \mathrm{proj}_2, \refl{\sh_G'})$
  (where $F$ is the equivalence defined in \cref{lem:SubG=MonoG}) in
  $\typemono(G\times G')$, as promised.
\end{example}
    
    
\subsection{The Lagrange construction}

In this section we give a general version of Lagrange's Theorem.
It serves as a basis for more traditional versions, such as the
counting version in \cref{xca:lagrange} below.

\begin{construction}\label{con:lagrange}
Let $G$ be a group. For every subgroup $(X,\pt):\Sub(G)$ of $G$,
with underlying group called $H$, we have a function $L_H$ of type
\[
\bigl(\prod_{x:X(\sh_G)}\sum_{g:\USymG} g\cdot_X x = \pt \,\bigr) \to
\bigl(\USymG \eqto (X(\sh_G) \times \USymH)\bigr).
\]
\end{construction}

\begin{implementation}{con:lagrange}
Define the map $[\blank] : \USymG \to X(\sh_G)$ by $[g]\defeq g\cdot_X \pt$
for all $g:\USymG$. Then \cref{lem:sum-of-fibers} yields an
equivalence from $\USymG$ to the sum of fibers
$\sum_{x:X(\sh_G)}\inv{[x]}$.
For every $x:X(\sh_G)$, the fiber $\inv{[x]}$ of $[\blank]$ at $x$ is 
$\sum_{g:\USymG} (x = g\cdot_X \pt)$, and the latter subset
of $\USymG$ is equivalent to the set $(\sh_G,\pt)\eqto(\sh_G,x)$.
So we get an equivalence from $\USymG$ to 
$\sum_{x:X(\sh_G)}((\sh_G,\pt)\eqto(\sh_G,x))$. We are done if
we can replace this irritating little last $x$ with $\pt$, since
$\USymH\defeq((\sh_G,\pt)\eqto(\sh_G,\pt))$.
We use the premiss 
$\prod_{x:X(\sh_G)}\sum_{g:\USymG} (g\cdot_X x = \pt)$.
Applying \cref{xca:AC-in-TT} to this premiss, we obtain
a function $g:X(\sh_G)\to\USymG$ such that $g(x)\cdot_X x = \pt$
for all $x:X(\sh_G)$. In other words, $(g(x),!)$ is a path of type
$(\sh_G,x)\eqto(\sh_G,\pt)$, and hence postconcatenation\footnote{%
\label{ft:lagrange-dep-sum}
Preconcatenation with the inverse gives an equivalence between
$(\sh_G,\pt)\eqto(\sh_G,x)$ and $(\sh_G,x)\eqto(\sh_G,x)$,
leading to the equivalence $L'_H$ in \cref{cor:lagrange-dep-sum}.}
gives the desired equivalence between $(\sh_G,\pt)\eqto(\sh_G,x)$ and 
$(\sh_G,\pt)\eqto(\sh_G,\pt)$. Thus we get in total an equivalence  
between $\USymG$ and $X(\sh_G) \times \USymH$, and we define $L_H(g)$
to be that equivalence.\footnote{This construction also works 
for \inftygps acting on types. However, the premiss may be harder to
fulfill in such general cases.}
\end{implementation}

A minor modification of the above implementation,
indicated in \cref{ft:lagrange-dep-sum} gives \cref{cor:lagrange-dep-sum}, 
which is sometimes more convenient, \eg in the proof of \cref{lem:burnside}.

\begin{construction}\label{cor:lagrange-dep-sum}
  Let conditions be as in \cref{con:lagrange}. Then, for each
  $f : \prod_{x:X(\sh_G)}\sum_{g:\USymG} (g\cdot_X x = \pt)$, we have an
  equivalence $L'_H(f)$ between $\USymG$ and
  $\sum_{x:X(\sh_G)}((\sh_G,x)\eqto(\sh_G,x))$.
\end{construction}

\begin{xca}\label{xca:lagrange}
The goal of this exercise is to state and prove the traditional
formulation of  Lagrange's Theorem. 
Let $G$ be a finite group and $(X,\pt):\Sub(G)$ a subgroup,
whose underlying group we call $H$. 
Assume that $X$ is a finite $G$-set. Show that $H$ is finite and
that $\Card(G) = \Card(X) \times \Card(H)$.
\end{xca}

\begin{xca}\label{xca:lagrange-Z-action-Rm}
The goal of this exercise is to illustrate that \cref{con:lagrange}
also can be applied to infinite groups. 
Recall the group of integers $\ZZ\jdeq\mkgroup{(\Sc,\base)}$ and the $\ZZ$-set
$R_m: \Sc\to\Set$ from \cref{def:RmtoS1}, defined by $R_m(\base) \defeq \bn m$
and $R_m(\Sloop) \defis \zs$, for $m>0$. Let $H_m$ be the underlying group
of $(R_m,0,!)$. Identify $\USym\ZZ$ with $\bn m \times \USymH_m$.
\end{xca}

\begin{xca}\label{xca:lagrange-if-subgr-not-normal}
  Recall the example from \cref{fig:not-normal}. The group here is
  $G\defeq \mkgroup (\Sc \vee \Sc)$, the free group on two generators
  (see~\ref{rem:F2-first-look}). Consider the $G$-set $X$ constructed by
  induction as follows: the base point $\base$ is mapped to $\bn 3 = \{0,1,2\}$;
  the first loop $\Sloop_1$ is mapped to the cyclic successor
  $s: \bn 3 \to \bn 3$; the second loop $\Sloop_2$ is mapped to the
  transposition that swaps $0$ and $1$ and leaves $2$ untouched. Our subgroup of
  interest here is $(X,0):\Sub(G)$. Find a ``choice map''
  $f:\prod_{x:\bn 3}\sum_{g:\USymG} (g\cdot x = 0)$ so that Lagrange's
  construction apply.\footnote{We shall later see that the underlying group of
    $(X,0)$ is a free group as well, see \cref{thm:nielsen-schreier}.}
\end{xca}


    % commented out by BID 211117 Some examples and references should be included when the cyclic subgroups are fully developed
    % \subsection{The geometry of subgroups: some small examples}\footnote{this subsection is not touched: it needs attention}
% \label{smallsubgpex}

% As a teaser, and in order to get a geometric feel for the subgroups and their intricate interplay, it can be useful to have some fairly manageable examples to stare at.
% Some of the main tools for analyzing the geometry of subgroups are collected in \cref{sec:fingp} on finite groups, and we hope the reader will be intrigued by our mysterious claims and go on to study \cref{sec:fingp}.
% That said, the examples we'll present are possible to muddle through by hand without any fancy machinery, but brute force is generally not an option and even for the present examples it is not something you want to show publicly.

% When presenting the subgroups of a group $G$, three types are especially revealing: the set of subgroups $\Sub(G)(\sh_G)$, the \emph{groupoid of subgroups} $\typesubgroup(G)\defequi\sum_{y:\BG}\Sub(G)(y)$ and what we for now call the ``set of normal subgroups'' $\prod_{y:\BG}\Sub(G)(y)$.   Our local use of ``normal subgroup'' is equivalent to the official definition to come.

% The first projection $\typesubgroup(G)\to \BG$ is referred to as the \emph{\covering of subgroups}.

% \footnote{Write out and fix the concrete examples (cyclic groups and $\Sigma_3$) commented out}
% % \begin{remark}
% % In  \cref{cha:circle} we studied the subgroups of the group of integers $G\eqto{}\ZZ$ through \coverings over the circle $S^1$ (which we showed was equivalent to $B\ZZ$).
% % We discovered a subgroup $n\ZZ$ for each natural number $n:\NN$ and in the groupoid $\typesubgroup({\ZZ})$ these sit as elements in separate components.  Each of these components are contractible (because addition is commutative: $\ZZ$ is an abelian group).

% % In general, a component $K$ of the groupoid $\sum_{y:\BG}\Sub(G)(y)$ of subgroups of a group $G$ may be much more interesting. For one thing the, $K$ can contain many subgroups in the sense that the preimage of the first projection $K\to \BG$ is a set that may have many different elements; each representing a subgroup.  However, this set of subgroup will be a \emph{conjugacy class} of subgroups: the different subgroups are related by the conjugation action of $G$.

% % If $G$ is abelian this action is trivial, and $\sum_{y:\BG}\Sub(G)(y)$ consists of contractible components indexed over the subgroups of $G$.  Otherwise different subgroups may live in the same component of the groupoid of subgroups -- we'll see examples in a moment.

% % In addition, the components will not in general be contractible, revealing the symmetries of the subgroups under the conjugation action.
% % \end{remark}


% % \begin{example}
% %   The trivial group only has itself as a subgroup; the groupoid of subgroups and the set of normal subgroups are singletons.
% % \end{example}
% % \begin{example}
% %   The cyclic group $C_p$ of prime order $p$ has only two subgroups, the trivial and the full subgroup itself and both are normal.  In fact, all subgroups of abelian groups are normal.

% % In general, the cyclic group $C_n$ of order $n$ has exactly one subgroup for each divisor $i$ of $n$.
% % \end{example}


% % \begin{example}
% %   The group $C_2\times C_2$ has has no less than five subgroups: the trivial one, three subgroups that as groups (as opposed as \emph{sub}groups) are equivalent to $C_2$ and the full group $C_4$ itself.
% % \end{example}
% % \begin{remark}
% %   The permutation group $\Sigma_3$ has four nontrivial proper subgroups.  Three conjugate subgroups isomorphic as groups to $C_2$ and one normal one which is as a group is isomorphic to $C_3$.  The component containing the copies of $C_2$ is equivalent to a circle.
% % \end{remark}

\section{Invariant maps and orbits}
\label{sec:fixpts-orbits}
We now return to some important constructions involving $G$-sets 
for a group $G$.
Some of these make equally good sense for \emph{$G$-types} for \aninftygp
$G$, in which case we add a footnote to this effect.

We are particularly interested in what happens when a $G$-set
is not transitive, that is, does not satisfy the requirement of 
\cref{def:transitiveGset}. In \cref{cha:circle}, under the name
of \coverings over the circle, we have already seen examples of
transitive and non-transitive $\ZZ$-sets:
In \cref{fig:covering} the left picture exhibits a transitive one,
and the right picture  a non-transitive one. Also, 
\cref{fig:two-comp-S1-cover} shows a non-transitive $\ZZ$-set,
whereas the $m$\th power bundle over the circle in
\cref{fig:m-th-power} is a transitive $\ZZ$-set.
\cref{lem:conistrans} gives a good explanation of these pictures:
A $G$-set is transitive if and only if the associated \covering is connected.
In other words, if a $G$-set $X:\BG\to\Set$ is transitive, then the
group action connects\footnote{In the sense that $\Trunc{(z,x)\eqto(w,y)}$
if and only if there exists a $g:z\eqto w$ such that $g\cdot_X x = y$.} 
any two elements in the total type $\Tot(X) \jdeq \sum_{x:\BG} X(z)$.
If $X$ is not transitive, then the latter total type falls apart 
in different components. Since these components are themselves
connected, the choice of an element of them gives rise to a subgroup of $G$
in the sense of \cref{def:set-of-subgroups}.

% perhaps with some warnings $\USymG$, $=$, \cref{lem:conistrans}, ...
\begin{definition}\label{def:actiontype} 
  \begin{enumerate}
  \item Let $G$ be a group and $X : \BG\to\Set$ a $G$-set,\footnote{%
      This definition can be generalized to \inftygps $G$ and $G$-types $X$.}
    then the \emph{action type}\index{action type}
    of $X$, denoted\footnote{%
      The superscripts and subscripts are decorated with ``$hG$'',
      following a convention in homotopy theory.
      The action type is sometimes denoted $X \dblslash G$.}
    \[
      X_{hG} \defeq \sum_{z:\BG} X(z),
    \]
    is the total type of $X$, see \cref{sec:sum-types}.
   
  \item The type of \emph{invariant maps}\footnote{%
      Invariant maps are dependent functions $f$ and the reason for the new name in
      this context is that $f(z) = g \cdot_X f(z)$ for any $z:\BG$ and $g:z\eqto
      z$. Cf.~\cref{lem:fixed-char}. Note that there need not be any invariant maps:
      the type of invariant maps of $\princ{\ZZ}$, \ie $\prod_{z:\Sc}\base\eqto z$,
      is empty.  Using \cref{lem:univisexp}, \cref{fig:transportalongloop} explains
      why: the successor function has no fixed point.%
    } \index{invariant map type} is
    \[
      X^{hG} \defeq \prod_{z:\BG} X(z).
    \]

  \item The \emph{set of orbits}\index{set!of orbits}\index{orbit set} is the
    subset of $\Sub_G(X)$ consisting of all $G$-subsets $P$ of $X$ such that the
    underlying $G$-subset $X_P$ is transitive:\footnote{See
      \cref{def:transitiveGset}.}
    \[
      X / G \defeq \sum_{P:\Sub_G(X)}\istrans(X_P).\qedhere
    \]
  \end{enumerate}
\end{definition}

\begin{remark}\label{rem:path-in-action-type}
  Note that, by \cref{def:pathover-trp} and \cref{def:pairtopath}, we get an
  equivalence that allows for a more practical form of identifications in the
  action type:
  \[
    \bigl((z,x)\eqto_{X_{hG}}(w,y)\bigr) \equivto
    \sum_{g:z\eqto w} g\cdot x = y,
  \]
  which also goes for their (often used) propositional truncations.
\end{remark}

These various definitions are closely related. For example, we will soon
identify the set of orbits with the set truncation of the action type (see
\cref{lem:X/G=setTruncX_hG}).

We have seen many instances of action types before:
When $G$-sets are considered as \coverings $f : A \to \BG$,
they are the domains $A$.
Recall for example~\cref{fig:two-comp-S1-cover},
showing an action of $\ZZ$ on $\set{1,2,3,4,5}$ with no invariant maps
and an action type equivalent to a sum of two circles.
In~\cref{fig:ZZ-set-orbits}, we show a similar $\ZZ$-set,
with underlying set $\set{0,1,2,3,4,5}$, three orbits,
and $5$ corresponding to the only invariant map.\footnote{%
Sending $\base$ to $5$ and $\Sloop$ to $\refl{5}$.}

\begin{marginfigure}
  \begin{tikzpicture}[scale=.15]
    \node (Sc) at (0,-5) {$\B\ZZ$};
    \node[dot,label=left:$5$] (five)   at (-10,30) {};
    \node[dot,label=left:$4$] (four)   at (-10,22) {};
    \node[dot,label=left:$3$] (three)  at (-10,18) {};
    \node[dot]                (base)   at (-10,-5) {};
    \node[label=left:$\Sloop$] (Sloop) at (10,-5) {};

    \pgfmathsetmacro\cc{.55228475}% = 4/3*tan(pi/8)
    \pgfmathsetmacro\cy{2*\cc}%
    \pgfmathsetmacro\cx{10*\cc}%
    \pgfmathsetmacro\intx{3.5}%
    \pgfmathsetmacro\inty{1.5}%
    \pgfmathsetmacro\ay{.35165954}%

    \draw (-10,18) .. controls (-10,18 - \cy + \ay) and (-\cx,17 - \ay)
    .. (0,17) .. controls (\cx,17 + \ay) and (10,20 - \cy - \ay) .. (10,20)
    .. controls (10,20 + \cy + \ay) and (\cx,23 - \ay)
    .. (0,23) .. controls (-\cx,23 + \ay) and (-10,22 + \cy - \ay)
    .. (-10,22) .. controls (-10,22 - \cy + \ay) and (-\cx,21 - \ay)
    .. (0,21) .. controls (\cx,21 + \ay) and (10,24 - \cy - \ay)
    .. (10,24)
    .. controls (10,24 + \cc) and (\intx + \cx, 24 + \inty + \ay)
    .. (\intx,24 + \inty) .. controls (\intx - \cx,24 + \inty - \ay)
    and (-\intx + \cx,20 + \ay)
    .. (-\intx,18 + \inty) .. controls (-\intx - \cx,18 + \inty - \ay)
    and (-10,18 + \cc) .. (-10,18);
    \draw[casblue] (-10,2) .. controls (-10,2 - \cy + \ay) and (-\cx,1 - \ay)
    .. (0,1) .. controls (\cx,1 + \ay) and (10,4 - \cy - \ay)
    .. (10,4)
    \foreach \y in {4,8} {
      .. controls (10,\y + \cy + \ay) and (\cx,3 + \y - \ay)
      .. (0,3 + \y) .. controls (-\cx,3 + \y + \ay) and (-10,2 + \y + \cy - \ay)
      .. (-10,2 + \y) .. controls (-10,2 + \y - \cy + \ay) and (-\cx,1 + \y - \ay)
      .. (0,1 + \y) .. controls (\cx,1 + \y + \ay) and (10,4 + \y - \cy - \ay)
      .. (10,4 + \y) }
    .. controls (10,12 + \cc) and (\intx + \cx, 12 + \inty + \ay)
    .. (\intx,12 + \inty) .. controls (\intx - \cx,12 + \inty - \ay)
    and (-\intx + \cx,4 + \ay)
    .. (-\intx,2 + \inty) .. controls (-\intx - \cx,2 + \inty - \ay)
    and (-10,2 + \cc) .. (-10,2);
    \draw (10,-5) .. controls ++(0,\cy) and ++(\cx,0)
    .. (0,-3) .. controls ++(-\cx,0) and ++(0,\cy)
    .. (-10,-5) .. controls ++(0,-\cy) and ++(-\cx,0)
    .. (0,-7) .. controls ++(\cx,0) and ++(0,-\cy) .. (10,-5);

    \draw (10,30) .. controls ++(0,\cy) and ++(\cx,0)
    .. (0,32) .. controls ++(-\cx,0) and ++(0,\cy)
    .. (-10,30) .. controls ++(0,-\cy) and ++(-\cx,0)
    .. (0,28) .. controls ++(\cx,0) and ++(0,-\cy) .. (10,30);

    \node[dot,label=left:$2$,casred] (two)   at (-10,10) {};
    \node[dot,label=left:$1$,casred] (one)   at (-10, 6) {};
    \node[dot,label=left:$0$,casred] (zero)  at (-10, 2) {};
  \end{tikzpicture}
  \caption{A $\ZZ$-set with three orbits and one invariant map.}
  \label{fig:ZZ-set-orbits}
\end{marginfigure}

In~\cref{fig:ZZ-set-orbits} we have highlighted one single component of
the action type in blue (\ie corresponding to an element of the set of orbits),
and we see that it contains a subset of the underlying set,
the three red elements $\set{0,1,2}$.
Such a set is what is traditionally called an orbit.
This connection is emphasized in \cref{cor:orbit-equiv}.

\begin{definition}\label{def:orbit-map}
Let $G$ be a group and $X : \BG\to\Set$ a $G$-set.
We define the map $[\blank]_0$ from the action type $X_{hG}$ of
$X$ to $\Sub_G(X)$, the set of $G$-subsets of $X$, as follows.
For any $u:X_{hG}$, let $[u]_0$ be the $G$-subset of $X$ that
sends $z:\BG$ to
\[
(x \mapsto \Trunc{u\eqto(z,x)}) : X(z) \to \Prop.\qedhere
\]
\end{definition}

Recall from \cref{rem:path-in-action-type} the equivalence of
$\Trunc{(z,x)\eqto(w,y)}$ and $\exists_{g:z\eqto w}(g\cdot x = y)$.
The next lemma follows easily from the properties of $\Trunc{u\eqto(z,x)}$.

\begin{lemma}\label{lem:[]0-maps-to-X/G}
  Let $G$ be a group and $X : \BG\to\Set$ a $G$-set.  For every $u:X_{hG}$, the
  underlying $G$-set\footnote{See \cref{def:Gsubset}.} of $[u]_0$, mapping $z$
  to $\sum_{x:X(z)}\Trunc{u\eqto(z,x)}$, is transitive.  Hence $[\blank]_0$ is
  a map from $X_{hG}$ to $X/G$.
\end{lemma}

In view of the above lemma, we call $[u]_0$ \emph{the orbit
through} $u$.\glossary(1orbit0){$[u]_0$}{orbit through $u:X_{hG}$}
The following lemma implies that the set of orbits can 
be identified with the set truncation of the action type.

\begin{lemma}\label{lem:X/G=setTruncX_hG}
  Let $G$ be a group and $X : \BG\to\Set$ a $G$-set.\footnote{%
  This lemma can be generalized to \inftygps $G$ and $G$-types $X$.}
  Then the map $[\blank]_0 :X_{hG}\to X/G $ is surjective.
  Moreover, we have a (unique) identification of
  $(X/G,[\blank]_0)$ and $(\setTrunc{X_{hG}},\settrunc{\blank})$
  in the type $\sum_{S:\Set}(X_{hG}\to S)$.
\end{lemma}

\begin{proof}
  Consider an orbit $O:X/G$, \ie
  $O$ is a $G$-subtype of $X$ such that $X_O$ is transitive. 
  We have to show that there exists a $u:X_{hG}$ such that $O=[u]_0$.
  By the connectivity of $\BG$ it suffices to show 
  $O(\sh_G)=_{X(\sh_G)\to\Prop}[u]_0(\sh_G)$ for some $u$.
  Transitivity of $X_O$ means that there exists an $x:X(\sh_G)$
  that belongs to the subset $O(\sh_G)$ and for all $y:X(\sh_G)$ in that subset
  $O(\sh_G)$ there
  exists a $g:\USymG$ such that $g\cdot x = y$, \ie $[(\sh_G,x)]_0(\sh_G)(y)$.
  Since we are proving a proposition (namely, that the preimage of $[\blank]_0$
  at $O$ is non-empty), we can assume an actual $x:X(\sh_G)$ with the property
  above, and then we take $u\defeq(\sh_G,x)$ and have to show $O(\sh_G)$ is logically equivalent to
  $[u]_0(\sh_G)$. But this follows directly
  from the observation made just above the lemma (see also
  \cref{rem:equivalents-of-[x]=[y]} below).
  
  The second part of the lemma follows from \cref{rem:set-trunc-as-quotient}. 
\end{proof}

Another way to state the above lemma is that the map 
$[\blank]_0 :X_{hG}\to X/G$ factors as the composite of $\settrunc\blank$
followed by a unique equivalence: $X_{hG}\to \setTrunc{X_{hG}}\equivto X/G$.

\begin{corollary}\label{cor:orbit-equiv}
Define the map $[\blank] : X(\sh_G) \to X/G$ by $[x]\defeq[(\sh_G,x)]_0$.
Then $[x] = [y]$ is equivalent to $\exists_{g:\USymG}(g\cdot x = y)$.
Moreover, $[\blank]$ is surjective and factors (uniquely) 
by an equivalence through the
quotient set of $X(\sh_G)$ modulo the equivalence relation 
$\exists_{g:\USymG}(g\cdot x = y)$, \cf \cref{xca:map-induces-quotient}. 
\end{corollary}

In view of this corollary, we call $[x]$ \emph{the orbit
through} $x$.\glossary(1orbit){$[x]$}{orbit through $x:X(\sh_G)$}
\begin{proof}
In the proof of surjectivity in \cref{lem:X/G=setTruncX_hG} we used
$u\defeq(\sh_G,x)$ to get $O=[u]_0$, so $[\blank]$ is surjective. 
The last statement follows since
both propositions are equivalent to $\Trunc{(\sh_G,x)\eqto(\sh_G,y)}$.
\end{proof}

\begin{remark}\label{rem:SubGX=Sub(X/G)}
Let $G$ be a group and $X : \BG\to\Set$ a $G$-set.
We have the following chain of definitions and equivalences:
\begin{align*}
\Sub_G(X)&\jdeq \prod_{z:\BG}(X(z)\to\Prop) 
\\
         & \we \bigl(X_{hG}\to\Prop\bigr)
               \quad\text{(by \cref{ft:SubTotX,xca:Sigma-curry})}
\\
         & \we \bigl(\setTrunc{X_{hG}}\to\Prop\bigr) 
               \quad\text{(since $\Prop$ is a set)}
\\        
         & \we \bigl(X/G\to\Prop\bigr)
               \quad\text{(by \cref{lem:X/G=setTruncX_hG})}
\\
         & \jdeq \Sub(X/G).\qedhere
\end{align*}
\end{remark}

\begin{xca}\label{xca:transX-just1orbit}
Show: $X/G$ is contractible if and only if $X$ is transitive.
%That $X/G$ is contractible encodes that there is just one orbit.
%Hint: use \cref{lem:conistrans}.
\end{xca}


\begin{remark}\label{rem:equivalents-of-[x]=[y]}
Given a group $G$, a $G$-set $X$ and $x,y:X(\sh_G)$,
the following propositions are all equivalent and we may pass from
one to another without mention:
\begin{itemize}
\item $[x]=_{X/G}[y]$;
\item $[x](\sh_G)=_{X(\sh_G)\to\Prop}[y](\sh_G)$;
\item $\exists_{g:\USymG}(g\cdot x = y)$;
\item $\Trunc{(\sh_G,x)\eqto(\sh_G,y)}$;
\item $[x](\sh_G,y)$;
\item $[y](\sh_G,x)$.
\end{itemize}
As functions of $x$ and $y$, all of the above define the equivalence
relation on $X(\sh_G)$ induced by the surjection $[\blank]$. 
\end{remark}

Thus, both the underlying set $X(\sh_G)$ and the action type
$X_{hG}$ have equivalence relations (induced by the surjections $[\blank]$
and $[\blank]_0$, respectively) with quotient set $X/G$.\footnote{%
  \label{ft:orbit-surj}
  This also justifies the notation $X/G$.
  We have a diagram of surjective maps:
  \[
    \begin{tikzcd}[ampersand replacement=\&]
      X(\sh_G) \ar[rr,"{x\mapsto(\sh_G,x)}"]\ar[dr,"{[\blank]}"']
      \& \& X_{hG}\ar[dl,"{[\blank]_0}"] \\
      \& X/G \&
    \end{tikzcd}
  \]}
  We can write $X(\sh_G)$ and $X_{hG}$ as sums of the respective fibers,
  which we will elaborate in the next paragraphs.

  Let $O:X/G$ be an orbit and consider
  $\inv{[O]_0} \jdeq \sum_{u:X_{hG}}(O=[u]_0)$. Note that the underlying $G$-set
  of O, \ie $X_O\jdeq(z:\BG \mapsto \sum_{y:X(z)}O(z)(y))$, is transitive. It
  follows that $O(u)$ holds\footnote{Notice the abuse here where $O(u)$ stands
    for $O(z)(x)$ where $u \jdeq (z,x)$.} if and only if $O=[u]_0$, for all
  $u:X_{hG}$.%
  \footnote{We use the first step of \cref{rem:SubGX=Sub(X/G)}.  If $O(u)$ and
    $O(v)$, then $\Trunc{u\eqto v}$ by the transitivity of $X_O$. The rest is
    obvious.}  Therefore, the fiber $\inv{[O]_0}$ is equivalent to the action
  type $(X_O)_{hG}\jdeq\sum_{z:\BG} X_O(z)$.

After the previous paragraph, the elaboration of 
$\inv{[O]} \jdeq \sum_{x:X(\sh_G)}(O=[x])$ is easy.
Recall that $[x]\jdeq[(\sh_G,x)]_0$, so that the fiber $\inv{[O]}$
is equivalent to the underlying set of $X_O$, \ie
$X_O(\sh_G)\jdeq\sum_{x:X(\sh_G)} O(\sh_G)(x)$
via identity on first components.
We depict the situation in the diagram%
in the margin. Note how the role of $X$ in \cref{ft:orbit-surj} 
is taken over by $X_O$. 

\begin{definition}\label{def:orbit-stabilizer}
  Let $G$ be a group, $X : \BG \to \Set$ a $G$-set, 
  and $x : X(\sh_G)$ an element.\footnote{%
  This definition can be generalized to \inftygps $G$ and $G$-types $X$.}
  \begin{enumerate}
  \item Define the group $G_x \defeq \Aut_{X_{hG}}(\sh_G,x)$.
% $(\sum_{z:\BG}\sum_{y:X(z)}\Trunc{(\sh_G,x) \eqto (z,y)},(\sh_G,x,!))$.
    Clearly, $\fst:\BG_x \to \BG$ is a set bundle: each fiber at $z:\BG$ is a
    subset of $X(z)$. It is pointed trivially. Hence
    $(G_x,\mkgroup(\fst)):\typemono(G)$ is a monomorphism into $G$.  We call the
    subgroup $G_x$ of $G$ the \emph{stabilizer (sub)group}\index{stabilizer}%
    \index{group!stabilizer} at $x$. We shall write $i_x : \Hom(G_x,G)$ for
    $\mkgroup(\fst)$.

    Notice that for $(X,\pt):\Sub(G)$, the group $G_\pt$ is the underlying group
    of the subgroup $(X,\pt)$, and $(G_\pt, i_\pt)$ is the corresponding
    monomorphism into $G$.
  \item Define $G\cdot x \defeq \setof{y : X(\sh_G)}{[x] =_{X/G} [y]}$
    to be the \emph{underlying set of the orbit through $x$}.\footnote{%
    This is short for the underlying set of the underlying $G$-set of the
    orbit $[x]$ of $X$.}
    \qedhere
  \end{enumerate}
\end{definition}

\begin{remark}\label{rem:orbit-fibs}
In the above definition, the underlying $G$-set of the orbit $[x]$, that is
\begin{displaymath}
  X_{[x]} \jdeq (z:\BG)\mapsto\sum_{y:X(\sh_G)}\Trunc{(\sh_G,x)\eqto(z,y)}
\end{displaymath}
plays an important double role: On one hand its action type $(X_{[x]})_{hG}$,
%$\sum_{z:\BG}\sum_{y:X(z)}\Trunc{(\sh_G,x)\eqto(z,y)}$,
pointed at $(\sh_G,x)$, is the classifying type $\BG_x$ of the stabilizer group
$G_x$. On the other hand it is a transitive $G$-set whose underlying set
$\sum_{y:X(\sh_G)}\Trunc{(\sh_G,x)\eqto(\sh_G,y)}$ is precisely $G \cdot
x$. %
In particular, $G \cdot x$ and $(\BG_x)_\div$ depends only on the class $O\jdeq[x]$,
since they can be identified with $X_O(\sh_G)$ and $(X_O)_{hG}$,
respectively. The situation is depicted in \cref{fig:fibs-at-[x]}.
\begin{marginfigure}
  \[\footnotesize %MB20250324: footnotesize should be default
    \begin{tikzcd}[ampersand replacement=\&,column sep=tiny]
      \displaystyle\sum_{O:X/G} X_O(\sh_G) \ar[rr]\ar[dr,"\fst"']
      \& \& \displaystyle\sum_{O:X/G} (X_O)_{hG}\ar[dl,"\fst"] \\
      \& X/G \& \\
      ([x], y) \ar[rr,mapsto]\ar[dr,mapsto,"\fst"']
      \& \& ([(\sh_G,x)]_0,(\sh_G,y))\ar[dl,mapsto,"\fst"] \\
      \& {[x]} \&
    \end{tikzcd}
  \]
  \caption{\label{fig:fibs-at-[x]}Along the horizontal arrow,
  the second component $y:G\cdot x$ is mapped to $(\sh_G,y):\BG_x$.}
\end{marginfigure}
Note furthermore that the base point of $\BG_x$ depends on the choice of $x$.
\end{remark}

\begin{xca}\label{xca:[x]=[y]-implies-||Gx=Gy||}
Let $G$ be a group and $X: \BG\to\Set$ a $G$-set. Show:
if $[x]=[y]$, then $\Trunc{G_x\eqto G_y}$, for any $x,y:X(\sh_G)$.
%Solution: conclusion is prop, so we may assume we have a $g:\USymG$
%such that $g\cdot x = y$. Hence we have $p: (\sh_G,x)\eqto(\sh_G,y)$,
%so $\BG_x\eqto\BG_y$ by application on this path $p$.
\end{xca}

\begin{remark}\label{rem:subgrp-is-stabsubgr}
  % In fact, every subgroup of $G$ is a stabilizer subgroup.
  The element of $\Sub(G)$ associated with the stabilizer subgroup
  $(G_x,i_x) : \typemono(G)$ of $x$ is the transitive $G$-set $X_{[x]}$, pointed
  by $x$ as element of the subset $X_{[x]}(\sh_G)$ of $X(\sh_G)$.  If $X$ is
  transitive to begin with, then all orbits are equal (\cref{xca:transX-just1orbit}) and the
  stabilizer subgroup of $x$ simplifies to $(X,x):\Sub(G)$.
\end{remark}

The following lemma states that the orbits of a $G$-set $X$
sum up to its underlying set, with the sum taken over the set
of orbits $X/G$.

\begin{lemma}
  \label{lem:splitting into orbits}
  The inclusions of the orbits form an equivalence
\[
  (O,x)\mapsto x \quad:\quad
  \bigl(\sum_{O:X/G} \inv{[O]} \bigr) \we X(\sh_G).
\]
\end{lemma}
\begin{proof}
Recall that $\inv{[O]}\jdeq \sum_{x:X(\sh_G)}(O=[x])$, 
and then abstract
away the $O$ using \cref{lem:contract-away}.\footnote{%
In fact, for every set $A$ and every equivalence relation on $A$,
the equivalence classes sum up to $A$.}
\end{proof}

There are two possible extreme cases for $G_x$ that are important:
\begin{definition}\label{def:fixed-free}
  Let $G$ be a group, $X$ a $G$-set and $x:X(\sh_G)$ an element of the underlying set.\footnote{%
  This definition can be generalized to \inftygps $G$ and $G$-types $X$.}
  We say that
  \begin{enumerate}
  \item $x$ is \emph{fixed}\index{fixed}
    if $i_x$ is an isomorphism (so ``$G_x$ is all of $G$''), and
  \item $x$ is \emph{free}\index{free}\index{action!free}
    if $G_x$ is trivial.
  \end{enumerate}
  We say that $X$ itself is \emph{free} if each $x:X(\sh_G)$ is free.
\end{definition}

\begin{example}\label{exa:fixed-free-neither}
Let $G$ be a group. For every set $S$, every element $s:S$ is fixed
under the trivial $G$-set $\triv_G S$, since the group action is the 
identity function. In contrast,
every element $g:\USymG$ is free under the $G$-set 
$\princ G \jdeq(\sh_G \eqto \blank)$, 
as $\sum_{z:\BG}\princ G(z)$ is contractible.
For an example with more variation, see \cref{exa:prep-burnside} upto
\cref{fig:C4-action-on-4-bits}. Find the fixed elements, the free elements
and those that are neither fixed nor free.
\end{example}

\begin{xca}\label{xca:fixed-free-neither}
Make sure you understand \cref{exa:fixed-free-neither} by
elaborating:
\begin{itemize}
\item $\BG_s$ (for $s:S$) in the case of $\triv_G S$,
\item $\BG_g$ (for $g:G$) in the case of $\princ G$,
\item $\BG_f$ for each $f:\bn4\to\bn2$
in the case of \cref{exa:prep-burnside}, see \cref{fig:C4-action-on-4-bits}.\qedhere
\end{itemize}
\end{xca}

\begin{lemma}\label{lem:free-pt-char}
  Let $G$ be a group and $X$ a $G$-set. Then we have for all $x:X(\sh_G)$
  that $x$ is free if and only if the (surjective) map
  $(\blank \cdot x) : \USymG \to (G\cdot x)$ is injective
  (and hence a bijection).
\end{lemma}
\begin{proof}
  Consider two elements of the orbit, say $g\cdot x,g'\cdot x$ for $g,g':\USymG$.
  We have $g\cdot x=g' \cdot x$ if and only if $x = \inv{g} g'\cdot x$
  if and only if $\inv{g} g'$ lies in $\USymG_x$.
  Hence the map $(\blank \cdot x)$ is injective iff $\USymG_x$ is contractible.
  Now use \cref{xca:connected-trivia} yielding that that 
  $G_x$ is trivial iff $\USymG_x$ is contractible.
\end{proof}

\begin{lemma}\label{lem:X_hG-set-iff-Xfree}
Let $G$ be a group and $X:\BG\to\Set$ a $G$-set.\footnote{%
This lemma can be generalized to \inftygps $G$ and $G$-types $X$.}
Then the following propositions are equivalent:
\begin{enumerate}
\item The action type $X_{hG}$ is a set;
\item The map $[\blank]_0 :X_{hG}\to X/G$ from 
      \cref{def:orbit-map} is an equivalence;
\item The $G$-set $X$ is free.
\end{enumerate}
\end{lemma}
\begin{proof}
We prove the relevant implications in circular order.
\begin{enumerate}
\item Assume $X_{hG}$ is a set. The map  $[\blank]_0 :X_{hG}\to X/G$ is
surjective by \cref{lem:X/G=setTruncX_hG}, so it suffices to show
it is injective. Since $X_{hG}$ is a set, it suffices to
show that $[u]_0=[v]_0$ implies $u=v$, for all $u,v:X_{hG}$.
This follows immediately from the definition of $[\blank]_0$, 
as the propositional truncation plays no role.
\item Assume $[\blank]_0 :X_{hG}\to X/G$ is an equivalence.
Then $X_{hG}$ is a set since $X/G$ is a set, and hence
all components of $X_{hG}$ are contractible.
It follows that all stabilizer groups $G_x$ are trivial,
and hence $X$ is free.
\item Assume $X$ is free. Then all stabilizer groups $G_x$ are trivial,
so all identity types $(\sh_G,x)\eqto (\sh_G,x)$ are contractible.
Since $\BG$ is connected we get that $u\eqto u$ is contractible
for all $u:X_{hG}$. Hence $X_{hG}$ is a set.\footnote{%
A type $T$ is a set 
if and only if all identity types $t\eqto_T t$ are contractible.}\qedhere
\end{enumerate}
\end{proof}

\begin{lemma}\label{lem:fixed-char}
  Given a group $G$ and a $G$-set $X$, an element $x:X(\sh_G)$ is
 fixed if and only if the orbit $G\cdot x$ is contractible,
  \ie $x = g\cdot x$ for all $g:\USymG$.\footnote{%
  This lemma can be generalized to \inftygps $G$ and $G$-types $X$.
  In that case $\USymG\defeq\loops\BG$ is the underlying \emph{type} of $G$.}
\end{lemma}
\begin{proof}
  The orbit $G\cdot x$ of $x$ is the fiber of $\Bi_x : \BG_x \ptdto \BG$
  at $\sh_G$. Since $\BG$ is connected,
  this is contractible if and only if all fibers of $\Bi_x$ are contractible,
  \ie $\Bi_x$ is an equivalence, which in turn is equivalent to $i_x$
  being an isomorphism.
\end{proof}

When $X : \BG \to \Set$ is a $G$-set for an ordinary group $G$, the subset
\[
  \setof{x : X(\sh_G)}{\text{$x$ is fixed}}
\]
is closely related to the type $X^{hG}$ of invariant maps.
If we evaluate an invariant map $f : \prod_{z:\BG}X(z)$ at $\sh_G$
we do indeed land in this subset:
Letting $x\defeq f(\sh_G)$,
and taking the dependent action on paths,
$\apd{f}(g) : \pathover x X g x$,
we can use~\cref{def:pathover-trp} to conclude
$\trp[X] g(x)\jdeq g\cdot x = x$, for all $g:\USymG$.
The following lemma states that, conversely, each fixed $x$ uniquely
determines an invariant map.

\begin{lemma}\label{lem:fixpts-are-fixed}
  Let $G$ be a group and $X$ a $G$-set,% \marginnote{\MB{%
  % We use the connectivity of $\BG$ and that $X(\sh_G)$ is a set.}}
  % with $X^{hG} \jdeq \prod_{z:\BG}X(z)$
  % the set of invariant maps
  . 
  The evaluation map $\ev : X^{hG} \to X(\sh_G)$ at $\sh_G$, defined by $\ev (f) \defeq f(\sh_G)$,
  \begin{enumerate}
  \item\label{it:ev-is-inj} 
  is injective, and
\item\label{it:ev-is-eq-on-inv} has image $\setof{x :
    X(\sh_G)}{\textup{$x$ is fixed}}.$
  \end{enumerate}
\end{lemma}
\begin{proof}
  Let $f,g : X^{hG}$ be invariant maps such that $f(\sh_G) = g(\sh_G)$. Since
  $\BG$ is connected, this inplies $f(z) = g(z)$ for all $z:\BG$. By
  extensionality for dependent functions, we get $f = g$. This thus proves
  \ref{it:ev-is-inj}.
  
  For \ref{it:ev-is-eq-on-inv}, assume that
  $x : X(\sh_G)$ is fixed, so that $\Bi_x\jdeq\fst:\BG_x \to \BG$
  is an equivalence. This means that $\inv\fst(z)$ is contractible
  for all $z:\BG$. Spelling out $\inv\fst(z)$, and using \cref{lem:contract-away},
  we get an identification of $\inv\fst(z)$ with
  $\sum_{y:X(z)}\Trunc{(\sh_G,x)\eqto(z,y)}$.
  Projecting on the first component of each center of contraction
  gives a dependent map $f:X^{hG}$ such that $\Trunc{(\sh_G,x)\eqto(\sh_G,f(\sh_G))}$,
  from which the proposition $x=f(\sh_G)$ follows.
\end{proof}

\begin{xca}\label{xca:Gset-A->B}
Let $G$ be the group $\SG_2\times\SG_2$ and $X$ the $G$-set mapping
any pair $(A,B)$ of $2$-element sets to the set $A\to B$.
Elaborate the action of $G$ on $X(\sh_G)$ and determine the
set of orbits and the set of invariant maps. You can
do the same exercise for the following easier cases first:
the $G$-set that is constant $\bn2\times\bn2$, and
the $\SG_2$-sets $X(\blank,\bn2)$ and $X(\bn2,\blank)$.
\end{xca}

\subsection{The Orbit--Stabilizer theorem}
\label{sec:orbit-stabilizer-theorem}

Consider a group $G$, a $G$-set $X$ and an element $x:X(\sh_G)$,
and recall \cref{def:orbit-stabilizer}.
The classifying type of the stabilizer group $G_x$
is the component of $X_{hG}\jdeq\sum_{z:\BG} X(z)$ 
pointed by the shape $(\sh_G,x)$.
The first projection of a symmetry of $(\sh_G,x)$ is a symmetry of 
$\sh_G$, and the second projection is a proof of a proposition.
This suggest the following simple way for $G_x$ to act on the
symmetries of $\sh_G$, by just ignoring the second projection:


\begin{definition}\label{def:Gx-action-on-G}
  Let $G$ be a group, $X$ a $G$-set and $x:X(\sh_G)$ an element of the
  underlying set. Recall that the stabilizer group $G_x$ comes with a
  monomorphism $i_x : \Hom(G_x,G)$. Define the $G_x$-set
  $\tilde G_x : \BG_x \to \UU$ as the restriction of $\princ G$ along $i_x$, \ie
  $\tilde G_x \defeq \princ G \circ \Bi_x$.\footnote{%
    Spelled out: for all $(z,y):X_{hG}$ in the same component as $(\sh_G,x)$,
    $\tilde G_x(z,y)\jdeq(\sh_G\eqto z)$.  In \cref{def:restrictandinduce} we
    will give a general definition of restrictions of $G$-sets along
    homomorphisms in $\Hom(G,H)$.}
\end{definition} 

The underlying set of $\tilde G_x$ is $\USymG$. The group action of
$\tilde G_x$ is explored in the following exercise.

\begin{xca}\label{xca:Gx-action-on-G}
Let $s:(\sh_G,x,!)\eqto(z,y,!)$ be a path in $\BG_x$ with first component
$s_1$, and let $g:\USymG$. Show that $s\cdot_{\tilde G_x} g = s_1 g$, \ie
the group action of $\tilde G_x$ is path composition. 
\end{xca}

The following exercise prepares for the subsequent Orbit--Stabilizer Theorem.

\begin{xca}\label{xca:fixed-free-neither-action-types}
Elaborate the action type of $\tilde G_x$ from \cref{def:Gx-action-on-G}
in each of the cases of \cref{xca:fixed-free-neither}, that is, elaborate
\begin{itemize}
\item $(\tilde G_s)_{hG_s}$ in the case of $\triv_G S$,
\item $(\tilde G_g)_{hG_g}$ in the case of $\princ G$, 
\item $(\tilde G_f)_{hG_f}$ for each $f:\bn4\to\bn2$,
in the case of \cref{exa:prep-burnside}, \cref{fig:C4-action-on-4-bits}.
\end{itemize}
Compare your findings with $G \cdot s$, $G \cdot g$,
and each $G \cdot f$, respectively.
\end{xca}

The action type of $\tilde G_x$ can be identified with the 
underlying set of the orbit through $x$ under $X$. This is 
achieved by a chain of easy equivalences, spelled
out in the following construction.

\begin{construction}[Orbit--Stabilizer theorem]
  \label{con:orbit-stabilizer}
  Let $G$ be a group, $X$ a $G$-set $X$,
  $x : X(\sh_G)$ an element of the underlying set of $X$.\footnote{%
  This construction can be generalized to \inftygps $G$ and $G$-types $X$.}
  Recall the $G_x$-set $\tilde G_x$ from \cref{def:Gx-action-on-G}.
  Then we have an equivalence from the action type $(\tilde G_x)_{hG_x}$
  to the underlying set $(G\cdot_X x)$ of the orbit through $x$.
\end{construction}
\begin{implementation}{con:orbit-stabilizer}
  The desired equivalence is the composition of elementary equivalences
  for sums and products, followed by contracting away the variable $z$:
  \begin{align*}
    (\tilde G_x)_{hG_x}
    &\jdeq \sum_{u : \BG_x}\tilde G_x(u) \\
    &\equivto \sum_{z:\BG}\sum_{y:X(z)}
      \Trunc{(\sh_G,x) \eqto (z,y)}\times (\sh_G \eqto z) \\
    &\equivto \sum_{y:X(\sh_G)}[x] =_{X/G} [y]\quad\jdeq\quad(G\cdot_X x).\qedhere
  \end{align*}
\end{implementation}

The above construction has some interesting consequences.  One is that
$(\tilde G_x)_{hG_x}$ is a set\footnote{\label{ft:action-type-tildeGx-set} Note
  that \cref{xca:Gx-action-on-G} already implies that $(\tilde G_x)_{hG_x}$ is a
  set: given $(\sh_G,x,!,g)$ and $(z,y,!,g')$, there can be at most one
  $s : \sh_G \eqto z$ such that $s\cdot_{\tilde G_x} g = sg = g'$.}, so that
\cref{lem:X_hG-set-iff-Xfree} applies:

\begin{corollary}\label{cor:action-subgrp-free}
The $G_x$-set $\tilde G_x$ is free.
\end{corollary}

We further obtain that the underlying set of the orbit of $\tilde G_x$ through $g$ can be
identified with the underlying set of $G_x$.

\begin{corollary}\label{lem:cosets-Gx.g}
%  Let $G$ be a group and $X$ a $G$-set. Given $x:X(\sh_G)$,
%  consider the stabilizer subgroup $G_x$ and the $G_x$-set $\tilde G_x$.
  For any $g:\USymG$, the map $(\blank \cdot_{\tilde G_x} g)$ is
  an equivalence from $\USymG_x$ to $(G_x \cdot_{\tilde G_x} g)$.
\end{corollary} 
 
\begin{proof}
This follows directly from \cref{lem:free-pt-char}, 
applied to $G_x$ and $\tilde G_x$, using that $\tilde G_x$ is free.
\end{proof}

In the case of a subgroup of $G$, we have the following result.\footnote{In
    ordinary group theory, a subgroup $H$ of a group $G$ always acts on $G$ by
    multiplication: this is the $G_{\pt}$-set $\tilde G_{\pt}$ in our
    presentation. The set of orbits for this action is the set $G/H$ of cosets:
    in our presentation, this is $\tilde G_{\pt} /G_{\pt}$. Each element of
    $G/H$ is of the form $Hg$ for some $g\in G$ and is in bijection with $H$:
    such an element is $G_{\pt} \cdot g$ is our presentation, and the
    bijectivity is \cref{lem:cosets-Gx.g}. The group $G$ acts transitively on
    the set of cosets $G/H$: this is the $G$-set $X$ in our presentation.}

\begin{construction}\label{con:preLagrange}
  Let $G$ be a group and let $(X,\pt):\Sub(G)$ be a subgroup of $G$ as defined
  in \cref{def:set-of-subgroups}. Then we have an equivalence
  $[\blank]_{\pt}$ from the underlying set $X(\sh_G)$ of $X$ to 
  $\tilde G_{\pt} /G_{\pt}$, the set of orbits of $\tilde G_{\pt}$. 
\end{construction}
\begin{implementation}{con:preLagrange}
The function $[\blank]_{\pt}$ is the composition of three equivalences.
Since $X$ is transitive, $\fst: (G\cdot_X {\pt})\to X(\sh_G)$ is an equivalence.
The Orbit--Stabilizer \cref{con:orbit-stabilizer} gives us an equivalence $o$
from $(G\cdot_X {\pt})$ to $(\tilde G_{\pt})_{hG_{\pt}}$. Since the latter type is a set,
the function $[\blank]_0$ from \cref{lem:X/G=setTruncX_hG} 
is an equivalence from $(\tilde G_{\pt})_{hG_{\pt}}$ to $\tilde G_{\pt} /G_{\pt}$.
Now define $[x]_{\pt} \defeq [o(\inv\fst(x))]_0$ for any $x:X(\sh_G)$.
\end{implementation}

The reader may notice that the last two results contain some of
the ingredients of the traditional formulation of Lagrange's Theorem:
the group $G$, the subgroup $G_{\pt}$, the orbits 
(cosets) $(G_{\pt} \cdot_{\tilde G_{\pt}} g)$ 
and the set of orbits $\tilde G_{\pt} /G_{\pt}$.
It is in fact possible to obtain the counting version of
Lagrange's Theorem, \cref{xca:lagrange}, from the above results:

\begin{xca}\label{xca:lagrange2}
Let $G$ be a finite group and $(X,\pt):\Sub(G)$ a subgroup,
whose underlying group we call $H$. 
Assume that $X$ is a finite $G$-set. Show that 
$\Card(G) = \Card(X) \times \Card(H)$ using \cref{lem:splitting into orbits},
\cref{lem:cosets-Gx.g} and \cref{con:preLagrange}, instead of
\cref{con:lagrange}.
\end{xca}

\section{The classifying type is the type of torsors}
\label{sec:torsors}
Recall the definition of the principal $G$-torsor
$\princ G \jdeq (\sh_G \eqto \blank)$ from \cref{def:principaltorsor}. In this
section we elaborate the concept of {\em torsor} and give one example of its
use. In \cref{sec:Gsetforabstract} we'll use abstract torsors to prove that the
type of groups and the type of abstract groups are equivalent by classifying
abstract groups via their pointed connected groupoid of abstract torsors. In the
end we want to prove that the type of torsors of $\abstr(G)$ is equivalent to
$\BG$. To do so we will need an intermediate equivalent type, namely that of
torsors for the concrete group $G$. We will then prove~\cref{lem:BGbytorsor},
establishing that showing that $\BG$ is equivalent to the type of
$G$-torsors. In \cref{sec:Gsetforabstract} we'll then only have to show that
abtract and concrete torsors are equivalent.
\begin{definition}\label{def:Gtorsor}
  Given a group $G$, the type of \emph{$G$-torsors}%
  \index{torsor@torsor}%
  \glossary(TorsorG){$\protect\typetorsor_G$}{the type of $G$-torsors}%
  \footnote{This works equally well with $\infty$-groups: $G$-torsors are in that case $G$-types in the component of the principal torsor $\princ G:\BG\to\UU$. There is no conflict with the case when the $\infty$-group $G$ is actually a group since then any $G$-type in the component of the principal $G$-torsor will be a $G$-set.}
  is
  \[
    \typetorsor_G\defequi\sum_{X:\GSet}\Trunc{\princ G \eqto X},
  \]
  where $\princ G \jdeq (\sh_G \eqto \blank)$ is the 
  principal $G$-torsor of \cref{def:principaltorsor}.
\end{definition}

\begin{xca}\label{xca:torsor=free+transitive}
  Show that a $G$-set is a $G$-torsor if and only if it is free and transitive.
\end{xca}

\begin{remark}
  For $G$ a group, the type of $G$-torsors is just another name for the component of the type of \coverings over $\BG$ containing the universal \covering.

  Observe that for a group $G$, $\typetorsor_G$ is a connected groupoid\footnote{Admittedly in a higher universe, but we can use the
    Replacement~\cref{pri:replacement} to see that $\typetorsor_G$ is equivalent
    to a type in the same universe as $G$ -- even before we
    have~\cref{lem:BGbytorsor} showing we can take $\BG$.}
  and so -- by specifying the base point $\princ G$ -- it classifies a group.
  Guess which one!\footnote{%
    By the way, the name ``torsor'' is a translation from the French \emph{torseur},
    introduced by \citeauthor{giraud1971},\footnotemark{} who
    related them to “twisting” operations on bundles.
    Since $BG$ is equivalent to the type of $G$-torsors,
    we can also think of shapes $t:\BG$ as giving rise to “twists”.
    Indeed, for a $G$-set $X$,
    we can think of $X(x)$ as a ``twisted'' version of the underlying set,
    $X(\sh_G)$.}\footcitetext{giraud1971}.
\end{remark}


\begin{definition}
  \label{def:BG2TorsG}
  Recall\footnote{See \cref{def:principaltorsor} and \cref{eq:pathsp}.} that the
  $G$-set $\pathsp y:\BG\to\Set$ is defined for any $y:\BG$ by
  $\pathsp y(z)\jdeq(y\eqto z)$.  Note that $\pathsp y$ is a $G$-torsor since
  its action type $(\pathsp y)_{hG}$ is contractible, so we can define
  \[
    \pathsp{\blank}:\BG\ptdto(\typetorsor_G,\princ G): y\mapsto P_y,
  \]
  pointed by reflexivity.\footnote{%
    That is, we have classified a homomorphism from $G$
    to $\Aut_{\GSet}(\princ G)$. It'll turn out to be an isomorphism.}
If $G$ is not clear from the context, 
we may choose to write $\pathsp{\blank}^G$ instead of $\pathsp{\blank}$.
\end{definition}

\begin{remark}\label{rem:pathsptransport}
  We will use several variants of $\pathsp{\blank}$, in combination with some of
the conventions introduced back in \cref{ch:univalent-mathematics}.  In this
remark, to avoid confusion, we explain these variants.
  
  First, we also use $\pathsp{\blank}$ to denote its induced action on paths:
for $y,z:\BG$ we have
  \[ \pathsp{\blank}:(y\eqto z)\to (\pathsp y\eqto \pathsp z),
  \] defined by path induction as in \cref{def:ap}.
  
  Then, as $\pathsp y\eqto \pathsp z$ is an identity between families of types,
function extensionality (\cref{def:funext}) applies.  For $q:y\eqto z$, we may
also use $\pathsp q$ to denote the corresponding function of type
$\prod_{x:\BG}(\pathsp y(x)\eqto \pathsp z(x))$.
  
  Finally, as $\pathsp y(x)$ and $\pathsp z(x)$ are types, univalence
(\cref{def:univalence}) applies.  Therefore we may use $\pathsp q(x)$ to denote
the corresponding equivalence, \ie transport in the type family
$\pathsp{\blank}(x)$, sending $p:\pathsp y(x)\jdeq (y\eqto x)$ to
$pq^{-1}:\pathsp z(x)\jdeq(z\eqto x)$.\footnote{In a commutative diagram,
    \[
      \begin{tikzcd}[ampersand replacement=\&] y \ar[rr,eqr,"q"]\ar[dr,eql,"p"']
\& \& z \ar[dl,eqr,"{\pathsp q(p)}"] \\ \& x.
      \end{tikzcd}
    \]}
\end{remark}

For connoisseurs of category theory, the following lemma is a corollary of a
\emph{type-theoretic Yoneda lemma},\footnote{This is more than an analogy: using
  \cref{thm:1types-are-groupoids} and \cref{thm:yoneda-lemma}, we obtain almost
  the same result. The only difference lies in that our $\princ G$ here
  translates to a \emph{covariant} representable instead of a contravariant
  one.} and the proof is \cref{xca:TTYoneda}.\footnote{%
  It is also possible to prove the lemma directly by an application of
  \cref{lem:weq-iso}: Take as inverse equivalence the map $Q$ mapping any
  $f:\pathsp y \eqto \pathsp z$ to
  $Q(f) \defequi \inv{(f_y(\refl y))} : (y\eqto z)$.}

\begin{lemma}\label{lem:pathsptransportiseq}
  Let $G$ be a group. For all $y,z:\BG$ the induced map of identity types
  \[
    \pathsp{\blank}:(y\eqto z)\to (\pathsp y\eqto \pathsp z)
  \]
  is an equivalence.
\end{lemma}

The following theorem justifies the title of this section, stating that the
classifying type of a group is the type of its torsors.

\begin{theorem}\label{lem:BGbytorsor}
  Let $G$ be a group. Then the function $\pathsp{\blank}^G:\BG\to\typetorsor_G$
  from~\cref{def:BG2TorsG} is an equivalence.\footnote{A similar results holds
    for $\infty$-groups.}
\end{theorem}

\begin{proof}
  Since both $\typetorsor_G$ and $\BG$ are pointed and connected,
  it suffices by
  \cref{cor:fib-vs-path}\ref{conn-fib-vs-path-point} to show that 
  $\pathsp{\blank}^G:(\sh_G\eqto\sh_G)\to(\princ G\eqto \princ G)$
  is an equivalence.
  %\footnote{This holds for all variants of $\pathsp{\blank}$ from \cref{rem:pathsptransport}.}
  This follows directly from \cref{lem:pathsptransportiseq}.
\end{proof}

\subsection{Homomorphisms and torsors}
\label{sec:homotor}
In view of the equivalence $\pathsp{\blank}^G$ between $\BG$ and 
$(\typetorsor_G,\princ G)$ of \cref{lem:BGbytorsor} one might 
ask what a group homomorphism  $f:\Hom(G,H)$ translates to on 
the level of torsors.  Off-hand, the answer is the round-trip 
$(\pathsp{\blank}^H)\Bf(\pathsp{\blank}^G)^{-1}$, but we can be more concrete than that.
We do know that for $z:\BG$ the $G$-torsor $\pathsp z^G$ should be sent to
$\pathsp {\Bf(z)}^H$, but how do we express this for an arbitrary $G$-torsor?
\begin{definition}
  \label{def:restrictandinduce}
  Let $f:\Hom(G,H)$ be a group homomorphism.  If $Y:\BH\to\Set$ is an $H$-set,
  then the \emph{restriction}\index{action!restricted}\index{restriction}
  $f^*Y$ of $Y$ to $G$ is the $G$-set given by precomposition\footnote{%
  Example: $\tilde G_x$ from \cref{def:Gx-action-on-G} can
  be written as $i_x^* \princ G$, \ie as the restriction of the
  principal $G$-torsor to the stabilizer group $G_x$ using $i_x:\Hom(G_x,G)$.}
  \[
    f^*Y\defequi (Y\circ\Bf) :\BG\to\Set.
  \]

  If $X:\BG\to\Set$ is a $G$-set, we define
  the \emph{induced $H$-set}\index{action!induced}\index{induced action}
  $f_!X : \BH\to\Set$ by setting, for $w:\BH$,\footnote{%
\label{ft:f_!X(w)-orbitset}
Note that the set $f_!X(w)$ can also be identified as the set of orbits of the
$G$-set $(z:\BG) \mapsto (\Bf(z) \eqto w)\times X(z)$, whose underlying set is
equivalent to $\princ H(w)\times X(\sh_G)$.}
  \[
    f_!X(w)\defeq\setTrunc[\bigg]%
    {\sum_{z:\BG}\bigl((\Bf(z) \eqto w)\times X(z)\bigr)}.\qedhere
  \]
\end{definition}

The following exercise shows that the set-truncation in the definition
of $f_!$ above really makes a difference.\footnote{%
    This situation is common in algebra and is often referred to by saying
    that some construction, in this case the untruncated
    definiens of $f_!X$, is not ``exact''. %
    % See also \cref{xca:why-setTrunc_f_!}. <-- already referenced in text
  }

\begin{xca}\label{xca:why-setTrunc_f_!}
Find groups $G,H$, a homomorphism $f:\Hom(G,H)$ and a $G$-set $X$ such
that $(w:\BH)\mapsto\sum_{z:\BG}\bigl((\Bf(z) \eqto w)\times X(z)\bigr)$
is an $H$-type that is not an $H$-set.
\end{xca}
% Solution: $G=\ZZ$, $H=\TG$, $f$ the unique homomorphism $f: \Hom(G,H)$,
% $X$ constant $\bn 1$. Then
% $\sum_{z:\Sc}(0=0)\times \bn 1$ is a circle, so not a set.

\begin{xca}\label{xca:adjunction-_!-^*}
    Give an equivalence from $f_!X$ to $X\circ\inv\Bf$ (\ie $(\inv f)^*\,X$)
    if $f$ is an isomorphism.
    Give an equivalence between the
    types $\Hom_H(f_!X,Y)$ and $\Hom_G(X,f^*\,Y)$,
    for all $G$-sets $X$ and $H$-sets $Y$.
\end{xca}
% Solution 2nd part: elaborating $\Hom_H(f_! X,Y)$ you can leave out
% the set truncation since $Y$ is an $H$-set. Now turn the sum
% over $z:\BG$ into an upfront product. Then eliminate the product
% over $w:\BH$ by using $f(z)\eqto w$.

\begin{marginfigure}
  \noindent\begin{tikzpicture}
    % Name the coordinates so they are easy to change
    % first the base : B
    \coordinate (B-left) at (-1,-1);
    \coordinate (B-X) at (0,-1.1);
    \coordinate (B-mid) at (1,-1.3);
    \coordinate (B-Y) at (1.9,-1.2);
    \coordinate (B-right) at (2.8,-1);
    \node (B) at (3.2,-1) {$B$};
    \node (A) at (3.2,1) {$A$};
    % then: A top and bottom
    \coordinate (A-top-left) at (-1,1.1);
    \coordinate (A-bot-left) at (-1,0.7);
    \coordinate (A-top-X) at (0,1.8);
    \coordinate (A-bot-X) at (0,.1);
    \coordinate (A-top-mid) at (1,1.9);
    \coordinate (A-bot-mid) at (1,-.1);
    \coordinate (A-top-Y) at (1.9,1.8);
    \coordinate (A-bot-Y) at (1.9,0);
    \coordinate (A-top-right) at (2.8,1.2);
    \coordinate (A-bot-right) at (2.8,0.8);
    % coordinates of X
    \coordinate (X-left) at (0,1);
    \coordinate (X-mid) at (0.5,1.3);
    \coordinate (X-right) at (1,0.9);
    \draw[dashed,fill=casblue!30] (X-left) .. controls +(-90:.2) and +(-90:.2) .. (X-right)
    .. controls +(90:.2) and +(0:.2) .. (X-mid)
    .. controls +(0:-.2) and +(90:.2) .. (X-left);
    \fill[casred!30] (A-top-Y) -- (A-bot-Y)
    .. controls +(15:.5) and +(90:-.5) .. (A-bot-right)
    -- (A-top-right)
    .. controls +(90:.5) and +(-10:.3) .. (A-top-Y);
    \draw[dashed] (A-top-Y) -- (A-bot-Y);
    \draw (A-bot-left) .. controls +(-90:.5) and +(10:-.5) .. (A-bot-X)
    .. controls +(10:.5) and +(0:-.5) .. (A-bot-mid)
    .. controls +(0:.5) and +(15:-.5) .. (A-bot-Y)
    .. controls +(15:.5) and +(90:-.5) .. (A-bot-right)
    -- (A-top-right)
    .. controls +(90:.5) and +(-10:.3) .. (A-top-Y)
    .. controls +(-10:-.3) and +(0:.4) .. (A-top-mid)
    .. controls +(0:-.4) and +(-1:.4) .. (A-top-X)
    .. controls +(-10:-.4) and +(90:.5) .. (A-top-left)
    -- (A-bot-left);
    \draw (B-left) .. controls +(20:.3) and +(-10:-.3) .. (B-X)
    .. controls +(-10:.3) and +(0:-.3) .. (B-mid)
    .. controls +(0:.3) and +(10:-.3) .. (B-Y)
    .. controls +(10:.3) and +(0:-.1) .. (B-right);
    % projection map f
    \draw[->,shorten <=.1cm,shorten >=.1cm] (A)
    -- node[anchor=west] {$f$} (B);
    % subset Y of B
    \draw[line width=2pt,casred!30] (B-Y) .. controls +(10:.3) and +(0:-.1) .. (B-right);
    \node (Y) at (2.45,-1.4) {$Y$};
    \node (fY) at (2.45,1) {$f^* Y$};
    % subset X of A
    \draw[line width=2pt,casblue!30] (B-X)
    .. controls +(-10:.3) and +(0:-.3) .. (B-mid);
    \node (X) at (.5,-1.6) {$\oldexists_fX$};
    \node (fX) at (.5,1.04) {$X$};
  \end{tikzpicture}
  \caption{Image ($\oldexists_f$) and preimage ($f^*$) of subsets for a map $f : A \to B$.}
  \label{fig:im-preim-subsets}
\end{marginfigure}
\begin{remark}\label{rem:^*-_!-as-(pre)image}
The purpose of this remark is to explain how
$f_!X$ and $f^*Y$ may be viewed as a certain kind of
image and preimage, respectively. In \cref{def:prop-image}
we defined the (propositional) image of a function $f:A\to B$,
and in \cref{sec:higher-images} the higher images. In these
definitions we used the whole domain $A$ of $f$.
In \cref{def:fiber} we defined the preimage, or fiber, of $f$,
for any element $b$ of the codomain $B$.

It is natural to generalize both image and preimage of
a function to subtypes of
the domain and codomain. Let $A$ and $B$ be types, $f:A\to B$
a function, and consider the types of subtypes $\Sub(A)\jdeq(A\to\Prop)$
and $\Sub(B)\jdeq(B\to\Prop)$. Given a subtype $Y:\Sub(B)$ of $B$,
consider $f^*Y \jdeq Y\circ f$. Then $f^*Y$ is subtype of $A$
consisting of precisely those $a:A$ for which $f(a)$ is in $Y$,
in other words, the preimage of $Y$ under $f$.

Now let $X:\Sub(A)$ be a subtype of $A$ and consider the subtype of
$B$ defined by the predicate
\[
  \oldexists_fX (b) \defeq \Trunc[\Big]{\sum_{a:A}\bigl((f(a) \eqto b)\times X(a)\bigr)}
  \jdeq \exists_{a:A}\bigl((f(a) \eqto b)\times X(a)\bigr).
\]
This is $f_!X$ with propositional truncation instead of set truncation,
and it holds precisely for all $b:B$ for which there
exists an $a$ in $X$ with $f(a)\eqto b$,
in other words, the image of $X$ under $f$.
These operations are illustrated in~\cref{fig:im-preim-subsets}
where we view $f$ as a projection, so the fibers are vertical.

Note that the above makes little sense for $A\jdeq\BG$ and $B\jdeq\BH$,
since predicates on connected types are constant. However, the intuition
carries over to $\Set$ valued functions $X$ and $Y$ and a set truncated
$f_!X$, analogously to higher images defined in \cref{sec:higher-images}.
\end{remark}

\begin{remark}\label{rem:coinduced-Hset}
  Dually to $f_!X(w)$ in \cref{def:restrictandinduce},
  there is also a \emph{coinduced $H$-set}\index{action!coinduced}
  $f_*:\BH\to\Set$ given by
  \[
    f_*X(w)\defeq\prod_{z:\BG}\bigl((\Bf(z)\eqto w) \to X(z)\bigr).
  \]
  Note that this always lands in sets since $X$ does.\footnote{%
The type $f_*X(w)$ can also be identified
as the set of invariant maps of the $G$-set $\pathsp{\Bf}(w)\to X$,
where $(\pathsp{\Bf}(w)\to X)(z) \defeq (\Bf(z) \eqto w)\to X(z)$ for $z:\BG$.}
\end{remark}

\begin{marginfigure}
  \noindent\begin{tikzpicture}
    % Name the coordinates so they are easy to change
    % first the base : B
    \coordinate (B-left) at (-1,-1);
    \coordinate (B-X) at (0,-1.1);
    \coordinate (B-mid) at (1,-1.3);
    \coordinate (B-Y) at (1.9,-1.2);
    \coordinate (B-right) at (2.8,-1);
    \node (B) at (3.2,-1) {$B$};
    \node (A) at (3.2,1) {$A$};
    % then: A top and bottom
    \coordinate (A-top-left) at (-1,1.1);
    \coordinate (A-bot-left) at (-1,0.7);
    \coordinate (A-top-X) at (0,1.8);
    \coordinate (A-bot-X) at (0,.1);
    \coordinate (A-top-mid) at (1,1.9);
    \coordinate (A-bot-mid) at (1,-.1);
    \coordinate (A-top-Y) at (1.9,1.8);
    \coordinate (A-bot-Y) at (1.9,0);
    \coordinate (A-top-right) at (2.8,1.2);
    \coordinate (A-bot-right) at (2.8,0.8);
    % coordinates of X
    \coordinate (X-left) at (.2,1);
    \coordinate (X-right) at (2.45,0.9);
    \draw[dashed,fill=casblue!30] (A-top-mid)
    .. controls +(0:-.4) and +(90:.5) .. (X-left)
    .. controls +(-90:.5) and +(0:-.5) .. (A-bot-mid)
    .. controls +(0:.5) and +(15:-.5) .. (A-bot-Y)
    .. controls +(15:.5) and +(90:-.5) .. (X-right)
    .. controls +(90:.5) and +(-10:.3) .. (A-top-Y)
    .. controls +(-10:-.3) and +(0:.4) .. (A-top-mid);
    \draw[dotted] (A-top-mid) -- (A-bot-mid);
    \draw[dotted] (A-top-Y) -- (A-bot-Y);
    \draw (A-bot-left) .. controls +(-90:.5) and +(10:-.5) .. (A-bot-X)
    .. controls +(10:.5) and +(0:-.5) .. (A-bot-mid)
    .. controls +(0:.5) and +(15:-.5) .. (A-bot-Y)
    .. controls +(15:.5) and +(90:-.5) .. (A-bot-right)
    -- (A-top-right)
    .. controls +(90:.5) and +(-10:.3) .. (A-top-Y)
    .. controls +(-10:-.3) and +(0:.4) .. (A-top-mid)
    .. controls +(0:-.4) and +(-1:.4) .. (A-top-X)
    .. controls +(-10:-.4) and +(90:.5) .. (A-top-left)
    -- (A-bot-left);
    \draw (B-left) .. controls +(20:.3) and +(-10:-.3) .. (B-X)
    .. controls +(-10:.3) and +(0:-.3) .. (B-mid)
    .. controls +(0:.3) and +(10:-.3) .. (B-Y)
    .. controls +(10:.3) and +(0:-.1) .. (B-right);
    % projection map f
    \draw[->,shorten <=.1cm,shorten >=.1cm] (A)
    -- node[anchor=west] {$f$} (B);
    % subset X of A
    \draw[line width=2pt,casblue!30] (B-mid) -- (B-Y);
    \node (fX) at (1.5,-1.6) {$\forall\!_fX$};
    \node (X) at (1.5,1.04) {$X$};
  \end{tikzpicture}
  \caption{Coinduced image of a subset for a map $f : A \to B$.}
  \label{fig:coinduced-image}
\end{marginfigure}
\begin{remark}\label{rem:coinduced-subset}
  As in~\cref{rem:^*-_!-as-(pre)image}, we can also consider
  $f_*$ as an operation on subsets $f_* : \Sub(A) \to \Sub(B)$
  for $f:A\to B$. The coinduced subset $\forall\!_fX \defeq f_*X$
  is less familiar than the image; it contains those $b:B$
  for which the entire preimage $\inv f(b)$ is contained in $X$,
  as illustrated in~\cref{fig:coinduced-image}.
\end{remark}

\begin{xca}\label{xca:adjunction-^*-_*}
    Give an equivalence between the
    types $\Hom_G(f^*Y,X)$ and $\Hom_H(Y,f_*X)$,
    for all $G$-sets $X$ and $H$-sets $Y$.
\end{xca}

When $X$ is the $G$-torsor $\pathsp x^G$, for some $x:\BG$,
the contraction (recall \cref{lem:contract-away})
of $\sum_{z:\BG}(x\eqto z)$ induces an equivalence $\eta_w$ of type
\[
  f_!\pathsp x^G(w) \jdeq
  \setTrunc[\bigg]{\sum_{z:\BG}\bigl((\Bf(z) \eqto w)\times(x \eqto z)\bigr)}
  \equivto (\Bf(x)\eqto w)\jdeq\pathsp{\Bf(x)}^H(w).
\]
Taking $x\jdeq\sh_G$, we get a
path $\eta:f_!\,\princ G\eqto \pathsp{\Bf(\sh_G)}^H$.
We also have the path $\Bf_\pt : \sh_H\eqto\Bf(\sh_G)$,
so that the action of $\pathsp{\blank}^H$ gives us a path
$\pi : \princ H \jdeq\pathsp{\sh_H}^H \eqto \pathsp{\Bf(\sh_G)}^H$.
Combining we get $\inv\eta\pi:\princ H \eqto f_!\,\princ G$.

If $X$ is a $G$-set such that $\Trunc{\princ G \eqto X}$, then $f_!X$
is an $H$-set such that $\Trunc{\princ H \eqto f_!X}$, so that
$f_! : \typetorsor_G \ptdto \typetorsor_H$, pointed by $\inv\eta\pi$.

Summing up, we have implemented the following:
\begin{construction}
  \label{con:inducedtorsor}
   Let $f:\Hom(G,H)$ be a group homomorphism. Then $f$ induces a
   pointed map $f_!:\typetorsor_G\ptdto\typetorsor_H$,
   and we have a path of type
   $f_!\,\pathsp{\blank}^G \eqto \pathsp{\blank}^H\,\Bf \jdeq
   f^*\,\pathsp{\blank}^H$,
   all represented by the following diagram:
   \[
     \begin{tikzcd}
     \sh_G \ar[ddd,mapsto] \ar[rrr,mapsto] &
     && \Bf(\sh_G)  & \ar[l,eql,"{\Bf_\pt}"'] \sh_H \ar[ddd,mapsto]\\
       &\BG \ar[r,"\Bf"]\ar[d,"{\pathsp{\blank}^G}"'] &
        \BH\ar[d,"{\pathsp{\blank}^H}"] \\
       &\typetorsor_G \ar[r,"f_!"'] & \typetorsor_H \\
     \princ G \ar[rrr,mapsto] &
     && f_!\,\princ G&  \ar[l,eqr,"{\inv\eta\pi}"] \princ H.
     \end{tikzcd}
   \]
\end{construction}

\section{Any symmetry is a symmetry in $\Set$}
\label{sec:groupssubperm}


For abstract groups there is a result, attributed to Cayley, which is often
stated as ``any group is a permutation group''.  In our parlance this translates
to ``any symmetry is a symmetry in $\Set$''.  The aim of this section is to give
a precise formulation of the latter and prove it, using what we learned in
\cref{sec:torsors}.
% \footnote{which reminds me of the following: my lecturer in cosmology once
% tried to publish a paper about rotating black holes, only to have it rejected
% because it turned out that it was his universe, not the black hole, that was
% rotating}

%which is equivalent to saying that $X$ is the universal \covering
Let $G$ be a group. Recall from \cref{lem:BGbytorsor} the equivalence
\begin{displaymath}
  \pathsp{\blank}^G:\BG\equivto\typetorsor_G : 
  z \mapsto (y \mapsto (z \eqto y)).
\end{displaymath}
Fixing the second argument $y$ to $\sh_G$, we obtain a map
$\pathsp{\blank}^G(\sh_G):\BG\to\Set$. And since
$\pathsp{\sh_G}^G(\sh_G)\jdeq \USymG$, we can point $\pathsp{\blank}^G(\sh_G)$
with reflexivity to make it a pointed function $\BG\ptdto \BSG_{\USymG}$, \ie it
classifies a homomorphism from $G$ to the permutation group
$\SG_{\USymG}\jdeq \Aut_{\Set}(\USymG)$, denoted by\footnote{The letter $\rho$
  commemorates the word ``regular''}
\begin{displaymath}
  \rho_G:\Hom(G,\SG_{\USymG}).
\end{displaymath}

\begin{theorem}[Cayley]
  \label{lem:allgpsarepermutationgps}
  For any group $G$, $\rho_G$ is a monomorphism.\footnote{By
    \cref{def:typeofmono}, $\rho_G$ is a monomorphism means that the induced map
    $\USym\rho_G$ from the symmetries of $\sh_G$ in $\BG_\div$ to the symmetries
    of $\USymG$ in $\Set$ is an injection, \ie ``any symmetry is a symmetry in
    $\Set$''.}
\end{theorem}

\begin{proof}
  In view of \cref{def:typeofmono} we need to show that
  $\B\rho_G\jdeq\pathsp{\blank}^G(\sh_G) :\BG \to \BSG_{\USymG}$ is a \covering.
  Note first that $\pathsp{\blank}^G(\sh_G)$ can be factorized, as a pointed
  map, as:\footnote{All the maps involved are pointed maps between groupoids, so
    the commutativity of this triangle is a proposition.}
  \[
    \begin{tikzcd}[ampersand replacement=\&]
      \BG \ar[r,equivr,"{\pathsp{\blank}^G}"]\ar[dr,"\B\rho_G"']
      \& (\typetorsor_G,\princ G) \ar[d,"{\ev_{\sh_G}}"]%\jdeq\bigl((\BG\to\Set)_{(\princ G)},\princ G\bigr)
      \\
      \& \BSG_{\USymG}%\jdeq(\Set_{(\USymG)},\USymG)
    \end{tikzcd}
  \]
  Here, $\ev_{\sh_G}: \typetorsor_G\to \conncomp{\Set}{\USymG}$ is the
  evaluation map defined by $\ev_{\sh_G}(X)\defeq X(\sh_G)$ and pointed by
  reflexivity.

  Since $\pathsp{\blank}^G$ is an equivalence, we are done if we show that, for
  any $S:\Set_{(\USymG)}$, the fiber $\inv{\ev_{\sh_G}}(S)$ is a set. This fiber
  is by definition $\sum_{X:\typetorsor_G}(S\eqto X(\sh_G))$, which is a subtype
  of $\sum_{X:\BG\to\Set}(S\eqto X(\sh_G))$. The latter is the type of pointed
  maps from $\BG$ to $(\Set,S)$ and hence a set by \cref{lem:hom-is-set} (more
  speicifcally \cref{ft:ptd-decr-h-lev}, since $\BG$ is connected and $\Set$ is a
  groupoid). Therefore the fiber $\inv{\ev_{\sh_G}}(S)$ a set.
\end{proof}
\wip{%
  Isn't it a bit overkill to use the equivalence with torsors? 
  I (MB too) feel the following proof is much simpler:
  \begin{proof}
    We need to show that $\B\rho_G :\BG \to \BSG_{\USymG}$ is a \covering. For
    any set $X:\BSG_{\USymG}$, the fiber of $\B\rho_G$ at $X$ is
    \begin{displaymath}
      \inv {\B\rho_G} (X) \jdeq \sum_{z : \BG}X \eqto (z\eqto\sh_G).
    \end{displaymath}
    Take two elements $(z_1,\phi_1)$ and $(z_2,\phi_2)$ of this fiber. Then an
    identification of type $(z_1,\phi_1) \eqto (z_2,\phi_2)$ is given by
    a path $p: z_1 \eqto z_2$ such that the following triangle commutes:
    \begin{displaymath}
      \begin{tikzcd}[ampersand replacement=\&]
        X \rar["\phi_1"] \drar["\phi_2"swap]
        \& (z_1 \eqto \sh_G) \dar["\blank\cdot\inv p"]
        \\
        \& (z_2 \eqto \sh_G)
      \end{tikzcd}
    \end{displaymath}
    In proving the proposition that such a $p$ is unique, we can use the
    connectedness of $\BG$ and suppose that $z_1,z_2\jdeq\sh_G$, 
    in which case we can see that $p$ is fully determined by the other data: 
    $X$ is non-empty and $\phi_1(x)=\phi_2(x)\cdot p$ for any $x:X$.
  \end{proof}%
}


Note that the above theorem yields that $(G,\rho_G)$ is a monomorphism into
$\SG_{\USymG}$. In other words, $G$ is a subgroup of $\SG_{\USymG}$. 
Note that the induced map on symmetries is:
\begin{displaymath}
  \USym\rho_G:\USym G \to (\USym G \eqto \USym G): 
  g \mapsto (x \mapsto x\inv g)
\end{displaymath}

\wip{The reader may ask for a more informative underlying group of this
subgroup than $G$ itself. To this end we apply $E$ from \cref{lem:SubG=MonoG}
to $(G,\rho_G)$ to obtain a subgroup in the sense of 
\cref{def:set-of-subgroups}. The underlying group of $E(G,\rho_G)$
has classifying type $\sum_{X:\BSG_{\USymG}} \inv{\B\rho_G} (X)$
and shape $(\USymG,(\sh_G,\id_{\USymG}))$.
As illustrated in the following diagram, the symmetries of this shape
are precisely those permutations $\pi$ of $\USymG$ for with there exist
a $g:\USymG$ such that $\pi$ is right multiplication by $\inv g$.\footnote{%
The other traditional way to understand Cayley's Theorem is
by left multiplication by $g$.}  

\[\begin{tikzcd}[ampersand replacement=\&, column sep={1cm, between origins}]
\USymG\arrow[d,equivl,"\pi"']\arrow[rr,equivr,"\id"]
 \&\&(\sh_G\arrow[d,dotted,"g","="']
 \&=\arrow[d,equivr,"\_\cdot\inv g"]
 \&\sh_G)\\
\USymG\arrow[rr,equivl,"\id"']\&\&(\sh_G\&=\&\sh_G)
\end{tikzcd}\]
}

%% --- This exercise is just plain false --- 
% \begin{xca}\label{xca:evP_isPrG}
% Show that $\princ G$ and ${\ev_{\sh_G}}\circ{\pathsp{\blank}}$ are
% equal as pointed maps.
% \end{xca}

\begin{remark}\label{rem:CayleyOversize}
  In many cases, the set $\USymG$ used in \cref{lem:allgpsarepermutationgps} is
  larger than necessary for obtaining the symmetries in $G$ as symmetries of a
  set. A case in point is the group $\SG_3$, where the symmetries \emph{are}
  already symmetries of a set, namely of the set $\bn3$. However,
  $\USG_3\jdeq(\bn3\eqto\bn3)$ is a $6$-element set.  Let's take a closer look
  at where and how this happens in the proof.

  As we noticed in the proof of \cref{lem:allgpsarepermutationgps}, the map
  $\pathsp{\blank}^G(\sh_G): \BG\ptdto\conncomp{\Set}{\USymG}$ classifying the
  monomorphism $\rho_G$ is decomposed as an equivalence $\pathsp{\blank}^G$
  (pointed by reflexivity) followed by the evaluation map $\ev_{\sh_G}$ (also
  pointed by reflexivity). This is depicted in the following diagram, where the
  second line shows the induced maps on the symmetries.
   \[
     \begin{tikzcd}
       \BG \ar[rr,equivr,"\pathsp{\blank}"] && (\typetorsor_G,\princ G)
       \ar[rr,"\ev_{\sh_G}"]&&
       \conncomp{\Set}{\USymG}\\
       \USymG \ar[rr,equivr,"\pathsp{\blank}"] && %\loops?
       (\princ G \eqto \princ G) \ar[rr,"\ev_{\sh_G}"]&& %\loops?
       (\USymG \eqto \USymG)
     \end{tikzcd}
   \]
   Let $\PP$ be the $G$-set given by
   $\PP(z)\defeq(\princ G(z) \eqto \princ G(z))$ for all $z:\BG$. By function
   extensionality, $\princ G \eqto \princ G$ is equivalent to
   $\prod_{z:\BG}\PP(z)$, the type of invariant maps of $\PP$. By
   \cref{lem:fixpts-are-fixed}\ref{it:ev-is-inj}, an invariant map $h$, and
   hence the corresponding symmetries of $\princ G$, is uniquely determined by
   its value $h_{\sh_G}$ at $\sh_G$.\footnote{%
     This is an alternative way to understand that $\ev_{\sh_G}$, and hence
     $\pathsp\blank^G$, classifies a monomorphism.%
   } %
   The underlying set of $\PP$ being $\PP(\sh_G)\jdeq(\USymG \eqto \USymG)$,
   \cref{lem:fixpts-are-fixed}\ref{it:ev-is-eq-on-inv} characterizes exactly the
   invariant maps of $\PP$ as corresponding via $\ev_{\sh_G}$ with elements of
   $\USymG \eqto \USymG$ fixed by the action of $\PP$. How does $\PP$ act on
   $\USymG$? That's for you to determine in \cref{xca:PP-fixed-permutations}. In
   other words, $\ev_{\sh_G}$ forgets about the extra structure of $\PP$ and
   sends invariant maps of $\PP$ to fixed permutations of $\USymG$. For example,
   in the case of $\SG_3$, the fixed permutations are the 6 elements of
   $\USG_3\eqto\USG_3$ given by $\blank\circ\inv\sigma$ for each of the 6
   possible $\sigma:\bn3 \eqto \bn3$.
\end{remark}

\begin{xca}\label{xca:PP-fixed-permutations}
  Let conditions be as in \cref{rem:CayleyOversize}.  By analyzing transport in
  the type family $\princ G$, show that a permutation $\pi$ of $\USymG$
  is fixed if and only if $\pi(gg')=g\pi(g')$ for all $g,g':\USymG$.  Show that
  the fixed permutations of $\USymG$ form an abstract group and that evaluation
  of such a permutation at $\refl{\sh_G}$ yields an abstract isomorphism from
  this group to $\abstr(G)$.
\end{xca}

\wip{%
  Double checking the exercise's solution:
  \begin{proof}
    Note that a quick induction on $p: z\eqto z'$ in $\BG$ shows that, for all
    $q: (\sh_G\eqto z') \eqto (\sh_G\eqto z')$, we have
    $\PP(p)(q) = p \cdot q(\inv p \cdot \blank)$. So the action of $g:\USymG$ on
    a symmetry $\varphi:\USymG \eqto\USymG$ is
    $g\cdot\pi = g\pi(\inv g \cdot \blank)$. When such a $\pi$ is fixed by the
    action, then $g\pi(\inv g \blank) = \pi$ for all $g$. That is
    $\forall g,g', \pi(g g') = g\pi(g')$. In other words,
    $\pi = \blank \pi(\refl{\sh_G})$. Since
    $\USym (\ev_G) (h) = x \mapsto x\cdot
    h_{\sh_G}(\refl{\sh_G})$. When $h \jdeq \pathsp g^G$ for some $g:\USymG$, we
    get $h_{\sh_G}(\refl{\sh_G}) = \refl{\sh_G} \inv g$, so the selected
    permutation is indeed the ``left action by multiplication on the right''
    $x \mapsto x \inv g$.
  \end{proof}%
}

\wip{Should we add a little remark on the other natural $\B\rho_G$ possible,
  namely $\princ G$? (It is equal to the current one if we point it with
  $\inv\blank$.)}

\section{The lemma that is not Burnside's}
\label{sec:burnsides-lemma}
\label{lem:burnsides-lemma}

\begin{example}\label{exa:prep-burnside}
Since the lemma to come is about counting orbits and
elements of orbits, we start by elaborating an example.
Recall from \cref{ex:cyclicgroups} the cyclic group 
$\CG_4 \jdeq\Aut_{\Cyc}(\bn4,s)$, where $\Cyc$ is defined
in \cref{def:Cyc} as the type of cycles, \ie pairs $(X,t)$ of 
a set $X$ and a permutation $t:X\equivto X$ such that any two
points of $X$ are some $t$-steps apart.
Let $X: \BCG_4\to\Set$ be the $\CG_4$-set mapping any $(A,f):\BCG_4$
to $A\to\bn2$. Then the underlying set of $X$ is $\bn4\to\bn2$,
\ie binary sequences of length $4$. The group action induced by $X$
cyclically rotates such sequences, by $0,1,2$ or $3$ positions.\footnote{%
Use \cref{cor:id-m-cycle}, univalence, and \cref{lem:trp-in-function-type}.}

By \cref{cor:orbit-equiv}, the set of orbits $X/\CG_4$ is
equivalent to the quotient of $\bn4\to\bn2$ induced by 
$[\blank] : (\bn4\to\bn2) \to X/\CG_4$. % from \cref{lem:X/G=setTruncX_hG}.
As also stated by that lemma, the equivalence class of any $x:\bn4\to\bn2$
consists precisely of all cyclic rotations of $x$. Clearly, 
$0000$ and $1111$ have singleton equivalence classes.
The equivalence class of $0001$ (resp.\ $0111$) consists of all four binary 
sequences with exactly one $1$ (resp.\ $0$).  Before you start thinking that
swapping $0$'s and $1$'s gives a new equivalence class, consider 
$0101$ that forms an equivalence class together with $1010$.
Finally, $0011$ forms an equivalence class together with $1001$,
$1100$ and $0110$. Thus we have distributed all 16 sequences over
six orbits, as in the left column of \cref{fig:C4-action-on-4-bits}.

\begin{margintable}
  \footnotesize
\begin{tabular}{ccc} \toprule
 orbit & stabilizers \\ \midrule
 $0000$ & $0,1,2,3$ \\
 $1111$ & $0,1,2,3$ \\
 $0001,0010,0100,1000$ & $0$ \\
 $0111,1011,1101,1110$ & $0$ \\
 $0101,1010$ & $0,2$ \\
 $1100,0110,0011,1001$ & $0$ \\ \bottomrule
\end{tabular}
\caption{\label{fig:C4-action-on-4-bits}
Underlying sets of orbits and the stabilizers of their elements.}
\end{margintable}

In the right column of \cref{fig:C4-action-on-4-bits}, we have
given in each row the respective stabilizing symmetries in $\BCG_4$.
\cref{xca:[x]=[y]-implies-||Gx=Gy||} tells us that it doesn't matter
too much which element in the orbit one chooses.\footnote{%
  Here it matters even less since $\CG_4$ is abelian.}
For the cardinality $\Card((\CG_4)_x)$ of the finite stabilizer groups, 
the particular $x$ one chooses within in each orbit is irrelevant,
but may vary from orbit to orbit. Now we can observe something
interesting: the product $\Card({\CG_4}\cdot {x}) \times \Card((\CG_4)_x)$
(\ie in each row, the number of elements on the left times that on the right)
is equal to $\Card(\CG_4) = 4$, for each $x$ in the underlying set of $X$.
This follows from Lagrange's Theorem, in particular \cref{xca:lagrange},
applied with $G\jdeq\CG_4$ and taking for $X$ the underlying $\CG_4$-set
of $[x]$, which is transitive.

Another observation in \cref{fig:C4-action-on-4-bits} is that, 
since there are six orbits and the orbits induce a disjoint partition
of $\bn4\to\bn2$, there are in total 24 pairs $(g,x)$ with $g\cdot x = x$.
This insight leads to the following lemma.
\end{example}


\begin{lemma}
  \label{lem:burnside}
  Let $G$ be a finite group and let $X:\BG\to\Set$ be a finite $G$-set.
  For any $g:\USymG$, define the set 
  $X^g \defeq \setof{x:X(\sh_G)}{g\cdot x = x}$
  of points fixed by $g$.
  Then each $X^g$, the sum type $\sum_{g:\USymG} X^g$, 
  and the set of orbits $X/G$ are finite sets, and we have
  \begin{align}\label{eq:burnside}
% \[
    \Card\Bigl(\sum_{g:\USymG} X^g\Bigr) = \Card(X/G) \times \Card(G).
% \]
  \end{align}
\end{lemma}
\begin{proof}
  We first need to make sure that the sets involved are finite. 
  Finite sets are decidable sets, see \cref{xca:finsets-decidable}.
  Hence each $X^g$ is a finite set, as it is a decidable subset of $X(\sh_G)$,
  see \cref{rem:subset-of-fin-set}.
 
  Finiteness of $\sum_{g:\USymG} X^g$ follows from 
  \cref{xca:fin-sum-of-finsets}. Regarding the set of orbits,
  note that \cref{cor:orbit-equiv} yield that $X/G$ is equivalent
  to the quotient of $X(\sh_G)$ modulo the equivalence relation
  $\exists_{g:\USymG} x=g\cdot y$. The latter proposition is decidable by
  \cref{xca:dec-quant-finset}. Now apply \cref{xca:dec-quot-finite-set}.
  
  Since the main statement \cref{eq:burnside} of the lemma is a proposition, 
  we may assume that, for both $X(\sh_G)$ and $\USymG$, 
  we have an equivalence to a standard finite set.  
  Rearranging sums and writing $X(\sh_G)$ as the sum of fibers
  of $[\blank]: X(\sh_G)\to X/G$ gives equivalences:
  \begin{align*}
 &\sum_{g:\USymG} X^g \jdeq
 \sum_{g:\USymG}\sum_{x:X(\sh_G)}(g\cdot x = x)\equivto 
 \sum_{x:X(\sh_G)}\sum_{g:\USymG}(g\cdot x = x) \equivto \\
 &\sum_{x:X(\sh_G)} \USymG_x\equivto 
 \sum_{O:X/G} \sum_{x : X(\sh_G)} \bigl((O=[x])\times\USymG_x\bigr)\equivto 
 \sum_{O:X/G} \sum_{x : X_O(\sh_G)} \USymG_x
  \end{align*}
  In the last step we have used that $O=[x]$ is equivalent to
  $O(\sh_G,x)$, which means that $x$ is in the underlying set
  $X_O(\sh_G)$ of the orbit $O$, see \cref{def:actiontype}
  and \cref{def:Gsubset}.
  
  Note that the last type in the chain above reflects how
  we counted in \cref{fig:C4-action-on-4-bits}: for every orbit,
  and every element in the underlying set of that orbit,
  we counted the stabilizers of that element.
  
  We aim to apply the Lagrange construction with subgroups 
  defined by $X_O$ and
  $x_O : X_O(\sh_G)$, for any orbit $O:X/G$. These points $x_O$ can
  be obtained as the `least' $x:X(\sh_G)$ such that $O=[x]$,
  where `least' means: corresponding to the smallest number under the
  equivalence of $X(\sh_G)$ with a standard finite set.
  We also have to give functions
  $f_O : \prod_{y:X_O(\sh_G)}\sum_{g:\USymG} g\cdot_{X_O} y = x_O$,
  for every $O:X/G$.
  Such functions are obtained by using the transitivity of $X_O$
  in combination with the equivalence between $\USymG$ and a 
  standard finite set: we can simply take the `least' $g:\USymG$
  such that $g\cdot_{X_O} y = x_O$.
  Applying \cref{cor:lagrange-dep-sum},
  we get an equivalence between $\USymG$ and $\sum_{x : X_O(\sh_G)} \USymG_x$. 
  We conclude that $\Card(\sum_{g:\USymG} X^g) = \Card(X/G) \times \Card(G)$,
  using \cref{xca:fin-sum-of-finsets}.
\end{proof}

As a first application of Burnside's Lemma, we note the following
number-theoretic consequence, which falls out when we consider
the analog of~\cref{exa:prep-burnside} for the case of $\CG_p$ acting
on base-$n$ sequences of length $p$.

\begin{theorem}[Fermat's Little Theorem]
  For any prime $p$ and natural number $n$, we have $p \mid {n^p-n}$.
\end{theorem}
\begin{proof}
  Consider the $\CG_p$-set $X : \BCG_p \to \Set$ defined by
  \[
    X(S,t) \defeq (S \to \bn n),
  \]
  for any $p$-cycle $(S,t)$. The underlying set is the type of functions
  $\bn p \to \bn n$, which is finite of cardinality $n^p$.


  For any $f : \bn p \to \bn n$, we can apply \cref{xca:lagrange} to the
  stabilizer subgroup $(\CG_p)_f$. But since $p = \Card(\CG_p)$ is prime,
  $(\CG_p)_f$ can only be trivial or all of $\CG_p$. The latter happens when $f$
  is one of the $n$ constant functions. Otherwise,\footnote{The set
    $\bn p \to \bn n$ is finite, hence decidable, so there is no risk to make a
    case-by-base analysis.} select $k$ such that $f(k) \neq f(0)$, and surely
  $s^k \cdot f \neq f$: then $s^k$ is not in $\UCG_p$\footnote{More rigorously,
    $s^k$ is not in the image of $\USymi_f$ where $i_f: (\CG_p)_f \to \CG_p$ is the
    monomorphism defining the stablizer.} the only option is that $(\CG_p)_f$ is
  trivial. In other words, $X^g$ is all of $X(\bn p,s)$ for
  $g = \refl{(\bn p, s)}$, and only the constant functions for anyof the other
  $p-1$ symmetries in $\UCG_p$.
  
  Now apply Burnside's~\cref{lem:burnside}. We get:
  \[
    n^p + n(p-1) = \Card\Bigl(\sum_{g:\UCG_p} X^g\Bigr) = \Card(X/\CG_p) \times
    \Card(\CG_p) = \Card(X/\CG_p) \times p,
  \]
  and so we conclude that $p$ divides $n^p-n$.
\end{proof}



%%% Local Variables:
%%% mode: LaTeX
%%% fill-column: 144
%%% latex-block-names: ("lemma" "theorem" "remark" "definition" "corollary" "fact" "properties" "conjecture" "proof" "question" "proposition" "exercise")
%%% TeX-master: "book"
%%% TeX-command-extra-options: "-fmt=macros"
%%% compile-command: "make"
%%% End:
