\section{The Limited Principle of Omniscience}

\begin{remark}\label{rem:LPO-solves-halting problem}
Recall the Limited Principle of Omniscience (LPO):
  for any function $P:\NN\to\bn 2$, 
  either there is a smallest number $n_0:\NN$ such that $P(n_0)=1$,
  or $P$ is a constant function with value $0$.
We will show that LPO is not provable in our theory.\footnote{%
The statement that something is not provable is a statement \emph{about}, 
and not \emph{in}, a theory. Proving such a statement often requires
properties of the theory that cannot be formulated nor proved in the theory itself.

One such `metaproperty' that we use here is \emph{canonicity}.
We call an element \emph{closed} if it does not contain free variables.
An example of canonicity is that every closed natural number is a \emph{numeral},
that is, either $0$ or $S(n)$ for some numeral $n$.
Another example of canonicity is that every closed element of $L\coprod R$
is either of the form $\inl{l}$ for some $l:L$ or 
of the form $\inr{r}$ for some $r:R$.
A second important metaproperty of our theory is that one can compute canonical forms.
}%end footnote

The argument is based on the Halting problem: given a Turing machine 
$M$ and an input $n$, determine whether $M$ halts on $n$.
It is known that the Halting problem cannot be solved by an algorithm.
%that can be implemented on a Turing machine.

We use a few more facts from computability theory.
First, Turing machines can be enumerated $M_0,M_1,\ldots$
Second, there exists a function $T(e,n,k)$ such that $T(e,n,k) = 1$
if $M_e$ halts on input $n$ in at most $k$ steps, and $T(e,n,k) = 0$
otherwise. This function $T$ can easily be implemented in our theory.

Towards a contradiction, assume we have a closed proof $t$ of LPO in our theory.
In particular, $t$ does not depend on any axiom.\footnote{It is possible to weaken the notion
of canonicity so that the argument still works even if the proof $t$ uses $\ua$.}
It is clear that $k\mapsto T(e,n,k)$ is a constant function with value $0$
if and only if $M_e$ does not halt on input $n$. Now consider $t(k\mapsto T(e,n,k))$,
which is an element of a type of the form $L\coprod R$.

We now explain how to solve the Halting problem.
Let $e$ and $n$ be arbitrary numerals. 
Then $t(k\mapsto T(e,n,k))$ is a closed element of $L\coprod R$.
Hence we can compute its canonical form. If $t(k\mapsto T(e,n,k))\jdeq\inr{r}$ for some
$r:R$, then $k\mapsto T(e,n,k)$ is a constant function with value $0$,
and $M_e$ does not halt on input $n$. If $t(k\mapsto T(e,n,k))\jdeq\inl{l}$ for some
$l:L$, then $M_e$ does halt on input $n$.
Thus we have an algorithm to solve the Halting problem
for all $e$ and $n$. Since this is impossible, we have refuted the assumption
that there is a closed proof $t$ of LPO in our theory.
\end{remark}
