\chapter{A categorical interlude}
\label{ch:cats}
\marginnote{%
  This chapter introduces some useful terminology that we'll use in the rest of the book.
  It can probably be skipped at a first reading, and only consulted as needed.}

We have seen that many types carry a notion of morphism between its elements:
\begin{itemize}
\item We have functions $f : A \to B$ between types $A$ and $B$ in a universe $\UU$
  (\cref{univalent-mathematics}),
\item We have identifications $p : x \eqto y$ between elements $x,y$ of any type $A$
  (\cref{sec:identity-types}),
\item We have pointed functions $f : A \ptdto B$ between pointed types $A$ and $B$ in $\UUp$
  (\cref{def:pointedtypes}),
\item We have fiberwise maps $f : \prod_{x:X}(A(x) \to B(x))$ between families $A,B : X \to \UU$ (\cref{def:fiberwise}),
\item We have homomorphisms $f : \Hom(G, H)$ between groups $G, H : \Group$
  (\cref{def:grouphomomorphism}),
\wip{\item We have maps of cycles $f : (X,t) \to (Y,u)$ and of bicycles $(X,a,b) \to (X',a',b')$,}
\item We have maps of $G$-sets $f : \Hom_G(X,Y)$ for $X,Y : \BG \to \Set$
  (\cref{def:map-of-Gsets}; a special case of fiberwise maps).
\end{itemize}
In all those cases, we have notions of identity morphism and composition of morphisms.
We have also seen that some maps between types are paired with maps on morphisms, for example,
taking underlying symmetries in groups, $\USym : \Group \to \Set$ (\cref{def:group-symmetries}),
comes with a corresponding operation of taking the underlying map of symmetries of a group homomorphism,
\[ \USym : \Hom(G,H) \to (\USymG \to \USymH) \]
(\cref{def:USym-hom}) satisfying $\USym(\id_G) = \id_{\USymG}$ and $\USym(\psi\circ\varphi) = \USym\psi \circ \USym\varphi$ (\cref{cor:USym-compose}).

It's very useful to develop some abstractions for types equipped with a notion of morphism and maps equipped with maps of morphisms like this.
These give the notions of (wild) \emph{categories} and \emph{functors}, respectively,
and \emph{category theory} is the study of these structures.

Here we give a brief primer\footnote{%
  The topic is of course too vast to cover in detail here,
  so we refer to the literature for more details.
  Category theory in univalent foundations is also treated in
  Chapter~10 of the HoTT book\footnotemark{}
  (based on~\citeauthor{AKS2015}\footnotemark{}),
  while \citeauthor{AwodeyCat}\footnotemark{}
  and \citeauthor{RiehlContext}\footnotemark{}
  give traditional expositions,
  and \citeauthor{MacLaneWorking}\footnotemark{}
  gives a comprehensive treatment.}%
\addtocounter{footnote}{-3}\footcitetext{hottbook}%
\stepcounter{footnote}\footcitetext{AKS2015}%
\stepcounter{footnote}\footcitetext{AwodeyCat}%
\stepcounter{footnote}\footcitetext{RiehlContext}%
\stepcounter{footnote}\footcitetext{MacLaneWorking}%
on category in order to systematize what we've done so far,
and prepare for the main result of the next chapter,
\cref{ch:absgroup}, which is to give an
\emph{equivalence of categories} between the categories of concrete and abstract groups.

\section{Brief overview of the chapter}

In~\cref{sec:categories} we define the kinds of categories we need, along with many examples (including the above).
Then we discuss various abstract notions in categories
(terminal and initial objects, monomorphisms and epimorphisms)
and remark on the importance of \emph{duality} in~\cref{sec:duality}.
In~\cref{sec:naturality} we cover functors and natural transformations,
and in~\cref{sec:adjunctions} we treat adjunctions;
we have already seen some examples of adjunctions,
for example in~\cref{xca:adjunction-_!-^*,xca:adjunction-^*-_*}.

We have also seen an incarnation of the Yoneda lemma in~\cref{xca:TTYoneda};
in~\cref{sec:yoneda} we treat the O.G.\ version.
This is leveraged both in~\cref{sec:cat-equiv},
where we cover equivalences of categories
and the relationship with univalence,
and in~\cref{sec:limits}, where we treat general limits and colimits.

We end with a brief introduction to monoidal categories in~\cref{sec:monoidal-cats}, as we'll see in~\cref{ch:abelian} that abelian groups form an example.

\section{Categories}
\label{sec:categories}

As mentioned above, many types come equipped with notion
of \emph{morphism} or \emph{arrow} between its elements
which is more general than identification or \emph{isomorphism}.
For instance
we have type of functions $A \to B$ for $A,B:\UU$
and the type of pointed functions $A \ptdto B$ for $A,B:\UUp$.
There are identity morphisms and composition of morphisms,
and this motivates the following definition:

\begin{definition}\label{def:wild-cat}
  A \emph{wild precategory}\index{category!wild precategory}\footnote{%
    See below for remarks on the terminology.
    Adding further properties to the data given here eventually
    recovers the notion of a category \emph{simpliciter},
    see \cref{def:category}.}
  consists of the following data:
  \begin{enumerate}
  \item\label{struc:cat-ob} A type $\var{Ob}$, called the \emph{type of objects}.
  \item\label{struc:cat-mor} For each pair of objects $A,B : \var{Ob}$,
    a type of \emph{morphisms} $\hom(A,B)$. These are also known as \emph{arrows},
    and written $A \to B$ when there's no danger of confusion.
    If $f : A \to B$ is such an arrow, then we say that the \emph{domain} of $f$ is $A$
    and the \emph{codomain} of $f$ is $B$.
  \item\label{struc:cat-id} For each object $A : \var{Ob}$,
    an \emph{identity arrow} $\id_A : A \to A$.
  \item\label{struc:cat-comp} For each pair of arrows $f : A \to B$ and $g : B \to C$,
    a \emph{composite arrow} $g\circ f : A \to C$.\footnote{%
      To be fully explicit, the composition operation has
      type
      \[
        \prod_{A,B,C:\var{Ob}}(B \to C) \to (A \to B) \to (A \to C),
      \]
      and we might denote it $g \circ_{A,B,C} f$.
      Since the objects $A$, $B$, and $C$ can often be inferred,
      we leave them out, lest the notation becomes too heavy.
      A similar remark goes for the other operations.}
  \item\label{struc:cat-unit-laws} For each arrow $f : A \to B$,
    a pair of identifications
    \[
      \lambda : \id_B \circ f \eqto f, \quad
      \rho : f \circ \id_A \eqto f.
    \]
  \item\label{struct:cat-assoc} For each triple of arrows
    $f : A \to B$, $g : B \to C$, and $h : C \to D$, an identification
    \[
      \alpha : h \circ (g \circ f) \eqto (h \circ g) \circ f.
    \]
  \end{enumerate}
  If $\mathcal C \jdeq (\var{Ob},\hom,\id,\lambda,\rho,\alpha)$
  is a wild precategory, then we write $A, B : \mathcal C$ instead of $A, B : \var{Ob}$
  to indicate that $A, B$ are elements of the underlying type of objects of $\mathcal C$.
  We also write $\Ob(\mathcal C)$ for this type.
  We may write $f, g : A \to_{\mathcal C} B$ to emphasize where the arrows $f$ and $g$ live, if needed,
  and sometimes $\hom_{\mathcal C}(A,B)$ or $\mathcal C(A,B)$, instead of $\hom(A,B)$.
\end{definition}
\begin{remark}[On the adjective ``wild'']\label{rem:wild}
  With this definition, we readily equip the universes of types $\UU$ and of pointed types $\UUp$
  with a structure of wild precategories.
  In the former case, we can use reflexivities for $\lambda$, $\rho$, and $\alpha$.
  In the latter, we leave their definition as an exercise.

  We use the adjective ``wild'' to highlight a deficiency of this definition as it stands:
  We haven't specified any further laws for the identifications $\lambda$, $\rho$, and $\alpha$.
  For example, it would be sensible to require an identification of $\lambda$ and $\rho$ at an identity:
  $\id_A \circ \id_A \eqto \id_A$,
  as well as a filler for the pentagonal diagram of $\alpha$'s coming from four composable arrows:
  \begin{equation}\label{eq:pentagon}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (P0) at (90:2.3cm) {$k \circ (h \circ (g \circ f))$};
      \node (P1) at (90+72:2cm) {$k \circ ((h \circ g) \circ f)$};
      \node (P2) at (90+2*72:2cm) {\makebox[5ex][r]{$(k \circ (h \circ g)) \circ f$}};
      \node (P3) at (90+3*72:2cm) {\makebox[5ex][l]{$((k \circ h) \circ g) \circ f$}};
      \node (P4) at (90+4*72:2cm) {$(k \circ h) \circ (g \circ f)$};
      \path[commutative diagrams/.cd, every arrow, every label]
      (P0) edge[eql] node[swap] {$\ap{k\circ\blank}(\alpha)$} (P1)
      (P1) edge[eql] node[swap] {$\alpha$} (P2)
      (P2) edge[eql] node[swap] {$\ap{\blank\circ f}(\alpha)$} (P3)
      (P4) edge[eqr] node {$\alpha$} (P3)
      (P0) edge[eqr] node {$\alpha$} (P4);
    \end{tikzpicture}
  \end{equation}
  Of course, we would on occasion then need a filler for a three dimensional diagram of pentagons
  for five composable arrows, etc., etc., \emph{ad infinitum}.\footnote{%
    There is a hierarchy of notions, $A_n$-precategories, for $n\ge 2$,
    with coherence conditions involving up to $n$ composable arrows.
    The wild precategories lie between $A_2$ and $A_3$ in this hierarchy.
    Besides the mentioned identification of $\lambda$ and $\rho$ for two identities,
    we'd also require fillers for diagrams like
    \[
      \begin{tikzcd}[ampersand replacement=\&,column sep=tiny]
        \& f\circ g \& \\
        f\circ(\id\circ g) \ar[rr,eql,"\alpha"']\ar[ur,eqr,"{\ap{f\circ\blank}(\lambda)}"]
        \& \& (f\circ\id)\circ g\ar[ul,eql,"{\ap{\blank\circ g}(\rho)}"']
      \end{tikzcd}
    \]
    as well as coherences when one or both of $f$ and $g$ are identities.}
  The corresponding structure is known to connoisseurs as an $(\infty,1)$-precategory.
  It is an open problem as we write this whether this notion can be defined in our type theory.
  Types and pointed types should furnish examples.
\end{remark}

However, when the types of morphisms $\hom(A,B)$ are \emph{sets}, then the types
of $\lambda$, $\rho$, and $\alpha$ are propositions, so any coherence conditions
are automatically fulfilled.
This motivates the following definition.

\begin{definition}\label{def:precategory}
  A \emph{precategory}\index{category!precategory}
  is a wild precategory $\mathcal C$ in which the types
  $A \to_{\mathcal C} B$ are sets, for all objects $A,B : \mathcal C$.
\end{definition}

Most (wild) precategories we shall meet satisfy a further condition
that makes them better behaved than arbitrary precategories:
a \emph{univalence} condition.
In fact, for the wild precategory of types and functions,
this condition is exactly the Univalence Axiom (\cref{def:univalence})!

In order to define this, we need the notion corresponding to equivalence
in an general wild precategory.

\begin{definition}\label{def:iso-in-cat}
  A morphism $f : A \to B$ in a wild precategory $\mathcal{C}$ is an
  \emph{isomorphism}\index{isomorphism!in a (wild pre-)category}
  if have $g, h : B \to A$ and identifications
  $\sigma : \id_B \eqto f\circ g$ and $\rho : \id_A \eqto h\circ f$.
  This condition is encoded by the type
  \[
    \isIso(f) \defeq \biggl(\sum_{g:B\to A}f\circ g\eqto\id_B\biggr)
    \times\biggl(\sum_{h:B\to A}h\circ f\eqto\id_A\biggr)
  \]
  If $f$ is an isomorphism, we also say that $f$ is \emph{invertible}.

  We define the type $A \isoto B$ of \emph{isomorphisms} from $A$ to $B$
  in $\mathcal C$ ($A \isoto_{\mathcal C} B$ if needed)
  by the following definition.%
  \glossary(2isoto){$\protect\isoto$}{type of isomorphisms}
  \[
    (A \isoto B) \defeq \sum_{f:A\to B} \isIso(f).\qedhere
  \]
\end{definition}
The type $\isIso(f)$
is indeed a proposition,\footnote{%
  This follows just as for functions: If $f$ is an isomorphism,
  then each factor in the product is contractible, as, e.g.,
  $\sum_{g:B\to A}f\circ g\eqto\id_B$ is the fiber
  of $f\circ\blank : (B \to A) \to (B \to B)$
  at $\id_B$, and all functions $f\circ\blank$ and $\blank\circ f$
  are equivalences of types using the data that makes $f$ an isomorphism
  along with $\lambda$, $\rho$, and $\alpha$.}
and every identity arrow is an isomorphism.
We write $\inv f$ for the inverse of an isomorphism $f$.
\begin{definition}\label{def:univalent-cat}
  A wild precategory $\mathcal C$ is \emph{univalent}\index{univalent} if
  for all objects $A,B : \Ob(\mathcal C)$, the function
  \[
    \constant{idtoiso}_{A,B} : (A \eqto_{\Ob(\mathcal C)} B) \to (A \isoto_{\mathcal C} B)
  \]
  defined by path induction sending $\refl A$ to $\id_A$,
  is an equivalence.
\end{definition}
\begin{definition}\label{def:category}
  A \emph{wild category}\index{category!wild category} is a univalent wild precategory,
  and a \emph{category}\index{category} is a univalent precategory.
\end{definition}
Now we are ready to restate the examples mentioned in the introduction to the chapter.
In each case we leave it to the reader to supply most of the data.
\begin{itemize}
\item
  The \emph{wild category of types} (in universe $\UU$) has $\UU$
  as its type of objects, and the function type $A \to B$ as its type of arrows
  for $A,B :\UU$. It is univalent by the Univalence Axiom.
\item
  The \emph{wild category of pointed types} (in a universe $\UU$) has
  $\UUp$ as its type of objects, and the type of pointed maps $A \ptdto B$ as its type of arrows.
\item
  The \emph{wild category of families} over a given type $X$ has
  $X \to \UU$ as its type of objects, and the type of fiberwise maps
  $(X \to_B Y) \defeq \prod_{b:B}X(b) \to Y(b)$
  as its type of arrows.
\item
  The \emph{category of sets}\index{category!of sets} has $\Set$
  as its type of objects, and the function type $A \to B$ as its type of arrows.
  It's a category since each type $A\to B$ is a set, and it's univalent.
\item
  The \emph{category of groups}\index{category!of groups} has $\Group$
  as its type of objects, and the homomorphism type $(G \to H) \defeq \Hom(G,H)$
  as its type of arrows.
\item
  The \emph{category of $G$-sets}\index{category!of $G$-sets}, for a group $G$,
  has $\GSet$ as its type of objects, and the type of morphisms of $G$-sets
  $\Hom_G(X,Y)$ (\cref{def:map-of-Gsets}) as its type of arrows.
\end{itemize}
Note that the category of sets sits inside the wild category of types.
This situation is quite typical is a rich source of further examples,
captured by the following definition.
\begin{definition}\label{def:full-subcat}
  If $\mathcal C$ is a wild precategory, and $P : \Ob(\mathcal C) \to \Prop$
  is a predicate on the objects, then we get a new wild precategory
  $\mathcal C_P$ with $\Ob(\mathcal C_P) \defeq \Ob(\mathcal C)_P$,
  using the notation of~\cref{def:subtype}, inheriting all the rest of the
  structure of $\mathcal C$. It is non-wild/univalent, if $\mathcal C$ is,
  and is called the \emph{full (wide) sub(pre)category}
  of $\mathcal C$ on the predicate $P$.\index{category!full subcategory}
\end{definition}

By the univalence condition, in a category each identity type $A \eqto B$
is equivalent to the set $A \isoto B$, and is hence a set.
Thus we get the following.
\begin{lemma}\label{lem:obj-gpd}
  The type of objects of a category form a groupoid.
\end{lemma}

And important special case is when each arrow type $A \to B$ is a \emph{proposition}.
\begin{example}\label{ex:poset}
  A \emph{preorder}\index{preorder} is precategory in which every arrow type is a proposition.
  In this case, the types of $\lambda$, $\rho$, and $\alpha$ are contractible, so the data
  of a preorder reduces to just a type $P$ and a binary relation,
  typically written $\le : P \to P \to \Prop$,
  that is reflexive, $x \le x$ (via the identities)
  and transitive, \ie if $x\le y$ and $y\le z$ implies $x\le z$ (via composition).

  A \emph{partial order}\index{partial order}, also known as a \emph{poset}\index{poset},
  is a univalent preorder.
  In this case, the type of objects is a set.
  This happens if and only if the relation is symmetric,
  \ie if $x \le y$ and $y \le x$ implies $x = y$.

  Typical examples are $(\NN,\le)$, $(\ZZ,\le)$, $(\Prop,\to)$,
  and $(\Sub(S),\subseteq)$ for a set $S$.
  A preorder that fails to be a poset is the two-element type $\bn 2$
  with the always true relation.
  This is hence also an example of a precategory that fails to be univalent.
\end{example}

\begin{xca}\label{xca:order-poset}
  Show that the type of orders, $\Order$,
  with the divisibility relation from~\cref{def:Order} forms a poset.
\end{xca}

Another important special case is when every morphism is an isomorphism.
\begin{definition}\label{def:wild-pre-groupoid}
  A (\emph{wild}) \emph{pregroupoid} is a (wild) precategory in which every
  arrow is invertible.
  A (\emph{wild}) \emph{groupoid} is a univalent (wild) pregroupoid.
\end{definition}

\begin{example}\label{ex:path-groupoid}
  Every type $X$ gives rise to a wild groupoid, its (wild) \emph{path groupoid},
  having $X$ as its type of objects and $x \eqto y$ as its type of arrows.
  The arrows are invertible, since paths are always invertible.

  If $X$ is a $1$-type, then this structure is a groupoid.
\end{example}
There's no conflict with the terminology introduced in~\cref{sec:props-sets-grpds},
because this construction gives an equivalence
from the type of $1$-types (in $\UU$) to the type of groupoids (in $\UU$),
as we shall see below in~\cref{thm:1types-are-groupoids}.

\Cref{def:wild-cat,def:precategory,def:category,def:wild-pre-groupoid}
are summarized as a diagram of inclusions in~\cref{fig:cat-summary}.

\begin{figure}
\begin{sidecaption}%
  {The various notions of categories arranged in a cube: on the bottom face
    we loose the adjective ``wild'' by requiring arrows to form \emph{sets};
    on the front left face we loose the prefix ``pre'' by requiring
    \emph{univalence}; and on the front right face we restrict to groupoids by
    requiring all arrows be \emph{invertible}.}[fig:cat-summary]
  \centering
  \footnotesize
  \begin{tikzpicture}
    \node (wpc) at (0,3) {wild precategory};
    \node (wc) at (-3,1) {wild category};
    \node (wpg) at (3,1) {wild pregroupoid};
    \node (wg) at (0,-1) {wild groupoid};
    \node (pc) at (0,1) {precategory};
    \node (c) at (-3,-1) {category};
    \node (pg) at (3,-1) {pregroupoid};
    \node (g) at (0,-3) {groupoid};
    \begin{scope}[->,commutative diagrams/hook]
      \draw (g) -- (pg);
      \draw (g) -- (c);
      \draw (pg) -- (pc);
      \draw (c) -- (pc);
      \draw (pc) -- (wpc);
      \draw (wc) -- (wpc);
      \draw (wpg) -- (wpc);
      \draw[commutative diagrams/crossing over] (wg) -- (wc);
      \draw[commutative diagrams/crossing over] (wg) -- (wpg);
      \draw (c) -- (wc);
      \draw (g) -- (wg);
      \draw (pg) -- (wpg);
    \end{scope}
    \node at (-4,-.5) [casred]{hom-\emph{sets}};
    \node at (-2,-2.5) [casred]{%
      \begin{tabular}{c}only \\ \emph{iso}morphisms\end{tabular}};
    \node at (2,-2.5) [casred]{\emph{univalence}};
  \end{tikzpicture}
\end{sidecaption}
\end{figure}

\begin{remark}\label{rem:cat-sizes}
  This is a good moment to remark on the size issues
  we have so far swept under the rug.
  The definition of a wild precategory $\mathcal C$ can be parametrized by
  \emph{two} universe levels: The type of objects belong to one, $\var{Ob}:\UU'$,
  while the family of arrows belong to another, $\hom : \var{Ob} \to \var{Ob} \to \UU$.

  If they coincide, then we call $\mathcal C$ a \emph{$\UU$-small category}\index{category!small}.
  For instance, a path groupoid for a type $X:\UU$ is $\UU$-small.

  The other common case is where $\UU:\UU'$, in which case
  we call $\mathcal C$ \emph{locally $\UU$-small}\index{category!locally small}.
  For example, the (wild) categories of types, sets, pointed types, groups, etc., built
  from types in $\UU$ are all locally $\UU$-small.
  This generalizes~\cref{def:ess-loc-small} for the case of path groupoids.
\end{remark}

Many notions in category theory work already at the level of wild categories,\footnote{%
  They \emph{all} work at the level of $(\infty,1)$-categories.
  We refer to \citeauthor{LurieHTT}\footnotemark{} and
  \citeauthor{LandInftyCat}\footnotemark{} for details.}%
\addtocounter{footnote}{-1}\footcitetext{LurieHTT}%
\stepcounter{footnote}\footcitetext{LandInftyCat}
but a notable exception is the construction of slice categories,
(also known as \emph{over categories}).
\begin{example}\label{def:slice-cat}
  The \emph{slice precategory} of a precategory $\mathcal C$ over an object $C : \mathcal{C}$,
  denoted $\mathcal C/C$,
  has as objects the type $\sum_{A:\mathcal C}A \to C$ of pairs $(A,f)$ of an object $A$ and an arrow $f:A\to C$
  with codomain $C$.
  An arrow from $(A,f)$ to $(A',f')$ is an arrow $g : A \to A'$ such that $f' \circ g=f$.\footnote{%
    This is a proposition since $\mathcal C$ is a precategory. We illustrate the arrow as
    a commuting triangle:
    \[
      \begin{tikzcd}[ampersand replacement=\&,column sep=small]
        A \ar[dr,"f"']\ar[rr,"g"] \& \& A' \ar[dl,"f'"] \\
        \& C \&
      \end{tikzcd}
    \]}
  The identities and compositions are inherited from $\mathcal C$.
  If $\mathcal C$ is univalent (and hence a category), then so is $\mathcal C/C$.

  If we try to define the slice $\mathcal C/C$ for an arbitrary wild precategory $\mathcal C$,
  using identifications $f' \circ g \eqto f$, we find that we need the pentagon coherence
  for $\alpha$ for $\mathcal C$ in order to define the $\alpha$ for $\mathcal C/C$.

  Of course, for \emph{particular} wild categories, it may very well happen
  that $\mathcal C/C$ is again a wild categories.\footnote{%
    This will be the case when $\mathcal C$ should be an $(\infty,1)$-category,
    carrying the whole hierarchy of coherences, which then
    carry over to $\mathcal C/C$.}
\end{example}
\begin{xca}\label{xca:univ-slice-cat}
  Construct a wild precategory structure on the \emph{slice of the universe} $\UU/B$
  over a fixed type $B:\UU$.
\end{xca}

\section{Abstract notions and duality}
\label{sec:duality}

Many concepts that we introduced in~\cref{ch:univalent-mathematics}
for the wild category of types make sense in arbitrary wild precategories.

\begin{definition}\label{def:terminal-obj}
  A \emph{terminal object}\index{terminal object}
  in a wild precategory $\mathcal C$ is an object $1$
  such that for any object $A:\mathcal C$,
  the arrow type $A \to 1$ is contractible.
\end{definition}
\begin{xca}\label{xca:terminal-prop}
  Show that if $\mathcal C$ is univalent, then
  the type of terminal objects is a proposition.
\end{xca}
The unit type $\bn 1$ is a terminal object in the wild category of types,
and in the category of sets,
while the trivial group $\TG$ is a terminal object in the category groups.

The is a ``dual'' notion as well.
\begin{definition}\label{def:initial-obj}
  An \emph{initial object}\index{initial object}
  in a wild precategory $\mathcal C$ is an object $0$
  such that for any object $A:\mathcal C$,
  the arrow type $0\to A$ is contractible.
\end{definition}
For example, the empty type $\bn 0$ is initial in the wild category of types,
while the trivial group $\TG$ is initial in the category of groups.

\begin{xca}\label{xca:unit-ptd-types}
  Show that the unit type $\bn 1$, pointed at its unique element,
  is both terminal and initial in the wild category of pointed types.
\end{xca}

The relationship between terminal and initial objects reflects
a deep aspect of category theory: Every concept comes with a dual version
obtained by ``reversing all the arrows''.
More formally, can introduce for every wild precategory its opposite category
that has its arrows reversed.
\begin{definition}\label{def:op-cat}
  For any wild precategory $\mathcal C\jdeq(\var{Ob},\hom,\id,\lambda,\rho,\alpha)$,
  define the \emph{opposite}\index{category!opposite} $\mathcal C\op$
  to have the same type of objects,
  morphisms $\hom_{\mathcal C\op}(A,B) \defeq \hom_{\mathcal C}(B,A)$,
  identities the same, and composition reversed:
  If $f : A \to_{\mathcal C} B$ and $g : B \to_{\mathcal C} C$,
  then $g\circ f : A \to_{\mathcal C} C$ works
  as the composite $f \circ_{\mathcal C\op} g$ of $g : C \to_{\mathcal C\op} B$
  and $f : B \to_{\mathcal C\op} A$.
  We then swap the roles of $\lambda$ and $\rho$,
  and $\inv\alpha$ plays the role of $\alpha$ in $\mathcal C\op$.
\end{definition}
\begin{lemma}\label{lem:op-idem}
  The operation of taking opposites defines an equivalence
  from the type of wild precategories to itself,
  with a trivially defined identification $(\mathcal C\op)\op \eqto C$.
\end{lemma}
It follows that any construction or theorem about wild precategories
has a dual version, obtained by precomposition with $(\blank)\op$.

For example, the dual of the slice category construction
is the \emph{coslice} category $C/\mathcal C$
(also known as the \emph{under category}).

As a further example of a pair dual notions, we consider that of
monomorphisms and epimorphisms.
\begin{definition}\label{def:mono-in-cat}
  An arrow $f : A \to B$ in a wild precategory $\mathcal C$
  is called a \emph{monomorphism}\footnote{For short: a mono} if
  post-composition with $f$ is an injection
  \[
    f\circ\blank : (C \to_{\mathcal C} A) \to (C \to_{\mathcal C} B)
  \]
  for all objects $C:\mathcal C$.
\end{definition}
\begin{definition}\label{def:epi-in-cat}
  Dually, $f$ is called an \emph{epimorphism}\footnote{For short: an epi}
  if pre-composition with $f$ is an injection
  \[
    \blank\circ f : (B \to_{\mathcal C} C) \to (A \to_{\mathcal C} C)
  \]
  for all objects $C:\mathcal C$.\footnote{%
    We can illustrate these in diagrams as saying that
    $f$ is a mono if a map into the codomain of $f$ factors in
    at most one way through the domain of $f$,
    \[
      \begin{tikzcd}[ampersand replacement=\&,column sep=small]
        \& A \ar[d,"f"] \\
        C \ar[ur,dashed]\ar[r] \& B
      \end{tikzcd}
    \]
    and dually, $f$ is an epi if a map out of the domain of $f$
    factors in at most one way through the codomain of $f$:
    \[
      \begin{tikzcd}[ampersand replacement=\&,column sep=small]
        A \ar[d,"f"']\ar[r] \& C \\
        B \ar[ur,dashed]
      \end{tikzcd}
    \]}
\end{definition}
If $C$ is a precategory, then these conditions reduce to the implications
\[
  f\circ g=f\circ h \to g=h, \quad\text{and}\quad g\circ f=h\circ f \to g=h,
\]
respectively.

We already met the monomorphisms in the category of groups in~\cref{def:typeofmono}
using a different definition.
\begin{xca}\label{xca:mono-cat-grp}
  Show that the monomorphisms in the category of groups are the same as
  those of~\cref{def:typeofmono}.
\end{xca}
\begin{xca}\label{xca:monos-epis-sets-types}
  Show that the monomorphisms in the wild category of types are just the injections,
  and the epimorphisms in the category of sets are just the surjections.
\end{xca}
The epimorphisms in the wild category of types are always surjections,
but are much more restricted. See~\citeauthor{BdJR2025}\footcite{BdJR2025} for details.
\begin{xca}\label{xca:monos-epis-preorder}
  Show that every morphism in a preorder is both a mono and an epi.
\end{xca}

\section{Functors and natural transformations}
\label{sec:naturality}

Not only do have arrows \emph{in} (wild pre-)categories, there's also
a notion of arrow \emph{between} them. These are called functors.
\begin{definition}\label{def:functor}
  A \emph{wild functor}\index{functor}
  $F : \mathcal C \to \mathcal D$
  between wild precategories $\mathcal C$ and $\mathcal D$
  consists of a function
  $F : \Ob(\mathcal C) \to \Ob(\mathcal D)$,
  mapping objects to objects,
  and a family of functions\footnote{%
    In practice, the functions on objects and arrows are named the same as
    the functor, but they could be disambiguated with subscripts,
    say, $F_0$ and $F_1$, if needed.}
  \[
    F : \prod_{A,B:\mathcal C}(A \to_{\mathcal C} B) \to (F(A) \to_{\mathcal D} F(B))
  \]
  together with identifications
  \[
    F_{\id} : F(\id_A) \eqto \id_A,\quad\text{and}\quad
    F_{\circ} : F(g\circ f) \eqto F(g)\circ F(f),
  \]
  for all objects $A$ and composable arrows $f$ and $g$ in $\mathcal C$.

  If $\mathcal D$ is a precategory, then the types of $F_{\id}$ and $F_{\circ}$
  are propositions, and in this case we just call $F$ a \emph{functor}.
\end{definition}
\begin{xca}\label{xca:wild-functor-isos}
  Show that every wild functor maps isomorphisms to isomorphisms.
\end{xca}
\begin{example}\label{ex:functor-preorders}
  A functor between preorders $(P,\le)$ and $(Q,\le)$ amounts to a monotone map
  $F : P\to Q$, \ie $p\le p'$ implies $F(p) \le F(p')$ for $p,p':P$.
\end{example}
\begin{example}\label{ex:USym-functor}
  Taking the underlying set of symmetries $\USym$
  gives a functor $\USym : \Group \to \Set$.
  It's easy to check that $\USym(\id_G) = \id_{\USymG}$,
  and we verified the preservation of composition
  in~\cref{cor:USym-compose}.
\end{example}
In~\cref{ch:absgroup} we shall define the category of abstract groups,
$\absGroup$,
in such a way that the map $\abstr : \Group \to \absGroup$
from~\cref{def:abstrG} becomes a functor
in line with the discussion from~\cref{rem:first-abs-hom}.

\begin{example}\label{ex:action-functors}
  Given a group homomorphism $f : G \to H$,
  we have three functors
  \[
    \begin{tikzcd}
      \GSet[G] \ar[r,bend left,"f_!"]\ar[r,bend right,"f_*"'] &
      \GSet[H] \ar[l,"f^*"' description]
    \end{tikzcd}
  \]
  with actions on objects described in~\cref{def:restrictandinduce,rem:coinduced-Hset}.
  The action on arrows of restriction is again given by restriction:
  If $g : X \to Y$ is a map of $H$-sets with $X,Y:\BH \to \Set$,
  then $f^*(g) : f^*X \to f^*Y$ maps $z:\BG$ to $g_{\Bf(z)} : X(\Bf(z)) \to Y(\Bf(z))$.

  The action on arrows of induction along $f$ takes a map of $G$-sets $g : X \to Y$,
  for $X,Y : \BG\to\Set$,
  to the functorial action of set truncation, for $w:\BH$:
\end{example}
\begin{example}\label{def:n-trunc-functor}
  Taking $n$-truncation gives a wild functor
  $\Trunc\blank_n : \UU \to \UU^{\le n}$,
  with action on arrows given by the universal property.\footnote{%
    Here, $\UU^{\le n}$ is the full subcategory (\cref{def:full-subcat})
    of $\UU$ on the $n$-truncated types.}
  For $n=0$, this is a functor from $\UU$ to $\Set_\UU$.
\end{example}
\begin{example}\label{ex:add-remove-basepoint}
  We can extend the operation of adding a default element
  from~\cref{def:pointedtypes}
  to a wild functor $(\blank)_+ : \UU \to \UUp$.
  It takes a function $f : A \to B$
  to the function $f_+ : A_+ \to B_+$
  with
  \[
    f_+(\inl a) \defeq \inl{f(a)},\quad\text{and}\quad
    f_+(\pt_{A_+}) \defeq \pt_{B_+}.
  \]
  The operation of taking underlying types of pointed types
  likewise extends to a wild functor $(\blank)_\div : \UUp \to \UU$.
\end{example}
\begin{example}\label{ex:loop-functor}
  Taking loop types extends to a wild functor $\Omega : \UUp \to \UUp$.
  We defined the action on maps in~\cref{def:loops-map},
  except we didn't equip $\Omega k$ with a pointing path, for $k : X \ptdto Y$.
  However, that's easily remedied using the path groupoids
  laws from back in~\cref{xca:path-groupoid-laws}:
  \[
    \pt_{\Omega Y} \jdeq \refl{\pt_Y}
    \eqto \inv{k_\pt} \cdot \refl{k(\pt_X)} \cdot k_\pt
    \jdeq \inv{k_\pt} \cdot \ap{k_\div}(\pt_{\Omega X}) \cdot k_\pt
    \jdeq \Omega k(\pt_{\Omega X})
  \]
  We leave it to the reader to fill in the remaining data.
\end{example}
\begin{example}\label{ex:slice-projection}
  For an object $C$ of a precategory $\mathcal C$,
  recall the slice precategory
  $\mathcal C/C$ of~\cref{def:slice-cat}.
  Taking the domain of an object $(A,f:A\to C)$ of the slice
  extends to a functor $\fst : \mathcal C/C \to \mathcal C$.
\end{example}
\begin{example}\label{ex:arrow-cat}
  Similar to the slice and coslice constructions,
  we have the formation of the \emph{arrow precategory} $\mathcal C^\to$
  of a precategory $\mathcal C$.
  It has as objects triples $(A,B,f)$ of two objects $A,B:\mathcal C$
  and an arrow $f: A\to B$.
  The arrows from $(A,B,f)$ to $(A',B',f')$ are pairs of arrows
  $g : A \to A'$ and $h : B\to B'$ making a commutative square:
  \[
    \begin{tikzcd}
      A \ar[d,"f"']\ar[r,"g"] & A'\ar[d,"f'"] \\
      B \ar[r,"h"'] & B'
    \end{tikzcd}
  \]
  Projecting out the domain and the codomain gives
  two functors
  \[
    \begin{tikzcd}
      \mathcal C & \mathcal C^\to \ar[l,"\dom"']\ar[r,"\cod"] & \mathcal C
    \end{tikzcd}
  \]
  Note that $\mathcal C^\to$ is a category if $\mathcal C$ is.
\end{example}
\begin{xca}\label{xca:wildprecat-of-wildprecats}
  Define identity wild functors and composition of wild functors,
  along with identifications $\lambda : \id_{\mathcal D}\circ F \eqto F$,
  $\rho : F \circ \id_{\mathcal C} \eqto F$,
  and $\alpha : H\circ (G \circ F) \eqto (H \circ G)\circ F$.
  This defines a wild precategory of (small) wild precategories.
  If you're feeling adventurous, define the fillers for the pentagon diagram~\eqref{eq:pentagon}
  for four composable wild functors
  \[
    \mathcal C_0 \xrightarrow F
    \mathcal C_1 \xrightarrow G
    \mathcal C_2 \xrightarrow H
    \mathcal C_3 \xrightarrow K
    \mathcal C_4.
  \]
  How much easier is this if $C_4$ is non-wild?
\end{xca}
\begin{xca}\label{xca:wildcat-of-precats}
  Show that the wild full subprecategory $\PreCat$ on the precategories of the
  wild precategory of small wild precategories is univalent.
  Conclude that the further wild full subcategory $\Cat$ on the categories
  is univalent as well.\footnote{%
    There is in fact also a wild category of all wild precategories,
    but its notion of arrow is more refined
    than the wild functors of~\cref{def:functor}:
    We would need to include components corresponding to
    the unit and associativity laws as well.}
\end{xca}

\begin{definition}\label{def:full-faithful}
  A functor $F : \mathcal C \to \mathcal D$ is \emph{faithful}
  if all the maps on morphisms, $F : \hom_{\mathcal C}(A, B) \to
  \hom_{\mathcal D}(F(A), F(B))$, are injective,
  and \emph{full} if they are surjective. We say that $F$ is
  \emph{fully faithful} if they are both (and hence equivalences).
\end{definition}
\wip{\begin{example}
    core inclusion is faithful,
    projections can be full, not faithful,
    full subcategory inclusions are fully faithful.
\end{example}}

There's also a notion of arrow \emph{between functors}.
These are called natural transformations.\footnote{%
  \citeauthor{Freyd1964}\footnotemark{} observed that ``categories are what one must define
  in order to define functors, and that functors are what one must define
  in order to define natural transformations.''
  In this sense, natural transformations are at the heart of category theory.}%
\footcitetext{Freyd1964}
\begin{definition}\label{def:nat-trans}
  A \emph{wild natural transformation}\index{natural transformation}
  $\alpha : F \to G$ between
  two wild functors $F,G : \mathcal C \to \mathcal D$ between
  wild precategories $\mathcal C$ and $\mathcal D$
  consists of a family of arrows
  \[
    \alpha_A : F(A) \to_{\mathcal D} G(A)
  \]
  indexed by objects $A:\mathcal C$,
  and a family of fillers for the squares
  \[
    \begin{tikzcd}
      F(A) \ar[r,"\alpha_A"]\ar[d,"F(f)"'] & G(A) \ar[d,"G(f)"] \\
      F(B) \ar[r,"\alpha_B"] & G(B)
    \end{tikzcd}
  \]
  in $\mathcal D$
  for each arrow $f : A \to_{\mathcal C} B$.\footnote{%
    These squares are called the ``naturality squares'' for $\alpha$.}

  If $\mathcal D$ is a precategory, then the types of the naturality square fillers
  are propositions. In this case, we just call $\alpha$ a \emph{natural transformation}.
\end{definition}
\begin{example}\label{ex:pt-unpt-unit}
  There is a wild
  natural transformation $\eta : \id_{\UU} \to ((\blank)_+)_\div$
  from the identity wild functor on the universe $\UU$
  to the composition
  \[
    \UU \xrightarrow{(\blank)_+} \UUp \xrightarrow{(\blank)_\div} \UU
  \]
  of the wild functors from~\cref{ex:add-remove-basepoint}.
  Its action on objects is $\inl{} : A \to (A_+)_\div$,
  where $(A_+)_\div \jdeq A \coprod\bn 1$.
  The naturality squares commute by reflexivity.
\end{example}
\begin{example}\label{ex:path-gpd-nat}
  Every function $f : A \to B$ between types in $\UU$
  becomes a wild functor between the corresponding wild path groupoids
  using the action on paths, $\ap f$ from~\cref{def:ap}.

  Likewise, given two functions $f,g : A \to B$, we get for
  every family of identifications $h : \prod_{x:A} f(x) \eqto g(x)$
  a wild natural transformation between the corresponding wild functors,
  using the naturality squares of~\cref{def:naturality-square}.
\end{example}

With natural transformations as arrows we can elevate the
type of functors to a precategory.
\begin{definition}\label{def:functor-cat}
  For a wild precategory $\mathcal C$ and a precategory $\mathcal D$
  we have the \emph{functor precategory}\index{functor category}%
  \index{functor precategory|see{functor category}}
  $\mathcal C \to \mathcal D$, also written
  $[\mathcal C,\mathcal D]$ or $\mathcal D^{\mathcal C}$, % or $\constant{Fun}(\mathcal C,\mathcal D)$,%
  with functors from $\mathcal C$ to $\mathcal D$ as objects
  and natural transformations as arrows.
  The identity arrow at $F$ is the identity natural transformation
  $\id_F: F \to F$ that assigns to each object $A:\mathcal C$ the identity
  arrow $\id_{F(A)}$.
  The composition likewise forms compositions objectwise.
\end{definition}
\begin{xca}\label{xca:funext-nat-trans}
  Show that a natural transformation $\alpha : F \to G$
  in a functor precategory $\mathcal C\to\mathcal D$
  is invertible if and only if
  each component $\alpha_A : F(A) \to_{\mathcal D} G(A)$ is.
\end{xca}
\begin{xca}\label{xca:functor-cat-univalent}
  Show that the functor precategory $\mathcal C\to\mathcal D$
  is univalent if $\mathcal D$ is.\footnote{%
    This is Thm.~9.2.5 in the HoTT book\footnotemark{}.}%
  \footcitetext{hottbook}
  In this case we call it the \emph{functor category}.
\end{xca}
\begin{xca}\label{xca:path-core-adj}
  Let $A$ be a type and $\mathcal C$ category.
  Show that restricting to the action on objects
  induces an equivalence
  \[
    [A, \mathcal C] \to (A \to \Ob(\mathcal C))
  \]
  from the functor category whose domain is the wild path category of $A$
  to the type of functions from $A$ to the objects of $\mathcal C$.
\end{xca}

A functor $F : \mathcal C\op \to \mathcal D$ whose domain is an opposite
category is called \emph{contravariant}, because it reverses the directions
of arrows: $f : C \to_{\mathcal C} C'$ in $\mathcal C$ maps to
$F(f) : F(C') \to_{\mathcal D} F(C)$ in $\mathcal D$.
If needed for emphasis, we may say that a functor
$F : \mathcal C \to \mathcal D$ is \emph{covariant} by contrast.

\begin{example}\label{ex:repr-functors}
  For an object $C$ of a \emph{locally $\UU$-small}
  wild precategory $\mathcal C$, we can form the
  co-
  and contravariant wild
  functors
  \begin{align*}
    k_C \defeq \Hom_{\mathcal C}(C, \blank) &: \mathcal C \to \UU, \\
    h_C \defeq \Hom_{\mathcal C}(\blank, C) &: \mathcal C\op \to \UU,
  \end{align*}
  whose actions on morphisms
  are by post- and precomposition, respectively.
  These are called \emph{representable} functors.
\end{example}

\section{Adjunctions}
\label{sec:adjunctions}

We have already seen two examples of an adjunctions
in~\cref{xca:adjunction-_!-^*,xca:adjunction-^*-_*}.
Given a group homomorphism $f : G \to H$,
there are families of bijections
\begin{align*}
  \alpha &: \Hom_H(f_!\,X,Y) \isoto \Hom_G(X,f^*\,Y), \\
  \beta &: \Hom_G(f^*\,Y,X) \isoto \Hom_H(Y,f_*\,X),
\end{align*}
for $G$-sets $X$ and $H$-sets $Y$
that are furthermore \emph{natural}.
This just means that if we fix either $X$ or $Y$,
then we get natural transformations of the corresponding functors.
For example,
fixing $X$, we can regard $\alpha$ as a natural transformation
\[
  \alpha : \Hom_H(f_!\,X, \blank) \to \Hom_G(X, f^*\,\blank)
\]
from the representable functor for $\GSet[H]$ at $f_!\,X$
to the composition of $f^*$ and the representable functor
for $\GSet$ at $X$.

\begin{definition}\label{def:adjunction}
  A (\emph{wild}) \emph{adjunction} between two (wild) precategories
  $\mathcal C$ and $\mathcal D$ consists of:
  \begin{itemize}
  \item a (wild) functor $F : \mathcal C \to \mathcal D$ (the \emph{left adjoint}),
  \item a (wild) functor $G : \mathcal D \to \mathcal C$ (the \emph{right adjoint}),
  \item a (wild) natural isomorphism $\alpha : \Hom_{\mathcal D}(F\blank,\blank) \isoto
    \Hom_{\mathcal C}(\blank, G\blank)$.
  \end{itemize}
  We write $F \dashv G$ to denote this situation.%
  \glossary(2dashv){$\protect\dashv$}{adjunction}\index{adjunction}
\end{definition}
The essence of an adjunction is thus the ability to
transpose between arrows $F(C) \to_{\mathcal D} D$ and
$C \to_{\mathcal C} G(D)$.
There is however another way of packaging this information.
We can transpose the identity $\id_{F(C)}$
to get an arrow $\eta_C : C \to_{\mathcal C} GF(C)$,
and the identity $\id_{G(D)}$ to get an arrow
$\varepsilon_D : FG(D) \to_{\mathcal D} D$.
Naturality of $\alpha$ makes these into natural transformations
\begin{equation}\label{eq:adj-unit-counit}
  \eta : \id_{\mathcal C} \to GF,\quad\text{and}\quad
  \varepsilon : FG \to \id_{\mathcal D}
\end{equation}
called the \emph{unit}\index{adjunction!unit of}
and \emph{counit}\index{adjunction!counit of} of the adjunction.
\begin{marginfigure}
  \[
    \begin{tikzcd}[ampersand replacement=\&,column sep=small]
      FGF(C) \ar[r,"\varepsilon_{F(C)}"] \& F(C) \\
      F(C) \ar[u,"F(\eta_C)"]\ar[ur,"\id_{F(C)}"'] \& \\[-5mm]
      \& G(D) \\
      G(D) \ar[ur,"\id_{G(D)}"]\ar[r,"\eta_{G(D)}"'] \& GFG(D) \ar[u,"G(\varepsilon_D)"']
    \end{tikzcd}
  \]
  \caption{Triangle laws for an adjunction $F \dashv G$.}
  \label{fig:adj-triangles}
\end{marginfigure}
\begin{xca}\label{xca:adj-triangles}
  Use naturality of $\alpha$, along with unit laws to fill the
  triangle laws in~\cref{fig:adj-triangles}.
\end{xca}
Conversely, given $F$, $G$, $\eta$, and $\varepsilon$, along with
fillers for the triangle laws, we can recover $\alpha$
by sending $f : F(C) \to_{\mathcal D} D$ to
the composite
\[
  C \xrightarrow{\eta_C} GF(C) \xrightarrow{G(f)} G(D).
\]
\begin{xca}\label{xca:adj-from-triangles}
  Use the functor and triangle laws to check that $\alpha$ thus
  defined is a natural isomorphism.\footnote{%
    We can thus go back and forth between
    the different bundles of data for a wild adjunction.
    We shall see later in~\cref{cor:adj-unique} for a functor
    whose domain is a category,
    that the data of a right adjoint is a proposition.}
\end{xca}

\begin{example}\label{ex:trunc-adj}
  We have a wild adjunction $\Trunc\blank_n \dashv \iota_n$
  with $\Trunc\blank_n : \UU \to \UU^{\le n}$ (cf.~\cref{def:n-trunc-functor})
  as the left adjoint
  and the inclusion $\iota_n : \UU^{\le n} \to \UU$ as the right adjoint.
  Indeed, the constructor $\trunc\blank_n$ acts as the unit, precomposition with which
  gives the equivalence
  \[
    (\Trunc X_n \to Y) \equivto (X \to Y)
  \]
  for $X : \UU$ and $Y : \UU^{\le n}$. Naturality follows from the universal property.
\end{example}
\begin{example}\label{ex:pt-unpt-adj}
  Also the add/forget base points functors
  from~\cref{ex:add-remove-basepoint} can be arranged
  into a wild adjunction $(\blank)_+ \dashv (\blank)_\div$.
  For $A : \UUp$ and $X : \UU$ we have
  \[
    \alpha : (A_+ \ptdto X) \equivto (A \to X_\div)
  \]
  given by precomposition with $\inl{} : A \to A_+$.
\end{example}

\begin{example}\label{ex:Sigma-Pi-adj}
  Generalizing~\cref{xca:adjunction-_!-^*,xca:adjunction-^*-_*}
  slightly, we have for any function $f : A \to B$,
  a string of adjunctions $f_! \dashv f^* \dashv f_*$,
  \[
    \begin{tikzcd}
      \Set^A \ar[r,bend left=30,"f_!"]\ar[r,bend right=30,"f_*"']
      \ar[r,bend left=20,phantom,"\scriptscriptstyle\bot"]
      \ar[r,bend right=20,phantom,"\scriptscriptstyle\bot"] &
      \Set^B, \ar[l,"f^*"' description]
    \end{tikzcd}
  \]
  where $f^*(Y) \defeq Y\circ f$, and we set
  \begin{align*}
    f_!(X)(b) &\defeq \setTrunc[\bigg]{\sum_{a:A} (f(a)\eqto b) \times X(a)},\\
    f_*(X)(b) &\defeq \prod_{a:A} (f(a)\eqto b) \to X(a).
  \end{align*}
  Generalizing a bit further, we have a similar string of wild adjunctions
  \[
    \begin{tikzcd}
      \UU^A \ar[r,<-,"f^*"' description]
      \ar[r,bend left=30,"f_!"]\ar[r,bend right=30,"f_*"']
      \ar[r,bend left=20,phantom,"\scriptscriptstyle\bot"]
      \ar[r,bend right=20,phantom,"\scriptscriptstyle\bot"] &
      \UU^B,
    \end{tikzcd}
  \]
  where we leave out the set truncation in the definition of $f_!$.

  For the special case of the terminal projection $! : A \to \bn 1$,
  these simplify to
  \[
    \begin{tikzcd}[column sep=large]
      \UU^A \ar[r,<-,"{\cst{}}"' description]
      \ar[r,bend left=30,"\Sigma_A"]\ar[r,bend right=30,"\Pi_A"']
      \ar[r,bend left=20,phantom,"\scriptscriptstyle\bot"]
      \ar[r,bend right=20,phantom,"\scriptscriptstyle\bot"] &
      \UU,
    \end{tikzcd}
  \]
  where $\Sigma_A(X) \defeq \sum_{a:A}X(a)$ and $\Pi_A(X) \defeq \prod_{a:A}X(a)$.
\end{example}

\begin{remark}\label{rem:adj-in-posets}
  For a monotone map $F : P \to Q$ of preorders,
  we have that $G : Q \to P$ is a right adjoint if and only if
  we have the logical biimplication
  \[
    F(p) \le q \leftrightarrow p \le G(q).
  \]
  For example, the inclusion of the integers into the rationals with their usual order,
  $\iota : \ZZ \hookrightarrow \QQ$ has both adjoints,
  the left adjoint being rounding up, and the right adjoint rounding down:
  $\ceil\blank \dashv \iota \dashv \floor\blank$.

  We also have the propositional analog of~\cref{ex:Sigma-Pi-adj},
  using the operations from~\cref{rem:^*-_!-as-(pre)image,rem:coinduced-subset}:
  \[
    \begin{tikzcd}
      \Sub(A) \ar[r,<-,"f^*"' description]
      \ar[r,bend left=30,"\oldexists_f"]\ar[r,bend right=30,"\forall\!_f"']
      \ar[r,bend left=20,phantom,"\scriptscriptstyle\bot"]
      \ar[r,bend right=20,phantom,"\scriptscriptstyle\bot"] &
      \Sub(B),
    \end{tikzcd}
  \]
  for any map $f : A \to B$.
\end{remark}

\section{The Yoneda Lemma}
\label{sec:yoneda}

\begin{definition}\label{def:yoneda}
  For a locally small\footnote{%
    We suppress the universe $\UU$,
    even though any precategory is $\UU$-small
    for some universe $\UU$;
    the point is that we should consider the category
    of $\UU$-small sets for the same universe
    as the one containing the hom-sets.}
  precategory $\mathcal C$,
  taking contravariant representable functors
  (recall \cref{ex:repr-functors}) induces a functor
  \[
    \yo : \mathcal C \to \Set^{\mathcal C\op} ,
    \qquad \yo(C) \defeq h_C \jdeq \Hom_{\mathcal C}(\blank, C).\qedhere
  \]
\end{definition}

\begin{theorem}\label{thm:yoneda-lemma}
  For any locally small precategory $\mathcal C$,
  functor $F : \Set^{\mathcal C\op}$, and $C:\mathcal C$,
  evaluation at $\id_C$ gives an isomorphism
  \[
    \Hom_{\Set^{\mathcal C\op}}(\yo(C), F) \isoto F(C)
  \]
  that is moreover natural in $F$ and $C$.
\end{theorem}

\begin{corollary}\label{cor:yo-ff}
  The Yoneda embedding $\yo : \mathcal C \to \Set^{\mathcal C\op}$
  is fully faithful.
\end{corollary}

\begin{corollary}\label{cor:yo-emb-obj}
  If $\mathcal C$ is a category, then the action on objects of the
  Yoneda embedding is an injection
  $\Ob(\mathcal C) \hookrightarrow \Ob(\Set^{\mathcal C\op})$.
\end{corollary}

\begin{definition}\label{def:repr}
  A functor $F : \Set^{\mathcal C\op}$ is called
  \emph{representable}\index{functor!representable}
  if there exists $C : \mathcal C$ together with an isomorphism
  $\alpha : \yo(C) \isoto F$.
\end{definition}

\begin{corollary}\label{cor:repr-prop}
  Representability is a proposition when $\mathcal C$ is a category, \ie
  the type of pairs $(C,\alpha)$ in~\cref{def:repr} is a proposition.
\end{corollary}

One benefit of the Yoneda lemma is that we can characterize adjoint functors
in terms of representability, and then derive the uniqueness of adjoint functors
in the case of categories.
\begin{lemma}\label{lem:adj-via-repr}
  Suppose $\mathcal C,\mathcal D$ are precategories ($\mathcal D$ could be wild),
  and $F : \mathcal C \to \mathcal D$ is a (wild) functor.
  If we have a map on objects $G : \Ob(\mathcal D) \to \Ob(\mathcal C)$
  along with natural isomorphisms
  $\alpha_D : \Hom_{\mathcal D}(F\blank, D) \isoto \yo(G(D))$
  for each $D : \Ob(\mathcal D)$, then there is a unique
  extension to the data of a wild adjunction, \ie
  extensions of $G$ to a functor $G : \mathcal D \to \mathcal C$,
  such that $\alpha_D$ is natural in $D:\mathcal D$.
\end{lemma}

\begin{corollary}\label{cor:adj-unique}
  If $F : \mathcal C \to \mathcal D$ is a (wild) functor defined on a category
  $\mathcal C$, then the data of a right adjoint $F \dashv G$ is a proposition.
\end{corollary}

\section{Equivalences of categories}

\begin{definition}\label{def:cat-equiv}
  A functor $F : \mathcal C \to \mathcal D$ is an \emph{equivalence of (pre)categories}%
  \index{equivalence!of categories}
  if it is a left adjoint whose unit and counit transformations $\eta$ and $\varepsilon$
  from~\eqref{eq:adj-unit-counit} are isomorphisms.
  We write $\mathcal C \equivto \mathcal D$ for
  the type of equivalences.
\end{definition}
From~\cref{cor:adj-unique} we immediately see the following:
\begin{lemma}\label{lem:cat-equiv-is-prop}
  If $\mathcal C$ is a category, then being an equivalence
  is a proposition for a functor $F : \mathcal C \to \mathcal D$.
\end{lemma}
It's often useful to show that a functor $F : \mathcal C \to \mathcal D$ is an equivalence by giving just a functor $G : \mathcal D \to \mathcal C$
along with natural isomorphisms $\eta : \id_{\mathcal C} \isoto GF$
and $\varepsilon : FG \isoto \id_{\mathcal D}$.
This is analogous to~\cref{lem:weq-iso}, and we refer to~\textcite[Lemma~6.2]{AKS2015}
for the implementation.

\begin{lemma}\label{lem:cat-equiv-ff}
  Any equivalence of precategories is fully faithful.
\end{lemma}

\begin{definition}\label{def:functor-split-eso}
  A functor $F : \mathcal C \to \mathcal D$ is
  \emph{split essentially surjective on objects}
  if for all objects $D : \mathcal D$ we have a pair of an object $C:\mathcal C$
  and a isomorphism $h : F(C) \isoto D$.

  We say that $F$ is \emph{essentially surjective on objects} if for all $D$,
  we merely have such a pair.
\end{definition}

\begin{lemma}\label{lem:equiv-precat-is-ff-split-eso}
  For precategories $\mathcal C$ and $\mathcal D$
  and a functor $F : \mathcal C \to \mathcal D$,
  the map from the type
  \begin{enumerate}
  \item $F$ is an equivalence of precategories
  \end{enumerate}
  to the type
  \begin{enumerate}[resume]
  \item $F$ is fully faithful and split essentially surjective on objects,
  \end{enumerate}
  defined using~\cref{lem:cat-equiv-ff} and sending $D:\mathcal D$
  to the pair $(G(D), \varepsilon_D)$,
  is an equivalence.
\end{lemma}
We again refer to~\textcite[Lemma~6.6]{AKS2015} for details.

\begin{lemma}\label{lem:ff-eso}
  If $F : \mathcal C \to \mathcal D$ is fully faithful and $\mathcal C$
  is a category, then for any $D : \mathcal D$ the type
  $\sum_{C:\mathcal C}F(C) \isoto D$ is a proposition.
\end{lemma}
From this we conclude that if $\mathcal C$ is category,
then split essential surjectivity
is equivalent to essential surjectivity,
and then~\cref{lem:equiv-precat-is-ff-split-eso}
gives one of the fundamental theorems of category theory:
\begin{theorem}\label{thm:ff-eso-equiv}
  If $\mathcal C$ is a category, then the projection
  $(\mathcal C\equivto \mathcal D) \to (\mathcal C \to \mathcal D)$
  is an injection with image given
  by the fully faithful and essentially surjective functors.\footnote{%
    Compare~\cref{lem:inj+surj}, stating that a function
    $f : A \to B$ of types is an equivalence if and only if it is
    injective and surjective.}
\end{theorem}
\begin{corollary}\label{cor:Cat-ua}
  For categories $\mathcal C$ and $\mathcal D$ we have that
  the map $(\mathcal C \eqto \mathcal D) \to (\mathcal C \equivto \mathcal D)$
  that sends $\refl{\mathcal C}$ to the identity equivalence on $\mathcal C$
  is an equivalence.\footnote{%
    There is also a version of this statement for precategories,
    but it concerns the stronger notion of \emph{isomorphism of precategories},
    \ie fully faithful functors that induce an equivalence on the types of objects.
    The point is that for categories, this is the same as equivalence.}
\end{corollary}

Since the type of functors between categories is $1$-truncated,
we conclude as well:
\begin{corollary}\label{cor:cat-2-type}
  The type of categories $\Cat$ is $2$-truncated.
\end{corollary}

This result is one of the reasons that the primary objects
of interest in category theory are the categories (\ie
univalent precategories), rather than the more general precategories.

However, it's occasionally useful to present a category
via a non-univalent precategory.
This will not be \emph{equivalent} to the corresponding category,
so we need a notion that establishes their connection.
\begin{definition}\label{def:we-cat}
  A functor $F : \mathcal C \to \mathcal D$ of precategories
  is a \emph{weak equivalence} if it is fully faithful
  and essentially surjective on objects.
\end{definition}
From the point of view of categories, however,
weak equivalences are as good as equivalences.
\begin{lemma}\label{lem:precomp-equiv-cat}
  If $H : \mathcal C \to \mathcal D$ is a weak equivalence of precategories,
  and $\mathcal E$ is a category,
  then $\blank\circ H : \mathcal E^{\mathcal D} \to \mathcal E^{\mathcal C}$
  is an equivalence of categories.
\end{lemma}
Our goal now is to construct for every precategory $\mathcal C$
a category $L(\mathcal C)$ along with a weak equivalence
$\eta : \mathcal C \to L(\mathcal C)$.
Here, $L(\mathcal C)$ is called the \emph{Rezk completion} of $\mathcal C$.

The idea is a generalization of the Cayley's~\cref{lem:allgpsarepermutationgps},
which is also how we'll go from abstract groups to concrete groups
in~\cref{ch:absgroup}.
The reader may want to read that first,
and then come back here before reading the comparison
between Rezk completion and delooping of abstract groups
in~\cref{sec:Rezk-delooping}.

Note that $\Set$ is a category,
and hence so is the functor category $\Set^{\mathcal C\op}$
by~\cref{xca:functor-cat-univalent}.
From~\cref{cor:yo-ff} we know that the Yoneda embedding
$\yo : \mathcal C \to \Set^{\mathcal C\op}$ is fully faithful,
motivating the following definition.
\begin{definition}\label{def:Rezk-completion}
  For any precategory $\mathcal C$,
  we define its Rezk completion\index{Rezk completion}
  as the full subcategory of $\Set^{\mathcal C\op}$
  on the representable functors,
  $L(\mathcal C)\defeq (\Set^{\mathcal C\op})_{\constant{isRepr}}$.
  We define $\eta : \mathcal C \to L(\mathcal C)$%
  \glossary(LC){$L(\mathcal C)$}{Rezk completion, \cref{def:Rezk-completion}}
  to be the factorization of $\yo$ via the trivial proofs that any $\yo(C)$
  is representable.
\end{definition}
\begin{theorem}\label{thm:Rezk-completion}
  The inclusion $\eta:\mathcal C \to L(\mathcal C)$ is a weak equivalence.
  Furthermore, if $\mathcal C$ is essentially $\UU$-small,
  then so $L(\mathcal C)$.
\end{theorem}
\begin{proof}
  We have that $\eta$ is fully faithful since $\yo$ is,
  and it's essentially surjective by definition.
  Suppose now that $\mathcal C$ is essentially $\UU$-small,
  meaning that the type of objects $\Ob(\mathcal C)$
  and all hom-sets are essentially $\UU$-small.
  Then we have the hom-functor
  $\hom_{\mathcal C} : \mathcal C\op \times \mathcal C \to \Set_\UU$,
  so we get a Yoneda embedding $\mathcal C \to \Set_\UU^{\mathcal C\op}$.
  Since $\Ob(\mathcal C)$ is essentially $\UU$-small,
  then $\Set_\UU^{\mathcal C\op}$ is locally $\UU$-small,
  and hence so is $L(\mathcal C)$.
  And now $\Ob(L(\mathcal C))$ is the image of
  $\Ob(\yo) : \Ob(\mathcal C) \to \Ob(\Set_\UU^{\mathcal C\op})$,
  which is then essentially $\UU$-small
  by the Replacement~\cref{pri:replacement}.
\end{proof}
\begin{xca}\label{xca:Rezk-completion-locally-small}
  Show that if $\mathcal C$ is a locally $\UU$-small,
  then so is $L(\mathcal C)$, even though $\Set_\UU^{\mathcal C\op}$
  is only locally $\UU$-small when $\mathcal C$ is essentially $\UU$-small
  (\citeauthor{FreydStreet1995}\footcite{FreydStreet1995}).
\end{xca}
% Solution: by Yoneda, Hom(\yo A, \yo B) is \UU-small

Another consequence of~\cref{cor:Cat-ua} is the following.
\begin{theorem}\label{thm:1types-are-groupoids}
  For $1$-types $A$ and $B$,
  the projection from the functor groupoid
  $[A,B]$, of the corresponding path groupoids,
  to the function type $(A\to B)$ is an equivalence.
  Consequently, the path groupoid construction (\cref{ex:path-groupoid})
  is an equivalence $\UU^{\le1} \to \Gpd$,
  where $\Gpd$ is the subtype of groupoids in $\Cat$.
\end{theorem}
\begin{proof}
  Let's write out the definition of the type of functors
  in the case of path groupoids:
  \begin{align*}
    [A,B] \jdeq &\sum_{f : A \to B}
    \sum_{g : \prod_{x,y:A}(x\eqto y) \to (f(x) \eqto f(y))}
    \biggl(\prod_{x:A} g(\refl x) = \refl{f(x)}\biggr) \\
    &\;\times \biggl(\prod_{x,y,z:A}\prod_{p:x\eqto y}\prod_{q:y\eqto z}
    g(qp) = g(q)g(p)\biggr)
  \end{align*}
  Here, the two bracketed components are propositions.
  By path induction, $g$ is determined by its action on reflexivity,
  so the first bracketed component nails it down,
  so only the first component, the function $f:A\to B$,
  remains up to equivalence.

  As a first consequence, the projection from the type of
  equivalences of groupoids, $A \equivto B$,
  to the type of equivalences of types, confusingly also denoted
  $A \equivto B$, is an equivalence. (Hence the confusion is harmless.)

  This shows that $\UU^{\le 1} \to \Gpd$ is injective, so by~\cref{lem:inj+surj},
  it remains to show it is surjective. Observe that for a groupoid,
  the hom-sets and the identity arrows
  are determined by the univalence condition. And relative to this,
  the composition operation is determined (by path induction) by its
  values at $\refl{y}p$, for all $p:x\eqto y$.
  But these values are forced to be $p$ by the left unit law.
  This identifies any groupoid with the path groupoid on its objects,
  showing that $\UU^{\le 1} \to \Gpd$ is surjective.
\end{proof}
\begin{xca}\label{xca:Rezk-completion-1trunc}
  Show that for any type $A$, we have a path pregroupoid
  with $\hom(x,y) \defeq \setTrunc{x\eqto y}$.
  Identify its Rezk completion with the $1$-truncation $\Trunc A_1$.
\end{xca}

% straightening–unstraightening example
Finally, let us use the Rezk completion to explore
the relationship between precategories and categories more deeply.
For example, we emphasize the fundamental importance of categories,
but as mentioned, (non-univalent) precategories do appear from time to time.
Can we think of them just in terms of categories?
It turns out we can.
The idea is that any precategory $\mathcal C$ presents a category $L(\mathcal C)$,
its Rezk completion, but it has more structure,
namely its type of objects, $\Ob(\mathcal C)$,
as well as the relation between these and those of $L(\mathcal C)$,
given by the functor $\eta : \mathcal C \to L(\mathcal C)$.
Since $\eta$ is fully faithful, this structure boils
down to the map on objects, $\Ob(\eta) : \Ob(\mathcal C) \to \Ob(L(\mathcal C))$.
This motivates the following definition.
\begin{definition}\label{def:flagged-cat}
  A \emph{flagged category}\index{category!flagged}\index{flagged category}
  is a triple $(\mathcal C,A,f)$ of a category $\mathcal C$, a type $A$,
  and a surjection $f : A \to \Ob(\mathcal C)$.\footnote{%
    The terminology is a bit whimsical:
    it's meant to indicate that we have a fat ``flag'' object, the category $\mathcal C$,
    equipped with a thinner ``flagpole'' object, the (higher) groupoid $A$,
    connected by the function $f$:\\
    \begin{center}
    \begin{tikzpicture}
      \draw[fill=casblue!30] (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
      \node at (.5,.5) {$\mathcal C$};
      \draw[fill=casred!30] (-.5,-1) -- (-.6,-1) -- (-.6,1) -- (-.5,1) -- cycle;
      \node at (-.8,0) {$A$};
      \draw[->] (-.4,.5) -- (-.1,.5);
      \node at (-.25,.7) {$f$};
    \end{tikzpicture}
    \end{center}
  }
  We define the type of flagged categories to be
  \[
    \FlCat \defeq \sum_{\mathcal C:\Cat}
    \sum_{A:\UU}\sum_{f:A\to\Ob(\mathcal C)}\issurj(f).\qedhere
  \]
\end{definition}
The flagged categories describe the fibers
of the Rezk completion.
\begin{theorem}\label{thm:flagged-cat-equiv}
  The map $F : \PreCat \to \FlCat$ sending a category $\mathcal C$
  to the triple $(L(\mathcal C), \Ob(\mathcal C), \Ob(\eta))$
  is an equivalence.\footnote{%
    We conclude that the following triangle commutes definitionally,
    displaying the fiber of Rezk completion
    at a category $\mathcal C$ as the type
    $\sum_{A:\UU}\sum_{f:A\to\Ob(\mathcal C)}\issurj(f)$:
    \[
      \begin{tikzcd}[column sep=small,ampersand replacement=\&]
        \PreCat\ar[dr,"L"']\ar[rr,equivr,"F"] \& \&
        \FlCat \ar[dl,"\fst"] \\
        \& \Cat \&
      \end{tikzcd}
    \]}
\end{theorem}
\begin{proof}
  The inverse at a triple $(\mathcal C,A,f)$
  ``pulls back'' the (pre)category structure $\mathcal C$
  along $f$ to get a precategory $\mathcal A$ on the object type $A$ and
  \[
    \hom(x,y) \defeq \hom_{\mathcal C}(f(x), f(y)),
  \]
  and likewise using $f$ to define identity arrows, composition, as well
  as unit and associativity laws. By construction, this extends $f$
  to a fully faithful functor $f : \mathcal A \to \mathcal C$.
  This is also essentially surjective by the surjectivity of $f$,
  hence inducing an equivalence $L(\mathcal A) \equivto \mathcal C$.

  For the other round trip, starting with a precategory $\mathcal C$,
  we use the fully faithfulness of $\eta : \mathcal C \to L(\mathcal C)$
  to identify the precategory structure on $\Ob(\mathcal C)$ with that
  pulled back from $L(\mathcal C)$.
\end{proof}
\begin{remark}\label{rem:chicken-or-egg}
  We have seen that precategories and categories
  are connected by an injection and retraction $L$:
  \[
    \begin{tikzcd}[column sep=large]
      \Cat\ar[r,bend left,hook] &
      \PreCat\ar[l,bend left,"L"]
    \end{tikzcd}
  \]
  From the point of view of precategories,
  categories are precategories with an extra property
  (univalence), while from the point of view of categories,
  precategories are categories equipped with extra (higher) structure
  (the flagged structure).
\end{remark}

\section{Limits and colimits}
\label{sec:limits}

We have also seen two instances of limits and colimits,
namely terminal and initial objects
(\cref{def:terminal-obj,def:initial-obj}, respectively).
We now build up to the general notions by first looking at
another dual pair of examples: products and coproducts.

\begin{marginfigure}
  \footnotesize\raggedright(1)
  $\begin{tikzcd}
    & P \ar[dl,"p_1"']\ar[dr,"p_2"] & \\
    A & & B
  \end{tikzcd}$\\
  (2)
  $\begin{tikzcd}
    & X \ar[ddl,"x_1"',bend right=5]\ar[ddr,"x_2",bend left=5]
    \ar[d,dashed,"f"] & \\[4mm]
    & P \ar[dl,"p_1"]\ar[dr,"p_2"'] & \\
    A & & B
  \end{tikzcd}$
  \caption{(1) A product cone, and (2) its universal property.}
  \label{fig:product-cone}
\end{marginfigure}
\begin{definition}\label{def:prod-in-a-cat}
  In a wild precategory $\mathcal C$, a \emph{product cone}
  for a pair of objects $A$ and $B$ is a triple $(P,p_1,p_2)$
  as in the diagram in~\cref{fig:product-cone}(1),
  satisfying the \emph{universal property of the product},
  namely that for any other cone of the same form $(X,x_1,x_2)$,
  there is a unique arrow $f : X \to P$ with $x_i \eqto p_if$, for $i=1,2$,
  \ie filling the triangles in~\cref{fig:product-cone}(2).
\end{definition}
\begin{xca}\label{xca:product-cone-prop}
  Show that if $\mathcal C$ is univalent, then the type of products cones
  on given objects $A,B$ is a proposition.
\end{xca}

We immediately recognize the product of types $A\times B$,
along with the first and second projections,
as a product cone $(A\times B,\fst,snd)$ in the wild category of types.
Likewise, in the category of groups,
the product $G\times H$ from~\cref{ex:productofgroups}
forms a product cone when equipped with the projection homomorphisms.

We say that a (wild) category \emph{has binary products}
if every pair of objects $A,B$ as a product cone.
In this case we are justified by~\cref{xca:product-cone-prop}
in writing $(A\times B,\fst,\snd)$ for the (unique) product cone
on $A,B$.

\begin{remark}\label{def:product-cones-as-terminal}
  We can recover the notion of product cone from that
  of terminal objects by considering the wild (pre)\emph{category of cones}
  given fixed objects $A,B$. This has a objects the cones
  of~\cref{fig:product-cone}(1) and as arrows
  the filled triangles of~\cref{fig:product-cone}(2).
  A product cone is then simply a terminal cone.
\end{remark}

Like every categorical notion, there is of course a dual notion,
obtained by turning all the arrows around.
\begin{marginfigure}
  \footnotesize\raggedright(1)
  $\begin{tikzcd}
    A & & B \\
    & C \ar[<-,ul,"i_1"]\ar[<-,ur,"i_2"'] &
  \end{tikzcd}$\\
  (2)
  $\begin{tikzcd}
    A & & B \\
    & C \ar[<-,ul,"i_1"']\ar[<-,ur,"i_2"] & \\[4mm]
    & X \ar[<-,uul,"x_1",bend left=5]\ar[<-,uur,"x_2"',bend right=5]
    \ar[<-,u,dashed,"f"'] &
  \end{tikzcd}$
  \caption{(1) A coproduct cocone, and (2) its universal property.}
  \label{fig:coproduct-cone}
\end{marginfigure}
\begin{definition}\label{def:coprod-in-a-cat}
  In a wild precategory $\mathcal C$, a \emph{coproduct cocone}
  for a pair of objects $A$ and $B$ is a triple $(C,i_1,i_2)$
  as in the diagram in~\cref{fig:coproduct-cone}(1),
  satisfying the \emph{universal property of the coproduct},
  namely that for any other cocone of the same form $(X,x_1,x_2)$,
  there is a unique arrow $f : C \to X$ with $x_j \eqto fi_j$, for $j=1,2$,
  \ie filling the triangles in~\cref{fig:product-cone}(2).
\end{definition}
In other words, a coproduct cocone is an initial cocone
under the pair of objects $A,B$.

In the wild category of types, we have coproduct cocones $(A \coprod B, \inl{}, \inr{})$. We haven't yet seen the coproduct of groups;
knowing about products and categorical duality, we'd be bound to discover
them! We shall study them in~\cref{sec:coprod}.

Let's look at another pair of examples: pullbacks and pushouts.

% products and coproducts, pullbacks and pushouts, (co)equalizers, (co)limits
% product and exponential in types

Given two functions $f:B\to D$ and $g:C\to D$ with common target,
the ``pullback'' which we will now define should be thought about as the type of all pairs of elements $(b,c):B\times C$ so that $f(b)\eqto g(c)$.
This construction is important in many situations also beyond group theory.

\begin{definition}
  \label{def:pullback}
  Let $B, C, D$ be types and let $f:B\to D$ and $g:C\to D$ be two maps.
  The \emph{pullback}\index{pullback!in types} of $f$ and $g$ is the type
  \[
    \prod(f,g)\defequi\sum_{(b,c):B\times C}(f(b)\eqto_Dg(c))
  \]
  together with the two projections $\prj_B:\prod(f,g)\to B$
  and $\prj_C:\prod(f,g)\to C$ sending $(b,c,p):\prod(f,g)$ to $b:B$ or $c:C$.
  If $f$ and $g$ are clear from the context,
  we may write $B\times_DC$ instead of $\prod(f,g)$ and summarize the situation by the diagram
  \[
    \begin{tikzcd}
      B\times_DC\ar[r,"\prj_C"]\ar[d,"prj_B"'] & C\ar[d,"g"] \\
      B\ar[r,"f"'] & D.\qedhere
    \end{tikzcd}
  \]
\end{definition}
\begin{xca}
  \label{xca:univpropofpullback}
  Let $f:B\to D$ and $g:C\to D$ be two maps with common target.
  If $A$ is a type show that
  \begin{align*}
    (A\to B)\times_{(A\to D)}(A\to C)\to &(A\to B\times_DC)\\
(\beta,\gamma,p:f\beta\eqto g\gamma)\,\mapsto\,&(a\mapsto (f(a),g(a),p(a):f\beta(a)\eqto g\gamma(a)))
  \end{align*}
  is an equivalence.\footnote{This verifies that our defined pullback is the pullback in the categorical sense:
    if the solid diagram commutes there is a unique dashed arrow making the resulting diagram commute:
    \[
      \begin{tikzcd}[ampersand replacement=\&,column sep=small]
        A\ar[dr,dashed]\ar[drr,bend left]\ar[ddr,bend right] \&\& \\
        \& B\times_DC\ar[r]\ar[d] \& C\ar[d] \\
        \& B\ar[r] \& D
      \end{tikzcd}
    \]}
\end{xca}
In view of \cref{xca:univpropofpullback} we will say that we have a \emph{pullback diagram}\index{pullback diagram}
\[
  \begin{tikzcd}
    A\ar[d,"g'"']\ar[r,"f'"]\pbcorner & C\ar[d,"g"] \\
    B\ar[r,"f"'] & D
  \end{tikzcd}
\]
to indicate that we have an element in $(A\to B)\times_{(A\to D)}(A\to C)$ such that the resulting map
$A\to B\times_DC$ is an equivalence.
This is emphasized with the little $\lrcorner$ symbol in the upper left corner.

\begin{example}
  If $d:\bn 1\to D$ denotes the constant function at $d:D$ and $g:C\to D$ is any map,
  then $\bn 1\times_D C$ is equivalent to the preimage $g^{-1}(d)\jdeq\sum_{b:B}d\eqto g(b)$.%
  \footnote{I.e., the preimage is a pullback:
    \[
      \begin{tikzcd}[ampersand replacement=\&,column sep=small]
        g^{-1}(d)\ar[d]\ar[r]\pbcorner \& C\ar[d,"g"] \\
        \bn 1\ar[r,"d"'] \& D
      \end{tikzcd}
    \]}
\end{example}


\section{Monoidal categories}
\label{sec:monoidal-cats}

% monoidal category = Mon(Cat)
% closed category (no need for HITs)

%% Move to Abstract Groups
% pregroupoids with Ob contr = abstract groups
% 

%%% Local Variables:
%%% mode: LaTeX
%%% latex-block-names: ("lemma" "theorem" "remark" "definition" "corollary" "fact" "properties" "conjecture" "proof" "question" "proposition" "exercise")
%%% TeX-master: "book"
%%% TeX-command-extra-options: "-fmt=macros"
%%% compile-command: "make book.pdf"
%%% End:
